# Energy Supply

## Investments Electricity

```{r 04_energy_supply/Investments Electricity}
items <- c(
#  "Investment|Energy Supply|Electricity|Transmission and Distribution",
  "Investment|Energy Supply|Electricity|Transmission and Distribution|BEV Chargers",
  "Investment|Energy Supply|Electricity|Transmission and Distribution|VRE support",
  "Investment|Energy Supply|Electricity|Transmission and Distribution|Normal",
  "Investment|Energy Supply|Electricity|Storage",
  "Investment|Energy Supply|Electricity|Hydrogen",
  "Investment|Energy Supply|Electricity|Solar",
  "Investment|Energy Supply|Electricity|Wind",
  "Investment|Energy Supply|Electricity|Hydro",
  "Investment|Energy Supply|Electricity|Nuclear",
  "Investment|Energy Supply|Electricity|Geothermal",
  "Investment|Energy Supply|Electricity|Biomass|CCS",
  "Investment|Energy Supply|Electricity|Biomass|w/o CCS",
  "Investment|Energy Supply|Electricity|Gas|CCS",
  "Investment|Energy Supply|Electricity|Gas|w/o CCS",
  "Investment|Energy Supply|Electricity|Oil",
  "Investment|Energy Supply|Electricity|Coal|CCS",
  "Investment|Energy Supply|Electricity|Coal|w/o CCS",
  NULL)

showAreaAndBarPlots(data, items, orderVars = "user", scales = "fixed")
```

## Capacities Electricity
### Mix of Electricity Capacities
```{r 04_energy_supply/Capacities Electricity}
items <- c(
  "Cap|Electricity|Solar|CSP",
  "Cap|Electricity|Solar|PV",
  "Cap|Electricity|Wind|Onshore",
  "Cap|Electricity|Wind|Offshore",
  "Cap|Electricity|Storage|Battery",
  "Cap|Electricity|Hydro",
  "Cap|Electricity|Hydrogen",
  "Cap|Electricity|Geothermal",
  "Cap|Electricity|Biomass",
  "Cap|Electricity|Gas|w/ CC",
  "Cap|Electricity|Gas|w/o CC",
  "Cap|Electricity|Oil|w/o CC",
  "Cap|Electricity|Coal|w/ CC",
  "Cap|Electricity|Coal|w/o CC",
  "Cap|Electricity|Nuclear",  
  NULL
  )
showAreaAndBarPlots(data, items, orderVars = "user", scales = "fixed")
```

### Coal
```{r 04_energy_supply/Cap Coal}
showLinePlots(data, "Cap|Electricity|Coal")
```

### Oil
```{r 04_energy_supply/Cap Oil}
showLinePlots(data, "Cap|Electricity|Oil")
```

### Gas
```{r 04_energy_supply/Cap Gas}
showLinePlots(data, "Cap|Electricity|Gas")
```

### Biomass
```{r 04_energy_supply/Cap Biomass}
showLinePlots(data, "Cap|Electricity|Biomass")
```

### Geothermal
```{r 04_energy_supply/Cap Geothermal}
showLinePlots(data, "Cap|Electricity|Geothermal")
```

### Nuclear
```{r 04_energy_supply/Cap Nuclear}
showLinePlots(data, "Cap|Electricity|Nuclear")
```

### Hydro
```{r 04_energy_supply/Cap Hydro}
showLinePlots(data, "Cap|Electricity|Hydro")
```

### Wind
```{r 04_energy_supply/Cap Wind}
showLinePlots(data, "Cap|Electricity|Wind")
showLinePlots(data, "Cap|Electricity|Wind|Onshore")
showLinePlots(data, "Cap|Electricity|Wind|Offshore")
```

### Solar
```{r 04_energy_supply/Cap Solar}
showLinePlots(data, "Cap|Electricity|Solar")
showLinePlots(data, "Cap|Electricity|Solar|PV")
showLinePlots(data, "Cap|Electricity|Solar|CSP")
```

### Hydrogen
```{r 04_energy_supply/Cap Hydrogen}
showLinePlots(data, "Cap|Electricity|Hydrogen")
```

### Storage Battery
```{r 04_energy_supply/Cap Hydro}
showLinePlots(data, "Cap|Electricity|Storage|Battery")
```


## Capacities Hydrogen
### Mix of Hydrogen Capacities
```{r 04_energy_supply/Capacities Hydrogen}
items <- c(
  "Cap|Hydrogen|Electricity",
  "Cap|Hydrogen|Biomass|w/o CC",
  "Cap|Hydrogen|Biomass|w/ CC",
  "Cap|Hydrogen|Gas|w/ CC",
  "Cap|Hydrogen|Gas|w/o CC",
  "Cap|Hydrogen|Coal|w/ CC",
  "Cap|Hydrogen|Coal|w/o CC",
  NULL
  )
showAreaAndBarPlots(data, items, orderVars = "user", scales = "fixed")
```

### Electricity
```{r 04_energy_supply/Cap Electricity}
showLinePlots(data, "Cap|Hydrogen|Electricity")
```

```{r 04_energy_supply/Cap Electricity in GW_el}
showLinePlots(data, "Cap (GWel)|Hydrogen|Electricity")
```

### Biomass
```{r 04_energy_supply/Cap Biomass}
showLinePlots(data, "Cap|Hydrogen|Biomass")
```

### Biomass w/o CCS
```{r 04_energy_supply/Cap Biomass without CCS}
showLinePlots(data, "Cap|Hydrogen|Biomass|w/o CC")
```

### Biomass w/ CCS
```{r 04_energy_supply/Cap Biomass with CCS}
showLinePlots(data, "Cap|Hydrogen|Biomass|w/ CC")
```

### Coal
```{r 04_energy_supply/Cap Coal}
showLinePlots(data, "Cap|Hydrogen|Coal")
```

### Gas
```{r 04_energy_supply/Cap Gas}
showLinePlots(data, "Cap|Hydrogen|Gas")
```

### Gas w/o CC
```{r 04_energy_supply/Cap Gas without CC}
showLinePlots(data, "Cap|Hydrogen|Gas|w/o CC")
```

### Gas w/ CC
```{r 04_energy_supply/Cap Gas with CC}
showLinePlots(data, "Cap|Hydrogen|Gas|w/ CC")
```

## Capacities Heat
### Mix of Heat Capacities
```{r 04_energy_supply/Capacities Heat}
items <- c(
  "Cap|Heat|Electricity|Heat Pump",
  "Cap|Heat|Biomass",
  "Cap|Heat|Gas",
  "Cap|Heat|Coal",
  NULL
  )
showAreaAndBarPlots(data, items, orderVars = "user", scales = "fixed")
```

### Electricity Heat Pump
```{r 04_energy_supply/Cap Electricity}
showLinePlots(data, "Cap|Heat|Electricity|Heat Pump")
```

### Biomass
```{r 04_energy_supply/Cap Biomass}
showLinePlots(data, "Cap|Heat|Biomass")
```

### Gas
```{r 04_energy_supply/Cap Gas}
showLinePlots(data, "Cap|Heat|Gas")
```

### Coal
```{r 04_energy_supply/Cap Coal}
showLinePlots(data, "Cap|Heat|Coal")
```


## Capacities other carriers
### Capacities Solids
```{r 04_energy_supply/Capacities Solids}
showAreaAndBarPlots(data, c(
  "Cap|Solids|Biomass",
  "Cap|Solids|Coal",
  NULL), tot = "Cap|Solids",
  orderVars = "user", scales = "fixed")
```

### Capacities Liquids
```{r 04_energy_supply/Capacities Liquids}
showAreaAndBarPlots(data, c(
  "Cap|Liquids|Hydrogen",
  "Cap|Liquids|Biomass",
  "Cap|Liquids|Gas",
  "Cap|Liquids|Oil",
  "Cap|Liquids|Coal",
  NULL), tot = "Cap|Liquids",
  orderVars = "user", scales = "fixed")
```

### Capacities Gases
```{r 04_energy_supply/Capacities Gases}
showAreaAndBarPlots(data, c(
  "Cap|Gases|Hydrogen",
  "Cap|Gases|Biomass",
  "Cap|Gases|Gas",
  "Cap|Gases|Coal",
  NULL), tot = "Cap|Gases",
  orderVars = "user", scales = "fixed")
```



## PE Extraction (before trade)
### PE Coal Extraction
```{r 04_energy_supply/PE Coal Extraction}
showLinePlots(data, "Res|Extraction|Coal")
```

### PE Oil Extraction
```{r 04_energy_supply/PE Oil Extraction}
showLinePlots(data, "Res|Extraction|Oil")
```

### PE Gas Extraction
```{r 04_energy_supply/PE Gas Extraction}
showLinePlots(data, "Res|Extraction|Gas")
```

### PE Biomass Production
```{r 04_energy_supply/PE Biomass Production}
items <- c(
  "PE|Production|Biomass|Lignocellulosic|Energy Crops",
  "PE|Production|Biomass|Lignocellulosic|Residues",
  "PE|Production|Biomass|1st Generation",
  NULL)
showAreaAndBarPlots(data, items, tot = "PE|Production|Biomass", orderVars = "user", scales = "fixed")
```

```{r 04_energy_supply/PE Biomass Production - details}
showLinePlots(data, "PE|Production|Biomass|Lignocellulosic|Energy Crops")
showLinePlots(data, "PE|Production|Biomass|Lignocellulosic|Residues")
showLinePlots(data, "PE|Production|Biomass|1st Generation|SugarAndStarch")
showLinePlots(data, "PE|Production|Biomass|1st Generation|Sunflowers_PalmOil_others")
```

### PE Uranium Extraction
```{r 04_energy_supply/PE Uranium Extraction}
showLinePlots(data, "Res|Extraction|Uranium [Energy]")
```
Note: multiply EJ by 7.16 to obtain kt of Uranium


## PE Consumption (after trade)
### PE Fossil
```{r 04_energy_supply/PE Fossil}
tot <- "PE|Fossil"
showLinePlots(data, tot)
showRegiLinePlots(data, tot)

items <- c(
  "PE|Gas|Electricity",
  "PE|Gas|Gases",
  "PE|Gas|Liquids",
  "PE|Gas|Heat",
  "PE|Gas|Hydrogen",
  "PE|Oil|Electricity",
  "PE|Oil|Liquids",
  "PE|Coal|Electricity",
  "PE|Coal|Gases",
  "PE|Coal|Liquids",
  "PE|Coal|Hydrogen",
  "PE|Coal|Heat",
  "PE|Coal|Solids",
  NULL)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")

items <- c(
  "PE|Gas|w/ CC",
  "PE|Gas|w/o CC",
  "PE|Oil|w/ CC",
  "PE|Oil|w/o CC",
  "PE|Coal|w/ CC",
  "PE|Coal|w/o CC",
  NULL)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

### PE Coal
```{r 04_energy_supply/PE Coal}
tot <- "PE|Coal"
items <- c(
  "PE|Coal|Hydrogen",
  "PE|Coal|Electricity",
  "PE|Coal|Heat",
  "PE|Coal|Gases",  
  "PE|Coal|Liquids",  
  "PE|Coal|Solids",
  NULL)

showLinePlots(data, tot)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

### PE Oil
```{r 04_energy_supply/PE Oil}
tot <- "PE|Oil"
items <- c(
  "PE|Oil|Electricity",
  "PE|Oil|Liquids",  
  NULL)

showLinePlots(data, tot)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

### PE Gas
```{r 04_energy_supply/PE Gas}
tot <- "PE|Gas"
items <- c(
  "PE|Gas|Hydrogen|w/ CC",
  "PE|Gas|Hydrogen|w/o CC",  
  "PE|Gas|Electricity|w/ CC",
  "PE|Gas|Electricity|w/o CC",
  "PE|Gas|Heat",  
  "PE|Gas|Gases",  
  "PE|Gas|Liquids|w/ CC",
  "PE|Gas|Liquids|w/o CC",
  NULL)

showLinePlots(data, tot)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

### PE Biomass
```{r 04_energy_supply/PE Biomass}
tot <- "PE|Biomass"
items <- c(
  "PE|Biomass|Hydrogen|w/ CC",
  "PE|Biomass|Hydrogen|w/o CC",
  "PE|Biomass|Electricity|w/ CC",
  "PE|Biomass|Electricity|w/o CC",  
  "PE|Biomass|Heat",
  "PE|Biomass|Gases|w/ CC",
  "PE|Biomass|Gases|w/o CC",  
  "PE|Biomass|Liquids|w/ CC",
  "PE|Biomass|Liquids|w/o CC",
  "PE|Biomass|Solids",
  "PE|Biomass|Biochar",
  NULL)

showLinePlots(data, tot)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

### PE Nuclear
```{r 04_energy_supply/PE Nuclear}
showLinePlots(data, "PE|Nuclear")
```

### PE Geothermal
```{r 04_energy_supply/PE Geothermal}
tot <- "PE|Geothermal"
items <- c(
  "PE|Geothermal|Heat",
  "PE|Geothermal|Electricity",
  NULL)

showLinePlots(data, tot)
showAreaAndBarPlots(data, items, tot = tot, orderVars = "user", scales = "fixed")
```

## SE Mix

### SE Totals 

```{r 04_energy_supply/SE Totals}

tot <- "SE"

items <- c(
  "SE|Electricity",
  "SE|Heat",
  "SE|Hydrogen",
  "SE|Gases",
  "SE|Liquids",
  "SE|Solids",
  NULL)
showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```

### SE Electricity
```{r 04_energy_supply/SE Electricity}

tot <- "SE|Electricity"

items <- c(
  "SE|Electricity|Net Imports",
  "SE|Electricity|Hydrogen",
  "SE|Electricity|Solar|CSP",
  "SE|Electricity|Solar|PV",
  "SE|Electricity|Wind|Onshore",
  "SE|Electricity|Wind|Offshore",
  "SE|Electricity|Hydro",  
  "SE|Electricity|Nuclear",  
  "SE|Electricity|Geothermal",  
  "SE|Electricity|Biomass|w/ CC",
  "SE|Electricity|Biomass|w/o CC",
  "SE|Electricity|Gas|w/ CC",
  "SE|Electricity|Gas|w/o CC",
  "SE|Electricity|Oil",  
  "SE|Electricity|Coal|w/ CC",
  "SE|Electricity|Coal|w/o CC",
  NULL)
showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```



### SE Electricity - Usage
```{r 04_energy_supply/SE Electricity - Usage}

tot <- "SE|Electricity"

items <- c(
  "SE|Input|Electricity|T&D losses",
  "SE|Input|Electricity|Self Consumption Energy System",
  "SE|Input|Electricity|Hydrogen|Synthetic Fuels|Liquids",
  "SE|Input|Electricity|Hydrogen|Synthetic Fuels|Gases",
  "SE|Input|Electricity|Hydrogen|direct FE H2",
  "SE|Input|Electricity|Hydrogen|Electricity Storage",
  "SE|Input|Electricity|Buildings",
  "SE|Input|Electricity|Industry",
  "SE|Input|Electricity|Transport",
  "SE|Input|Electricity|CDR"
)

showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```


### SE Heat
```{r 04_energy_supply/SE Heat}
tot <- "SE|Heat"
items <- c(
  "SE|Heat|Electricity|Heat Pump",
  "SE|Heat|Biomass",
  "SE|Heat|Gas",
  "SE|Heat|Coal",
  NULL)
showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```

### SE Heat - Usage
```{r 04_energy_supply/SE Heat - Usage}

tot <- "SE|Heat"

items <- c(
  "SE|Input|Heat|T&D losses",
  "FE|Industry|Heat",
  "FE|Buildings|Heat")
showAreaAndBarPlots(data, items, tot, scales = "fixed")
```

### SE Hydrogen
```{r 04_energy_supply/SE Hydrogen}

tot <- "SE|Hydrogen"

items <- c(
  "SE|Hydrogen|Net Imports",   
  "SE|Hydrogen|Electricity|VRE Storage Electrolysis",
  "SE|Hydrogen|Electricity|Standard Electrolysis",
  "SE|Hydrogen|Biomass|w/ CC",
  "SE|Hydrogen|Biomass|w/o CC",
  "SE|Hydrogen|Gas|w/ CC",
  "SE|Hydrogen|Gas|w/o CC",
  "SE|Hydrogen|Coal|w/ CC",
  "SE|Hydrogen|Coal|w/o CC",
  NULL
  )
showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```

### SE Hydrogen - Usage
```{r 04_energy_supply/SE Hydrogen - Usage}

tot <- "SE|Hydrogen"

items <- c(
  "SE|Input|Hydrogen|T&D losses",
  "FE|Industry|Hydrogen",
  "FE|Buildings|Hydrogen",
  "FE|Transport|Hydrogen",
  "FE|CDR|Hydrogen",
  "SE|Input|Hydrogen|Electricity|Normal Turbines",
  "SE|Input|Hydrogen|Electricity|Forced VRE Turbines",
  "SE|Input|Hydrogen|Synthetic Fuels|Liquids",
  "SE|Input|Hydrogen|Synthetic Fuels|Gases")
showAreaAndBarPlots(data, items, tot, scales = "fixed")
```



### SE Solids
```{r 04_energy_supply/SE Solids}

tot <- "SE|Solids"

items <- c(
  "SE|Solids|Traditional Biomass",  
  "SE|Solids|Biomass",
  "SE|Solids|Coal")
showAreaAndBarPlots(data, items, tot, scales = "fixed")
```

### SE Solids - Usage
```{r 04_energy_supply/SE Solids - Usage}

tot <- "SE|Solids"

items <- c(
  "SE|Input|Solids|T&D losses",
  "FE|Transport|Solids",  
  "FE|Buildings|Solids",  
  "FE|Non-energy Use|Industry|Solids",  
  "FE|w/o Non-energy Use|Industry|Solids")
showAreaAndBarPlots(data, items, tot, orderVars = "user", scales = "fixed")
```

### SE Liquids

```{r 04_energy_supply/SE Liquids}

tot <- "SE|Liquids"

items <- c(
  "SE|Liquids|Hydrogen",
  "SE|Liquids|Hydrogen|Net Imports",
  "SE|Liquids|Biomass|w/ CC",
  "SE|Liquids|Biomass|w/o CC",
  "SE|Liquids|Fossil|Oil",
  "SE|Liquids|Fossil|Gas|w/ CC",
  "SE|Liquids|Fossil|Gas|w/o CC",  
  "SE|Liquids|Fossil|Coal|w/ CC",
  "SE|Liquids|Fossil|Coal|w/o CC",
  NULL)

showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```

### SE Liquids - Usage
```{r 04_energy_supply/SE Liquids - Usage}

tot <- "SE|Liquids"

items <- c(
  "SE|Input|Liquids|T&D losses",
  "FE|CDR|Liquids",
  "FE|Buildings|Liquids",
  "FE|Non-energy Use|Industry|Liquids",
  "FE|w/o Non-energy Use|Industry|Liquids",
  "FE|Transport|Bunkers|Liquids",
  "FE|Transport|w/o Bunkers|Liquids")
showAreaAndBarPlots(data, items, tot, orderVars = "user", scales = "fixed")
```

### SE Gases
```{r 04_energy_supply/SE Gases}

tot <- "SE|Gases"

items <- c(
  "SE|Gases|Hydrogen",
  "SE|Gases|Biomass|w/ CC",
  "SE|Gases|Biomass|w/o CC",
  "SE|Gases|Fossil|Natural Gas",
  "SE|Gases|Fossil|Coal|w/ CC",
  "SE|Gases|Fossil|Coal|w/o CC",
  NULL)

showAreaAndBarPlots(data, items, tot, orderVars="user", scales = "fixed")
```

### SE Gases - Usage
```{r 04_energy_supply/SE Gases - Usage}

tot <- "SE|Gases"

items <- c(
  "SE|Input|Gases|T&D losses",
  "FE|CDR|Gases",
  "FE|Transport|Bunkers|Gases",
  "FE|Transport|w/o Bunkers|Gases",
  "FE|Buildings|Gases",
  "FE|Non-energy Use|Industry|Gases",
  "FE|w/o Non-energy Use|Industry|Gases")
showAreaAndBarPlots(data, items, tot, orderVars = "user", scales = "fixed")
```


## SE Totals - line plots

### SE Electricity
```{r 04_energy_supply/SE Electricity total line}
showLinePlots(data, "SE|Electricity")
```

### SE Heat
```{r 04_energy_supply/SE Heat total line}
showLinePlots(data, "SE|Heat")
```

### SE Hydrogen
```{r 04_energy_supply/SE Hydrogen total line}
showLinePlots(data, "SE|Hydrogen")
```

### SE Solids
```{r 04_energy_supply/SE Solids total line}
showLinePlots(data, "SE|Solids")
```

### SE Liquids
```{r 04_energy_supply/SE Liquids total line}
showLinePlots(data, "SE|Liquids")
```

### SE Gases
```{r 04_energy_supply/SE Gases total line}
showLinePlots(data, "SE|Gases")
```




## SE Details - line plots


### SE Electricity
```{r 04_energy_supply/SE Electricity detail line}

showLinePlots(data, "SE|Electricity|Coal")
showLinePlots(data, "SE|Electricity|Coal|w/ CC")
showLinePlots(data, "SE|Electricity|Coal|w/o CC", histVars = "SE|Electricity|Coal")
showLinePlots(data, "SE|Electricity|Oil")
showLinePlots(data, "SE|Electricity|Gas")
showLinePlots(data, "SE|Electricity|Gas|w/ CC")
showLinePlots(data, "SE|Electricity|Gas|w/o CC", histVars = "SE|Electricity|Gas")
showLinePlots(data, "SE|Electricity|Biomass")
showLinePlots(data, "SE|Electricity|Geothermal")
showLinePlots(data, "SE|Electricity|Nuclear")
showLinePlots(data, "SE|Electricity|Hydro")
showLinePlots(data, "SE|Electricity|Wind")
showLinePlots(data, "SE|Electricity|Wind|Onshore")
showLinePlots(data, "SE|Electricity|Wind|Offshore")
showLinePlots(data, "SE|Electricity|Solar")
showLinePlots(data, "SE|Electricity|Solar|PV", histVars = "SE|Electricity|Solar")
showLinePlots(data, "SE|Electricity|Solar|CSP")
showLinePlots(data, "SE|Electricity|Hydrogen")
```


### SE Heat
```{r 04_energy_supply/SE Heat detail line}

showLinePlots(data, "SE|Heat|Coal")
showLinePlots(data, "SE|Heat|Gas")
showLinePlots(data, "SE|Heat|Biomass")
showLinePlots(data, "SE|Heat|Electricity|Heat Pump")

```

### SE Hydrogen
```{r 04_energy_supply/SE Hydrogen detail line}

showLinePlots(data,  "SE|Hydrogen|Electricity|Standard Electrolysis")
showLinePlots(data,  "SE|Hydrogen|Electricity|VRE Storage Electrolysis")
showLinePlots(data,  "SE|Hydrogen|Biomass")
showLinePlots(data,  "SE|Hydrogen|Biomass|w/ CC")
showLinePlots(data,  "SE|Hydrogen|Biomass|w/o CC")
showLinePlots(data,  "SE|Hydrogen|Gas")
showLinePlots(data,  "SE|Hydrogen|Gas|w/ CC")
showLinePlots(data,  "SE|Hydrogen|Gas|w/o CC")
showLinePlots(data,  "SE|Hydrogen|Coal")
showLinePlots(data,  "SE|Hydrogen|Coal|w/ CC")
showLinePlots(data,  "SE|Hydrogen|Coal|w/o CC")
showLinePlots(data,  "SE|Hydrogen|Net Imports")

```

### SE Solids
```{r 04_energy_supply/SE Solids detail line}

showLinePlots(data, "SE|Solids|Coal")
showLinePlots(data, "SE|Solids|Biomass")
showLinePlots(data, "SE|Solids|Traditional Biomass")
```

### SE Liquids
```{r 04_energy_supply/SE Liquids detail line}

showLinePlots(data, "SE|Liquids|Fossil|Oil")
showLinePlots(data, "SE|Liquids|Fossil|Coal|w/ CC")
showLinePlots(data, "SE|Liquids|Fossil|Coal|w/o CC")
showLinePlots(data, "SE|Liquids|Fossil|Gas|w/ CC")
showLinePlots(data, "SE|Liquids|Fossil|Gas|w/o CC")
showLinePlots(data, "SE|Liquids|Biomass|w/ CC")
showLinePlots(data, "SE|Liquids|Biomass|w/o CC")
showLinePlots(data, "SE|Liquids|Hydrogen")
```


### SE Gases
```{r 04_energy_supply/SE Gases detail line}

showLinePlots(data,  c("SE|Gases|Fossil|Natural Gas", "SE|Gases|Gas"))
showLinePlots(data, "SE|Gases|Fossil|Coal")
showLinePlots(data, "SE|Gases|Biomass")
showLinePlots(data, "SE|Gases|Waste")

```
