---
title: "Implicit FE price target convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gdx)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r setup, echo=FALSE, include=FALSE}

emptyMsg <- function(msg) {
  cat(msg)
}

# getPlotData() errors when a symbol is entirely absent from the gdx (the price-target
# symbols are only declared when the module is active). Return a 0-row frame instead so
# the nrow() == 0 guards below handle the "not active" case gracefully.
getPlotDataSafe <- function(var, gdx) {
  tryCatch(mip::getPlotData(var, gdx), error = function(e) data.frame())
}

# Build a facet label from all dimension columns except iteration / period / value,
# so the report is robust to the exact gdx set names of the carrier/sector dims.
buildLabel <- function(df) {
  idx <- setdiff(names(df), c("iteration", "ttot", "value"))
  if (length(idx) == 0) return(rep("target", nrow(df)))
  do.call(paste, c(df[idx], sep = " · "))
}

# tolerance hardcoded to 0.05 (5%) in regiCarbonPrice/postsolve.gms.
# Read cm_implicitPriceTarget_tolerance from the gdx if exported, else fall back.
tol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "cm_implicitPriceTarget_tolerance", react = "silent")))
if (length(tol) == 0 || is.na(tol)) tol <- 0.05

dev <- getPlotDataSafe("p47_implicitPrice_dev_iter", params$gdx)
tax <- getPlotDataSafe("p47_implicitPriceTax_iter", params$gdx)

if (nrow(dev) > 0) {
  dev <- dev %>%
    rename("value" = "p47_implicitPrice_dev_iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "ttot" := as.numeric(as.character(.data$ttot))
    )
  dev$label <- buildLabel(dev)
}
if (nrow(tax) > 0) {
  tax <- tax %>%
    rename("value" = "p47_implicitPriceTax_iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "ttot" := as.numeric(as.character(.data$ttot))
    )
  tax$label <- buildLabel(tax)
}

```

The implicit FE price target adjusts a sectoral tax/subsidy on a final-energy carrier
until the achieved price matches the target. The criterion requires the **price
deviation to be below +-`r round(100 * tol, 2)`%** (`|p47_implicitPrice_dev| <= `r tol``)
for every region / sector / carrier / period.

Two cases are *excluded* from the check (see `pm_implicitPrice_ignConv`): a
non-existent price (carrier not used), or a price that has not changed for the last 3
iterations (stuck). Those show up in the final-iteration summary at the bottom.

## Convergence summary

Per iteration, the largest absolute FE price deviation across all targets, versus the
tolerance (dashed). Green = every target within tolerance, red = at least one outside.

```{r convsummary, results = "asis"}

if (nrow(dev) == 0) {
  emptyMsg("No implicit FE price target active in this run (p47_implicitPrice_dev_iter is empty).")
} else {
  s <- dev %>%
    group_by(.data$iteration) %>%
    summarise("worst" = max(abs(.data$value)), .groups = "drop") %>%
    mutate(
      "converged" = ifelse(.data$worst > tol, "no", "yes"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>Max |deviation|: ", round(100 * .data$worst, 2), "%<br>",
                         ifelse(.data$worst > tol,
                                paste0("not converged (> ", round(100 * tol, 2), "%)"),
                                "converged"))
    )

  p <- ggplot(s, aes_(x = ~iteration, y = ~worst, group = 1, text = ~tooltip)) +
    geom_hline(yintercept = tol, linetype = "dashed", color = "grey40") +
    geom_line(alpha = 0.6) +
    suppressWarnings(geom_point(aes_(fill = ~converged), size = 2, shape = 21, color = "white")) +
    scale_fill_manual(values = c("yes" = "#00BFC4", "no" = "#F8766D")) +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(x = "iteration", y = "worst |price deviation|")

  ggplotly(p, tooltip = c("text"))
}

```

## Price deviation per target

Deviation over iterations, one panel per region / sector / carrier. Dashed lines mark
the +-`r round(100 * tol, 2)`% tolerance band; colour distinguishes the target period.

```{r deviation, results = "asis"}

if (nrow(dev) == 0) {
  emptyMsg("No implicit FE price target active in this run (p47_implicitPrice_dev_iter is empty).")
} else {
  p <- ggplot(dev, aes_(x = ~iteration, y = ~value, color = ~factor(ttot),
                       group = ~interaction(ttot))) +
    geom_hline(yintercept = c(-tol, tol), linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, color = "grey80") +
    geom_line(alpha = 0.7) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~label, scales = "free_y") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(x = "iteration", y = "price deviation from target", color = "period")

  ggplotly(p)
}

```

## Applied tax / subsidy per target

The implicit tax (positive) or subsidy (negative) applied to reach the target, over
iterations. A converged target has a settled (flat) tax.

```{r tax, results = "asis"}

if (nrow(tax) == 0) {
  emptyMsg("p47_implicitPriceTax_iter is empty in this gdx.")
} else {
  p <- ggplot(tax, aes_(x = ~iteration, y = ~value, color = ~factor(ttot),
                       group = ~interaction(ttot))) +
    geom_hline(yintercept = 0, color = "grey80") +
    geom_line(alpha = 0.7) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~label, scales = "free_y") +
    theme_bw() +
    labs(x = "iteration", y = "implicit tax/subsidy on FE [T$2005/TWyr]", color = "period")

  ggplotly(p)
}

```

## Final-iteration status

Targets flagged in the last iteration: `pm_implicitPrice_NotConv` holds the residual
deviation of targets that did **not** converge; `pm_implicitPrice_ignConv` marks
targets **ignored** by the check (1 = non-existent price, 2 = price frozen 3 iterations).

```{r summary, results = "asis"}

notConv <- getPlotDataSafe("pm_implicitPrice_NotConv", params$gdx)
ignConv <- getPlotDataSafe("pm_implicitPrice_ignConv", params$gdx)

if (nrow(notConv) == 0) {
  cat("**All active FE price targets converged** (pm_implicitPrice_NotConv is empty).\n\n")
} else {
  notConv <- notConv %>%
    rename("deviation" = "pm_implicitPrice_NotConv") %>%
    filter(abs(.data$deviation) > 0) %>%
    mutate("deviation_pct" = round(100 * .data$deviation, 2))
  if (nrow(notConv) == 0) {
    cat("**All active FE price targets converged.**\n\n")
  } else {
    cat("**Not converged targets (residual deviation):**\n\n")
    print(knitr::kable(notConv, row.names = FALSE))
    cat("\n\n")
  }
}

if (nrow(ignConv) > 0) {
  ignConv <- ignConv %>%
    rename("ignore_reason" = "pm_implicitPrice_ignConv") %>%
    filter(.data$ignore_reason > 0)
  if (nrow(ignConv) > 0) {
    cat("**Ignored targets (1 = no price, 2 = price frozen):**\n\n")
    print(knitr::kable(ignConv, row.names = FALSE))
  }
}

```
