---
title: "Regional emission target convergence"
output:
  html_document:
    toc: false
    toc_float: true
    toc_depth: 5
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 16
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
  .code-folding-btn {
    display: none;
  }
  .collapsible-section {
    position: relative;
    margin-bottom: 8px;
  }
  .collapse-btn {
    position: absolute;
    top: 5px;
    right: 5px;
    z-index: 100;
    background: rgba(255,255,255,0.9);
    border: 1px solid #ccc;
    border-radius: 3px;
    cursor: pointer;
    font-size: 11px;
    padding: 2px 6px;
    line-height: 1.4;
    color: #555;
  }
  .collapse-btn:hover {
    background: #f0f0f0;
  }
</style>

<script type="text/javascript">
$(document).ready(function() {
  var $tocItems = $(".tocify-subheader li.tocify-item"); // selector for all TOC items
  var $tabs     = $("a[role=\"tab\"]");                  // selector for all tabs
  var $panels   = $("div[role=\"tabpanel\"]");           // selector for all tabpanels

  $tocItems.hide();  // hide all TOC items

  // get the tab name for each section header (e.g. mammals)
  // and assign it to its corresponding toc item
  $panels.find("div[data-unique]").each(function() {
    var key = $(this).attr("data-unique");
    var tab = $(this).closest("div[role=\"tabpanel\"]").attr("id");
    $tocItems.filter("[data-unique=\"" + key + "\"]").attr("tab", tab)
  })
  // now each toc item knows to which tab panel it is pointing

  // show the toc items that point to sections on the first panel
  $tocItems.filter(function() {
    return($(this).attr("tab") === $tabs.first().text());
  }).toggle();

  // assign an onclick event to the tabs..
  $tabs.on("click", function() {
    $tocItems.hide();  // ... hide all TOC items
    var key = $(this).text(); // ... get the name of the tab clicked
    $tocItems.filter(function() {  // ... filter for the toc items pointing to that tab
      return($(this).attr("tab") === key);
    }).toggle();  // ... and show them
  });
});

function toggleSection(btn) {
  var content = btn.nextElementSibling;
  if (content.style.display === 'none') {
    content.style.display = '';
    btn.innerHTML = '&#9660;';
  } else {
    content.style.display = 'none';
    btn.innerHTML = '&#9654;';
  }
}
</script>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(plotly)
library(mip)
library(gdx)
library(reactable)
library(htmltools)
library(tidyr)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r data, echo=FALSE, include=FALSE}

pm_emiMktTarget_tolerance <-
  mip::getPlotData("pm_emiMktTarget_tolerance", params$gdx)
pm_emiMktTarget_dev <-
  mip::getPlotData("pm_emiMktTarget_dev", params$gdx)
pm_factorRescaleemiMktCO2Tax <-
  mip::getPlotData("pm_factorRescaleemiMktCO2Tax", params$gdx)
pm_emiMktCurrent <- mip::getPlotData("pm_emiMktCurrent", params$gdx)
p47_emiMktCurrent_iter <-
  mip::getPlotData("p47_emiMktCurrent_iter", params$gdx)
pm_emiMktTarget <- mip::getPlotData("pm_emiMktTarget", params$gdx)
pm_emiMktRefYear <- mip::getPlotData("pm_emiMktRefYear", params$gdx)
pm_emiMktTarget_dev_iter <-
  mip::getPlotData("pm_emiMktTarget_dev_iter", params$gdx)
pm_taxemiMkt_iteration <-
  mip::getPlotData("pm_taxemiMkt_iteration", params$gdx)
pm_taxemiMkt <- mip::getPlotData("pm_taxemiMkt", params$gdx)
regi <- gdx::readGDX(params$gdx, "regi")
ext_regi <- gdx::readGDX(params$gdx, "ext_regi")
regi_groupExt <- gdx::readGDX(params$gdx, "regi_groupExt")
regi_group <- gdx::readGDX(params$gdx, "regi_group")
emiMktGroup <- gdx::readGDX(params$gdx, "emiMktGroup")

tgts <-
  pm_emiMktTarget %>% select(ext_regi, ttot2, emiMktExt) %>% mutate_at(c('ttot2'), as.numeric) %>% mutate(group = paste0(ttot2, "_", emiMktExt))
tgtRegions <- unique(tgts$ext_regi)

p47_targetConverged_iter <-
  mip::getPlotData("p47_targetConverged_iter", params$gdx)
p47_emiTargetMkt_iter <-
  mip::getPlotData("p47_emiTargetMkt_iter", params$gdx)
p47_factorRescaleSlope_iter <-
  mip::getPlotData("p47_factorRescaleSlope_iter", params$gdx)
p47_factorRescaleemiMktCO2Tax_iter <-
  mip::getPlotData("p47_factorRescaleemiMktCO2Tax_iter", params$gdx)
#p47_ClampedRescaleSlope <- mip::getPlotData("p47_ClampedRescaleSlope", params$gdx)
regiEmiMktRescaleType <-
  gdx::readGDX(params$gdx, "regiEmiMktRescaleType")

emiMktTarget_tolerance <-
  setNames(
    pm_emiMktTarget_tolerance$pm_emiMktTarget_tolerance,
    pm_emiMktTarget_tolerance$ext_regi
  )

tryGetPlotData <- function(name, gdx) {
  tryCatch(mip::getPlotData(name, gdx), error = function(e) data.frame())
}
p47_slopeReferenceIteration_iter     <- tryGetPlotData("p47_slopeReferenceIteration_iter",     params$gdx)
p47_slopeUpperClampBound_iter        <- tryGetPlotData("p47_slopeUpperClampBound_iter",        params$gdx)
p47_upperClampActive_iter            <- tryGetPlotData("p47_upperClampActive_iter",            params$gdx)
p47_clampedRescaleSlope_iter         <- tryGetPlotData("p47_clampedRescaleSlope_iter",         params$gdx)
p47_dampedFactorRescaleemiMktCO2Tax_iter <- tryGetPlotData("p47_dampedFactorRescaleemiMktCO2Tax_iter", params$gdx)

```

```{r initializeCharts, echo=FALSE, include=FALSE}

#aesthetics
booleanColor <- c("yes" = "#00BFC4", "no" = "#F8766D")
qttyPriceColor <- c("qtty" = "#00BFC4", "price" = "#F8766D")
methodPalette <- c(
  "sq·1st"   = "#e41a1c", "sq·match" = "#ff7f00", "sq·small" = "#f7c842",
  "sq·win"   = "#984ea3", "sq·+sl"   = "#a65628", "sq·clamp" = "#f781bf",
  "sq·degen" = "#aaaaaa",
  "sl·ref"   = "#1f78b4", "sl·1st"   = "#377eb8", "sl·rep"   = "#00BFC4",
  "converged"= "#93c47d"
)

g <- NULL

```



<!-- emissions and prices -->
```{r emiAndPrice chart, echo=FALSE, include=FALSE}

#Convert regi to ext_regi and vice-versa
convGroupExt <- NULL
for (tgtReg in unique(tgts$ext_regi)) {
  if (tgtReg %in% regi)
    convGroupExt <- c(convGroupExt, setNames(tgtReg, tgtReg))
  else
    convGroupExt <-
      c(convGroupExt,
        setNames(
          regi_groupExt %>% filter(ext_regi == tgtReg,!(all_regi %in% tgtReg)) %>% slice(1) %>% pull(all_regi),
          tgtReg
        ))
}

emissions <- p47_emiMktCurrent_iter %>%
  rename(emissions = p47_emiMktCurrent_iter) %>%
  mutate(emissions = emissions * 3.66667 * 1000) %>%
  select(-ttot)

convEmiMktGroup <-
  setNames(emiMktGroup$all_emiMktExt, emiMktGroup$all_emiMkt)
tmpPrice <- pm_taxemiMkt_iteration %>%
  mutate(pm_taxemiMkt_iteration = pm_taxemiMkt_iteration * 1000 * 12 / 44) %>%
  filter(all_regi %in% convGroupExt)
price <- rbind(
  tmpPrice %>%
    mutate(
      all_emiMktExt = convEmiMktGroup[all_emiMkt],
      ext_regi = setNames(names(convGroupExt), convGroupExt)[all_regi]
    ),
  tmpPrice %>%
    filter(all_emiMkt == "ETS") %>%
    mutate(
      all_emiMktExt = "all",
      ext_regi = setNames(names(convGroupExt), convGroupExt)[all_regi]
    )
) %>%
  rename(ttot2 = ttot,
         #ext_regi = all_regi,
         emiMktExt = all_emiMktExt) %>%
  select(-all_emiMkt,-all_regi)

plotData <-
  left_join(
    left_join(
      left_join(
        left_join(
          left_join(
            left_join(
              emissions,
              price,
              by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
            ),
            regiEmiMktRescaleType %>% select(-ttot) %>% rename(ttot2 = ttot1) %>% mutate(iteration = as.integer(iteration)),
            by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
          ),
          p47_factorRescaleSlope_iter %>% select(-ttot),
          by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
        ),
        p47_factorRescaleemiMktCO2Tax_iter %>% select(-ttot),
        by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
      ),
      p47_targetConverged_iter %>% rename(ttot2 = ttot),
      by = c("iteration", "ttot2", "ext_regi")
    ),
    pm_emiMktTarget_dev_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  mutate(
    tooltip = paste0(
      "Iteration : ",
      iteration,
      "<br>",
      "Emissions : ",
      round(emissions),
      " Mt CO2<br>",
      "Price : ",
      round(pm_taxemiMkt_iteration),
      " US$2005/t CO2<br>",
      "Convergence : ",
      ifelse(
        is.na(p47_targetConverged_iter),
        ifelse(
          is.na(rescaleType),
          "not active",
          paste0(
            "not converged<br>",
            "Deviation: ",
            round(100 * pm_emiMktTarget_dev_iter, 2),
            "%",
            ifelse(pm_emiMktTarget_dev_iter >= 0, " > ", " < -"),
            round(100 * emiMktTarget_tolerance[ext_regi], 2),
            "%<br>",
            "Rescale for next iteration: ",
            round(p47_factorRescaleemiMktCO2Tax_iter, 2),
            "<br>",
            "Rescale Type : ",
            ifelse(
              rescaleType %in% c("slope_prevIteration", "slope_firstIteration"),
              paste0(
                rescaleType,
                " (slope=",
                round(p47_factorRescaleSlope_iter, 2),
                ")"
              ),
              rescaleType
            )
          )
        ),
        "converged"
      )
    )
  )

g[["qtty_and_price"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    subGroup <-
      setNames(
        tgts %>% filter(ext_regi == reg) %>% pull(ttot2),
        tgts %>% filter(ext_regi == reg) %>% pull(group)
      )
    lapply(subGroup, function(tgtPeriod) {
      d <- plotData %>% filter(ext_regi == reg, ttot2 == tgtPeriod)
      secAxisScale <-
        max(d %>% pull(pm_taxemiMkt_iteration)) / max(d %>% pull(emissions))
      
      p <- plot_ly(data = d, height = 350) %>% #
        add_trace(
          x = ~ iteration,
          y = ~ emissions,
          name = "Emissions",
          mode = "lines",
          type = "scatter",
          line = list(color = qttyPriceColor["qtty"]),
          hoverinfo = "none",
          hoverinfo = 'skip'
        ) %>%
        add_trace(
          x = ~ iteration,
          y = ~ pm_taxemiMkt_iteration,
          name = "Price",
          yaxis = "y2",
          mode = "lines",
          type = "scatter",
          line = list(color = qttyPriceColor["price"]),
          hoverinfo = "none",
          hoverinfo = 'skip'
        ) %>%
        # add trace only for hover info (hack to disable line in hovermode = "x unified"
        add_trace(
          x = ~ iteration,
          y = ~ emissions,
          mode = "lines",
          type = "scatter",
          hoverinfo = "text",
          line = list(width = 0),
          hovertext = d$tooltip
        ) %>%
        layout(
          hovermode = "x unified",
          showlegend = FALSE,
          margin = list(r = 60),
          xaxis = list(
            zeroline = FALSE,
            range = c(0.5, max(d$iteration) + 0.5),
            title = "",
            gridcolor = "#e8e8e8",
            mirror = TRUE,
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          ),
          yaxis = list(
            zeroline = TRUE,
            zerolinecolor = "#e8e8e8",
            zerolinewidth = 3,
            title = list(text = "Emissions (Mt CO2)", font = list(color =
                                                                    qttyPriceColor["qtty"])),
            range = c(min(
              0, min(d %>% pull(emissions)) - max(d %>% pull(emissions)) * 0.05
            ), max(d %>% pull(emissions)) * 1.05),
            gridcolor = "#e8e8e8",
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          ),
          yaxis2 = list(
            zeroline = TRUE,
            zerolinecolor = "#e8e8e8",
            zerolinewidth = 3,
            showgrid = FALSE,
            title = list(text = "Price (US$2005/t CO2)", font = list(color =
                                                                       qttyPriceColor["price"])),
            overlaying = "y",
            side = "right",
            range = c(
              min(
                0,
                min(d %>% pull(emissions)) * secAxisScale - max(d %>% pull(emissions)) *
                  secAxisScale * 0.05
              ),
              max(d %>% pull(emissions)) * secAxisScale * 1.05
            ),
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          )
        )
    })
  })


```


 <!-- deviation -->
```{r deviation chart, echo=FALSE, include=FALSE}

# deviation chart
plotData <- pm_emiMktTarget_dev_iter %>%
  mutate(
    converged = ifelse((.data$pm_emiMktTarget_dev_iter <=  emiMktTarget_tolerance[ext_regi]) &
                         (.data$pm_emiMktTarget_dev_iter >= -emiMktTarget_tolerance[ext_regi]),
                       "yes",
                       "no"
    ),
    tooltip = paste0(
      "Iteration: ",
      .data$iteration,
      "<br>",
      ifelse((.data$pm_emiMktTarget_dev_iter <=  emiMktTarget_tolerance[ext_regi]) &
               (.data$pm_emiMktTarget_dev_iter >= -emiMktTarget_tolerance[ext_regi]),
             paste0(
               "Converged<br>",
               "Deviation: ",
               round(100 * .data$pm_emiMktTarget_dev_iter, 2),
               "%"
             ),
             paste0(
               "Not converged<br>",
               "Deviation: ",
               round(100 * .data$pm_emiMktTarget_dev_iter, 2),
               "%",
               ifelse(pm_emiMktTarget_dev_iter >= 0, " > ", " < -"),
               round(100 * emiMktTarget_tolerance[ext_regi], 2),
               "%"
             )
      )
    )
  )

g[["deviation"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    subGroup <-
      setNames(
        tgts %>% filter(ext_regi == reg) %>% pull(ttot2),
        tgts %>% filter(ext_regi == reg) %>% pull(group)
      )
    lapply(subGroup, function(tgtPeriod) {
      d <- plotData %>% filter(ext_regi == reg, ttot2 == tgtPeriod)
      p <-
        ggplot(d,
               aes_string(x = "iteration", y = "pm_emiMktTarget_dev_iter")) +
        suppressWarnings(geom_rect(
          aes(
            xmin = .data$iteration - 0.5,
            xmax = .data$iteration + 0.5,
            ymin = emiMktTarget_tolerance[ext_regi] * -1,
            ymax = emiMktTarget_tolerance[ext_regi],
            fill = converged,
            text = tooltip
          ),
          alpha = 0.2
        )) +
        suppressWarnings(geom_line(aes_string(group = 1, text = "tooltip"))) +
        scale_y_continuous(
          breaks = c(
            -5 * emiMktTarget_tolerance[ext_regi],
            -emiMktTarget_tolerance[ext_regi],
            0,
            emiMktTarget_tolerance[ext_regi],
            5 * emiMktTarget_tolerance[ext_regi]
          ),
          labels = scales::percent_format(accuracy = .1)
        ) +
        scale_fill_manual(values = booleanColor) +
        theme_bw() +
        theme(
          strip.background = element_blank(),
          axis.title.x = element_blank(),
          legend.position = "none"
        ) +
        coord_cartesian(ylim = c(-5 * emiMktTarget_tolerance[reg],+5 * emiMktTarget_tolerance[reg])) +
        ylab("Deviation (%)") +
        scale_x_continuous(expand = c(0, 0),
                           breaks = 1:max(d$iteration)) +
        theme(plot.margin = margin(0, 42, 0, 22, "pt")) # t,r,b,l
      
      tol_frac         <- emiMktTarget_tolerance[[reg]]
      tol_pct          <- round(100 * tol_frac, 2)
      emiMktExt_first  <- tgts %>% filter(ext_regi == reg, ttot2 == tgtPeriod) %>% slice(1) %>% pull(emiMktExt)
      refYear_val      <- pm_emiMktRefYear %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExt_first) %>%
        slice(1) %>% pull(pm_emiMktRefYear)
      tol_MtCO2 <- if (length(refYear_val) > 0 && !is.na(refYear_val))
        round(tol_frac * refYear_val * 3.66667 * 1000, 1) else NA
      tol_label <- paste0("Tolerance: ±", tol_pct, "%",
                          if (!is.na(tol_MtCO2)) paste0(" (±", tol_MtCO2, " MtCO₂)") else "")

      plotly::ggplotly(p, tooltip = c("text"), height = 200) %>%
        plotly::layout(annotations = list(list(
          text      = tol_label,
          xref      = "paper", yref = "paper",
          x = 0.99, y = 0.02,
          xanchor   = "right", yanchor = "bottom",
          showarrow = FALSE,
          font      = list(size = 10, color = "#666"),
          bgcolor   = "rgba(255,255,255,0.75)"
        ))) %>%
        plotly::config(displayModeBar = FALSE)
    })
  })


```



```{r rescaleTableData, echo=FALSE, include=FALSE}

rescaleTypeAbbrev <- c(
  squareDev_firstIteration       = "sq·1st",
  squareDev_perfectMatch         = "sq·match",
  squareDev_smallChange          = "sq·small",
  squareDev_outsideWindow        = "sq·win",
  squareDev_noNonPositiveSlope   = "sq·+sl",
  squareDev_adaptiveClamp        = "sq·clamp",
  squareDev_degenerateSlope      = "sq·degen",
  slope_refIteration             = "sl·ref",
  slope_firstIteration           = "sl·1st",
  slope_repeatPrev_positiveSlope = "sl·rep"
)

safeJoin <- function(base, extra, by, extra_col_rename = NULL) {
  if (nrow(extra) == 0) return(base)
  if (!is.null(extra_col_rename)) extra <- dplyr::rename(extra, !!extra_col_rename)
  dplyr::left_join(base, extra, by = by)
}

tableData <- pm_emiMktTarget_dev_iter %>%
  select(-ttot) %>%
  mutate(iteration = as.integer(iteration)) %>%
  left_join(
    regiEmiMktRescaleType %>%
      select(-ttot) %>% rename(ttot2 = ttot1) %>%
      mutate(iteration = as.integer(iteration)) %>% select(-any_of("value")) %>%
      group_by(iteration, ttot2, ext_regi, emiMktExt) %>%
      summarise(rescaleType = paste(sort(unique(rescaleType)), collapse = " + "), .groups = "drop"),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  left_join(
    p47_targetConverged_iter %>% rename(ttot2 = ttot),
    by = c("iteration", "ttot2", "ext_regi")
  ) %>%
  left_join(
    p47_factorRescaleSlope_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  left_join(
    p47_factorRescaleemiMktCO2Tax_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_slopeReferenceIteration_iter) > 0) p47_slopeReferenceIteration_iter %>% rename(ttot2 = ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi")
  ) %>%
  safeJoin(
    if (nrow(p47_slopeUpperClampBound_iter) > 0) p47_slopeUpperClampBound_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_upperClampActive_iter) > 0) p47_upperClampActive_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_clampedRescaleSlope_iter) > 0) p47_clampedRescaleSlope_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_dampedFactorRescaleemiMktCO2Tax_iter) > 0) p47_dampedFactorRescaleemiMktCO2Tax_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  arrange(ext_regi, ttot2, emiMktExt, iteration) %>%
  group_by(ext_regi, ttot2, emiMktExt) %>%
  mutate(delta_dev_raw = pm_emiMktTarget_dev_iter - lag(pm_emiMktTarget_dev_iter)) %>%
  ungroup() %>%
  mutate(
    method     = sapply(strsplit(rescaleType, " + ", fixed = TRUE), function(types) {
      codes <- rescaleTypeAbbrev[types]
      paste(codes[!is.na(codes)], collapse = " + ")
    }),
    dev_pct    = round(100 * pm_emiMktTarget_dev_iter, 2),
    converged  = ifelse(!is.na(p47_targetConverged_iter) & p47_targetConverged_iter == 1, "✓", "✗"),
    used_slope = round(p47_factorRescaleSlope_iter, 3),
    final_rsc  = round(p47_factorRescaleemiMktCO2Tax_iter, 3),
    delta_dev  = round(abs(100 * delta_dev_raw), 2)
  )

tableData$slope_ref  <- if ("p47_slopeReferenceIteration_iter" %in% names(tableData)) tableData$p47_slopeReferenceIteration_iter else NA_real_
tableData$raw_slope  <- if ("p47_clampedRescaleSlope_iter" %in% names(tableData)) round(tableData$p47_clampedRescaleSlope_iter, 3) else NA_real_
tableData$clamp_bnd  <- if ("p47_slopeUpperClampBound_iter" %in% names(tableData)) round(tableData$p47_slopeUpperClampBound_iter, 3) else NA_real_
tableData$clamp_act  <- if ("p47_upperClampActive_iter" %in% names(tableData)) {
  ifelse(is.na(tableData$p47_upperClampActive_iter), NA_character_,
         ifelse(tableData$p47_upperClampActive_iter == 1, "✓", "—"))
} else {
  NA_character_
}
tableData$pre_damp   <- if ("p47_dampedFactorRescaleemiMktCO2Tax_iter" %in% names(tableData)) round(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter, 3) else NA_real_
tableData$damped     <- if ("p47_dampedFactorRescaleemiMktCO2Tax_iter" %in% names(tableData)) {
  ifelse(is.na(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter), NA_character_,
         ifelse(abs(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter - tableData$p47_factorRescaleemiMktCO2Tax_iter) > 1e-4, "✓", "—"))
} else {
  NA_character_
}

tableData <- tableData %>%
  select(iteration, ttot2, ext_regi, emiMktExt,
         method, rescaleType, dev_pct, delta_dev, converged,
         slope_ref, raw_slope, clamp_bnd, clamp_act, used_slope,
         pre_damp, final_rsc, damped)

legendRow <- function(code, fullname, what, when) {
  htmltools::tags$tr(
    style = "border-bottom: 1px solid #f0f0f0;",
    htmltools::tags$td(style = "padding: 2px 8px; font-family: monospace; white-space: nowrap;", code),
    htmltools::tags$td(style = "padding: 2px 8px; font-family: monospace; white-space: nowrap; color: #888; font-size: 0.85em;", fullname),
    htmltools::tags$td(style = "padding: 2px 8px;", what),
    htmltools::tags$td(style = "padding: 2px 8px; color: #666;", when)
  )
}

makeRescaleTable <- function(data, reg, tgtPeriod, emiMktExtVal) {
  d <- data %>%
    filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
    arrange(iteration)

  tol_frac    <- emiMktTarget_tolerance[[reg]]
  tol_pct     <- round(100 * tol_frac, 2)
  refYear_val <- pm_emiMktRefYear %>%
    filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
    slice(1) %>% pull(pm_emiMktRefYear)
  tol_MtCO2 <- if (length(refYear_val) > 0 && !is.na(refYear_val))
    round(tol_frac * refYear_val * 3.66667 * 1000, 1) else NA
  caption_text <- paste0(
    "Tolerance: ±", tol_pct, "%",
    if (!is.na(tol_MtCO2)) paste0(" (±", tol_MtCO2, " MtCO2)") else ""
  )

  # ── parameter metadata ────────────────────────────────────────────────────
  param_order  <- c("method","dev_pct","delta_dev","converged",
                    "slope_ref","raw_slope","clamp_bnd","clamp_act","used_slope",
                    "pre_damp","final_rsc","damped")
  param_labels <- c(
    method     = "Method",
    dev_pct    = "Deviation (%)",
    delta_dev  = "|Δdev| (%)",
    converged  = "Converged",
    slope_ref  = "Ref. iter.",
    raw_slope  = "Raw slope",
    clamp_bnd  = "Clamp bound",
    clamp_act  = "Clamp active",
    used_slope = "Used slope",
    pre_damp   = "Pre-damp factor",
    final_rsc  = "Applied factor",
    damped     = "Damped"
  )
  param_groups <- c(
    method="Convergence", dev_pct="Convergence", delta_dev="Convergence", converged="Convergence",
    slope_ref="Slope", raw_slope="Slope", clamp_bnd="Slope",
    clamp_act="Slope", used_slope="Slope",
    pre_damp="Rescale Factor", final_rsc="Rescale Factor", damped="Rescale Factor"
  )
  group_colors <- c(Convergence="#4a9eca", Slope="#e07b39", `Rescale Factor`="#5cb85c")

  # ── transpose: rows = parameters, cols = iterations ───────────────────────
  iter_cols <- as.character(d$iteration)
  d_t <- d %>%
    select(iteration, all_of(param_order)) %>%
    mutate(across(everything(), as.character)) %>%
    tidyr::pivot_longer(cols = -iteration, names_to = "param", values_to = "value") %>%
    tidyr::pivot_wider(names_from = iteration, values_from = value) %>%
    mutate(group = param_groups[param], label = param_labels[param]) %>%
    arrange(match(param, param_order)) %>%
    select(group, param, label, all_of(iter_cols))

  # ── per-iteration lookups for cell styling ─────────────────────────────────
  method_per_iter    <- setNames(d$method,    iter_cols)
  converged_per_iter <- setNames(d$converged, iter_cols)

  # ── column definitions ─────────────────────────────────────────────────────
  label_coldef <- list(
    group = reactable::colDef(show = FALSE),
    param = reactable::colDef(show = FALSE),
    label = reactable::colDef(
      name  = "",
      width = 145,
      style = function(value, index) {
        list(fontWeight  = "600",
             fontSize    = "0.82em",
             borderLeft  = paste0("3px solid ", group_colors[d_t$group[index]]),
             paddingLeft = "6px",
             whiteSpace  = "nowrap")
      }
    )
  )

  iter_coldefs <- lapply(setNames(iter_cols, iter_cols), function(iter) {
    method  <- method_per_iter[[iter]]
    is_conv <- identical(converged_per_iter[[iter]], "✓")
    is_sq   <- !is.na(method) && grepl("^sq", method)

    reactable::colDef(
      name  = iter,
      width = 46,
      align = "center",
      html  = TRUE,
      header = htmltools::div(
        style = paste0("font-weight:bold; font-size:0.88em;",
                       if (is_conv) " color:#c0c0c0;" else ""),
        iter
      ),
      cell = function(value, index) {
        if (is.na(value) || value == "NA")
          return(htmltools::span(style = "color:#e8e8e8;", "—"))
        pname <- d_t$param[index]
        if (pname == "method")
          gsub(" \\+ ", "<br><span style='color:#aaa; font-size:0.85em;'>+</span>",
               value, perl = TRUE)
        else if (pname == "converged")
          if (value == "✓")
            htmltools::span(style = "color:#00BFC4; font-weight:bold;", value)
          else
            htmltools::span(style = "color:#F8766D; font-weight:bold;", value)
        else
          value
      },
      style = function(value, index) {
        pname <- d_t$param[index]
        is_slope_row <- param_groups[pname] == "Slope"
        if (is.na(value) || value == "NA") return(list(color = "#ebebeb"))
        if (is_conv)  return(list(color = "#c8c8c8", fontSize = "0.78em"))
        if (is_slope_row && is_sq) return(list(color = "#d0d0d0"))
        NULL
      }
    )
  })

  tbl <- reactable::reactable(
    d_t,
    columns  = c(label_coldef, iter_coldefs),
    rowStyle = function(index) {
      base <- list(fontSize = "0.84em", lineHeight = "1.4")
      if (index > 1 && d_t$group[index] != d_t$group[index - 1])
        base <- c(base, list(borderTop = "2px solid #d8d8d8", marginTop = "4px"))
      base
    },
    sortable   = FALSE,
    bordered   = FALSE,
    compact    = TRUE,
    highlight  = TRUE,
    pagination = FALSE,
    theme = reactable::reactableTheme(
      headerStyle = list(borderBottom = "2px solid #ccc", fontSize = "0.88em",
                         background = "#eeeeee")
    )
  )

  legend <- htmltools::div(
    style = "font-size:0.8em; color:#444; margin-top:8px;",
    htmltools::tags$details(
      htmltools::tags$summary(
        style = "cursor:pointer; font-weight:bold; color:#555; padding:4px 0;",
        "Method legend"
      ),
      htmltools::tags$table(
        style = "border-collapse:collapse; margin-top:6px; width:100%;",
        htmltools::tags$thead(
          htmltools::tags$tr(
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "Code"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "Full name"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "What it does"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "When it fires")
          )
        ),
        htmltools::tags$tbody(
          legendRow("sq·1st",    "squareDev_firstIteration",       "Rescale = (1 + dev)²",                                            "First iteration, or slope reference just reset — no slope history yet"),
          legendRow("sq·match",  "squareDev_perfectMatch",         "Rescale = (1 + dev)²",                                            "Tax identical to both reference and iteration 1 — division-by-zero guard"),
          legendRow("sq·small",  "squareDev_smallChange",          "Rescale = (1 + dev)²",                                            "Tax change vs both reference and iteration 1 is < 1e−2 — avoids stagnation from negligible price variation"),
          legendRow("sq·win",    "squareDev_outsideWindow",        "Rescale = (1 + dev)²",                                            "Reference near-zero denominator AND using iteration 1 would exceed the slope window"),
          legendRow("sq·+sl",    "squareDev_noNonPositiveSlope",   "Rescale = (1 + dev)²",                                            "Slope is positive (price rise → more emissions) AND no prior slope to repeat — sign assumption violated"),
          legendRow("sq·clamp",  "squareDev_adaptiveClamp",        "Rescale = (1 + dev)²; slope reference reset",                     "Adaptive Upper Clamp Bound fired twice consecutively at same level — slope-based convergence exhausted"),
          legendRow("sq·degen",  "squareDev_degenerateSlope",      "Rescale = (1 + dev)²; slope reference reset",                     "Emission change negligible relative to 2005 emissions — degenerate slope from near-net-zero starting point"),
          legendRow("sl·ref",    "slope_refIteration",             "Rescale = (target − current) / (slope × price) + 1",         "Normal path — slope computed against the current Slope Reference Iteration"),
          legendRow("sl·1st",    "slope_firstIteration",           "Rescale = (target − current) / (slope × price) + 1",         "Reference denominator near-zero, but iteration 1 is within slope window — uses iteration 1 as baseline"),
          legendRow("sl·rep",    "slope_repeatPrev_positiveSlope", "Rescale = (target − current) / (slope × price) + 1 (prev.)", "Computed slope is positive, but prior slope exists with same reference — repeats previous slope to avoid sign anomaly")
        )
      )
    )
  )

  htmltools::tagList(
    tbl,
    legend
  )
}

```

```{r rescaleFactorChart, echo=FALSE, include=FALSE}

g[["rescale_factor"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>% arrange(iteration)

      hasDamp <- !all(is.na(d$pre_damp))
      p <- plot_ly(data = d, height = 280) %>%
        add_segments(x = min(d$iteration) - 0.5, xend = max(d$iteration) + 0.5,
                     y = 1, yend = 1, showlegend = FALSE,
                     line = list(color = "#aaaaaa", dash = "dot", width = 1)) %>%
        add_trace(x = ~iteration, y = ~final_rsc,
                  name = "Final rescale", mode = "lines+markers", type = "scatter",
                  line = list(color = "#5cb85c", width = 2),
                  marker = list(color = "#5cb85c", size = 6),
                  hovertemplate = "Iter %{x}<br>Final: %{y:.3f}<extra></extra>")

      if (hasDamp) {
        d_damp <- d %>% filter(!is.na(pre_damp) & abs(pre_damp - final_rsc) > 1e-4)
        if (nrow(d_damp) > 0) {
          p <- p %>% add_trace(data = d_damp,
                               x = ~iteration, y = ~pre_damp,
                               name = "Pre-damp", mode = "markers", type = "scatter",
                               marker = list(color = "#e07b39", size = 10, symbol = "circle-open",
                                             line = list(width = 2, color = "#e07b39")),
                               opacity = 0.7,
                               hovertemplate = "Iter %{x}<br>Pre-damp: %{y:.3f}<extra></extra>")
        }
      }

      p %>% layout(
        hovermode = "x unified", showlegend = TRUE,
        legend = list(orientation = "h", y = -0.18),
        xaxis = list(title = "Iteration", zeroline = FALSE,
                     range = c(0.5, max(d$iteration) + 0.5),
                     gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                     showline = TRUE, linecolor = "#adadad", linewidth = 2),
        yaxis = list(title = "Rescale factor", gridcolor = "#e8e8e8",
                     ticks = "outside", showline = TRUE, linecolor = "#adadad", linewidth = 2)
      )
    })
  })

```


```{r phaseDiagramChart, echo=FALSE, include=FALSE}

g[["phase_diagram"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)

      emi_d <- p47_emiMktCurrent_iter %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        mutate(emissions = p47_emiMktCurrent_iter * 3.66667 * 1000) %>%
        select(iteration, emissions)

      price_d <- price %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        select(iteration, pm_taxemiMkt_iteration) %>%
        rename(carbon_price = pm_taxemiMkt_iteration)

      d_phase <- emi_d %>%
        inner_join(price_d, by = "iteration") %>%
        left_join(tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod,
                                       emiMktExt == emiMktExtVal) %>%
                    select(iteration, method, converged), by = "iteration") %>%
        arrange(iteration)

      tgt_emi_MtCO2 <- tryCatch({
        v <- pm_emiMktTarget %>%
          filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
          slice(1) %>% pull(pm_emiMktTarget)
        if (length(v) == 1 && !is.na(v)) v * 3.66667 * 1000 else NULL
      }, error = function(e) NULL)

      p <- plot_ly(data = d_phase, height = 350)

      if (!is.null(tgt_emi_MtCO2)) {
        p <- p %>% add_segments(
          x = min(d_phase$carbon_price, na.rm = TRUE),
          xend = max(d_phase$carbon_price, na.rm = TRUE),
          y = tgt_emi_MtCO2, yend = tgt_emi_MtCO2,
          showlegend = FALSE,
          line = list(color = "#e07b39", dash = "dash", width = 1.5)
        )
      }

      d_phase <- d_phase %>%
        mutate(
          primary_method = sub(" \\+.*", "", ifelse(is.na(method), "unknown", method)),
          primary_method = ifelse(primary_method %in% names(methodPalette), primary_method, "unknown"),
          point_color    = ifelse(primary_method %in% names(methodPalette),
                                  methodPalette[primary_method], "#aaaaaa")
        )

      p <- p %>%
        add_trace(x = ~carbon_price, y = ~emissions,
                  type = "scatter", mode = "lines",
                  line = list(color = "#cccccc", width = 1),
                  showlegend = FALSE, hoverinfo = "skip")

      for (meth in unique(d_phase$primary_method)) {
        d_sub <- d_phase %>% filter(primary_method == meth)
        col   <- if (meth %in% names(methodPalette)) methodPalette[[meth]] else "#aaaaaa"
        p <- p %>% add_trace(
          data = d_sub,
          x = ~carbon_price, y = ~emissions,
          type = "scatter", mode = "markers+text",
          name = meth,
          text = ~as.character(iteration), textposition = "top right",
          textfont = list(size = 9, color = "#555"),
          marker = list(size = 8, color = col),
          customdata = ~paste0("Iter: ", iteration,
                               "<br>Price: ", round(carbon_price), " US$2005/t CO2",
                               "<br>Emi: ", round(emissions), " Mt CO2",
                               "<br>Method: ", ifelse(is.na(method), "—", method)),
          hovertemplate = "%{customdata}<extra></extra>"
        )
      }

      seg_arrows <- if (nrow(d_phase) > 1)
        lapply(seq_len(nrow(d_phase) - 1), function(i)
          list(
            ax       = d_phase$carbon_price[i],
            ay       = d_phase$emissions[i],
            x        = d_phase$carbon_price[i + 1],
            y        = d_phase$emissions[i + 1],
            xref = "x", yref = "y", axref = "x", ayref = "y",
            arrowhead = 2, arrowsize = 0.8, arrowwidth = 1.2,
            arrowcolor = "#aaaaaa",
            showarrow = TRUE, text = ""
          )
        ) else list()

      p %>% layout(
          showlegend = TRUE,
          legend = list(orientation = "h", y = -0.22, font = list(size = 11)),
          annotations = seg_arrows,
          xaxis = list(title = "Carbon price (US$2005/t CO2)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "Emissions (Mt CO2)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2)
        )
    })
  })

```


```{r methodTimelineChart, echo=FALSE, include=FALSE}

g[["method_timeline"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        arrange(iteration) %>%
        mutate(
          primary = sub(" \\+.*", "", method),
          primary = ifelse(!is.na(converged) & converged == "✓", "converged",
                           ifelse(is.na(primary) | nchar(trimws(primary)) == 0, "inactive", primary)),
          color   = ifelse(primary == "inactive", "#eeeeee",
                           ifelse(primary %in% names(methodPalette), as.character(methodPalette[primary]), "#aaaaaa")),
          label = ifelse(is.na(method) | nchar(trimws(method)) == 0, "target not active", method)
        )

      present_methods <- intersect(names(methodPalette), unique(d$primary))

      # Single bar trace — continuous strip, per-bar color from d$color
      p <- plot_ly(height = 70) %>%
        add_bars(
          data = d,
          x = ~iteration, y = rep(1, nrow(d)),
          marker = list(color = d$color, line = list(width = 0)),
          hovertemplate = ~paste0("Iter ", iteration, ": ", label, "<extra></extra>"),
          showlegend = FALSE
        )

      # Circle legend entries — placed off-chart so they appear without fading
      for (meth in names(methodPalette)) {
        if (!meth %in% present_methods) next
        col <- methodPalette[[meth]]
        p <- p %>% add_trace(
          type = "scatter", mode = "markers",
          x = -999L, y = -999,
          name = meth,
          marker = list(color = col, symbol = "circle", size = 8),
          showlegend = TRUE,
          hoverinfo = "skip"
        )
      }

      p %>% layout(
        barmode  = "stack",
        showlegend = TRUE,
        legend = list(
          orientation = "h",
          x = 0.98, xanchor = "right",
          y = 0.45, yanchor = "top",
          font = list(size = 10),
          traceorder = "normal",
          bgcolor = "rgba(0,0,0,0)",
          borderwidth = 0
        ),
        xaxis = list(title = "", zeroline = FALSE,
                     range = c(0.5, max(d$iteration) + 0.5),
                     fixedrange = TRUE,
                     showticklabels = FALSE, showgrid = FALSE),
        yaxis = list(title = "", showticklabels = FALSE, zeroline = FALSE,
                     range = c(0, 1.01), fixedrange = TRUE, showgrid = FALSE),
        margin = list(t = 2, b = 50, l = 80, r = 60)
      )
    })
  })

```


```{r slopeClampChart, echo=FALSE, include=FALSE}

g[["slope_clamp"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>% arrange(iteration)

      p <- plot_ly(data = d, height = 280)

      if ("clamp_bnd" %in% names(d) && !all(is.na(d$clamp_bnd))) {
        p <- p %>% add_trace(x = ~iteration, y = ~clamp_bnd,
                             name = "Adaptive clamp bound", type = "scatter", mode = "lines",
                             line = list(color = "#e07b39", width = 1.5, dash = "dot", shape = "hv"),
                             hovertemplate = "Iter %{x}<br>Clamp bound: %{y:.3f}<extra></extra>")
      }
      if ("raw_slope" %in% names(d) && !all(is.na(d$raw_slope))) {
        p <- p %>% add_trace(x = ~iteration, y = ~raw_slope,
                             name = "Raw slope (pre-clamp)", type = "scatter", mode = "markers",
                             marker = list(color = "#e07b39", size = 9, symbol = "circle-open",
                                           line = list(width = 2, color = "#e07b39")),
                             opacity = 0.7,
                             hovertemplate = "Iter %{x}<br>Raw slope: %{y:.3f}<extra></extra>")
      }

      p %>%
        add_trace(x = ~iteration, y = ~used_slope,
                  name = "Used slope", type = "scatter", mode = "lines+markers",
                  line = list(color = "#4a9eca", width = 2),
                  marker = list(color = "#4a9eca", size = 6),
                  hovertemplate = "Iter %{x}<br>Used slope: %{y:.3f}<extra></extra>") %>%
        add_segments(x = min(d$iteration) - 0.5, xend = max(d$iteration) + 0.5,
                     y = 0, yend = 0, showlegend = FALSE,
                     line = list(color = "#e8e8e8", width = 2)) %>%
        layout(
          hovermode = "x unified", showlegend = TRUE,
          legend = list(orientation = "h", y = -0.2),
          xaxis = list(title = "Iteration", zeroline = FALSE,
                       range = c(0.5, max(d$iteration) + 0.5),
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "Slope", gridcolor = "#e8e8e8",
                       mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2)
        )
    })
  })

```


```{r convergenceRateChart, echo=FALSE, include=FALSE}

g[["convergence_rate"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        arrange(iteration) %>% filter(!is.na(delta_dev)) %>%
        mutate(abs_delta = delta_dev)

      plot_ly(data = d, height = 220) %>%
        add_bars(x = ~iteration, y = ~abs_delta,
                 marker = list(color = "#4a9eca", opacity = 0.85),
                 hovertemplate = "Iter %{x}<br>|Δdev|: %{y:.2f}%<extra></extra>",
                 showlegend = FALSE) %>%
        layout(
          xaxis = list(title = "Iteration", zeroline = FALSE,
                       range = c(0.5, max(d$iteration) + 0.5),
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "|Δdeviation| (%)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2)
        )
    })
  })

```


```{r summaryData, echo=FALSE, include=FALSE}

activity_long <- tableData %>%
  mutate(
    target_label = paste(ext_regi, ttot2, emiMktExt, sep = " · "),
    primary      = sub(" \\+.*", "", method),
    primary      = ifelse(!is.na(converged) & converged == "✓", "converged",
                          ifelse(is.na(primary) | nchar(trimws(primary)) == 0, "inactive", primary)),
    cell_color   = ifelse(primary == "inactive", "#eeeeee",
                          ifelse(primary %in% names(methodPalette), as.character(methodPalette[primary]), "#aaaaaa")),
    is_conv = !is.na(converged) & converged == "✓",
    tooltip = ifelse(
      primary == "inactive",
      paste0(
        gsub(" · ", " | ", target_label), " — Iter ", iteration,
        "<br>Method: not active"
      ),
      paste0(
        gsub(" · ", " | ", target_label), " — Iter ", iteration,
        "<br>Method: ", ifelse(is.na(method) | nchar(trimws(method)) == 0, "not active", method),
        "<br>Deviation: ", ifelse(is.na(dev_pct), "—", paste0(dev_pct, "%")),
        "<br>|Δdev|: ", ifelse(is.na(delta_dev), "—", paste0(delta_dev, "%")),
        "<br>Applied factor: ", ifelse(is.na(final_rsc), "—", as.character(final_rsc)),
        "<br>Converged: ", ifelse(is_conv, "✓", "✗")
      )
    )
  )

conv_first <- activity_long %>%
  filter(is_conv) %>%
  group_by(target_label) %>%
  summarise(first_conv = min(iteration), .groups = "drop")

target_summary_df <- tableData %>%
  group_by(ext_regi, ttot2, emiMktExt) %>%
  arrange(iteration) %>%
  summarise(
    n_total_iters   = max(iteration),
    n_active        = sum(!is.na(method)),
    n_conv_iters    = sum(!is.na(converged) & converged == "✓"),
    first_conv_iter = { ci <- iteration[!is.na(converged) & converged == "✓"]
                        if (length(ci) > 0) min(ci) else NA_integer_ },
    last_conv_iter  = { ci <- iteration[!is.na(converged) & converged == "✓"]
                        if (length(ci) > 0) max(ci) else NA_integer_ },
    pct_conv        = round(100 * sum(!is.na(converged) & converged == "✓") /
                            max(1, max(iteration)), 1),
    final_dev_pct   = last(dev_pct),
    final_conv      = last(converged),
    peak_abs_dev    = round(max(abs(dev_pct), na.rm = TRUE), 2),
    n_damped        = sum(!is.na(damped) & damped == "✓"),
    methods_used    = {
      pm <- sub(" \\+.*", "", method[!is.na(method)])
      pm <- pm[nchar(trimws(pm)) > 0]
      if (length(pm) == 0) "—"
      else { mc <- sort(table(pm), decreasing = TRUE)
             paste(paste0(names(mc), " (", as.integer(mc), ")"), collapse = " · ") }
    },
    .groups = "drop"
  ) %>%
  mutate(
    tol_pct         = round(100 * emiMktTarget_tolerance[ext_regi], 2),
    target_label    = paste(ext_regi, ttot2, emiMktExt, sep = " · "),
    in_tol          = !is.na(final_dev_pct) & abs(final_dev_pct) <= tol_pct,
    first_conv_disp = ifelse(is.na(first_conv_iter), "—", as.character(first_conv_iter)),
    last_conv_disp  = ifelse(is.na(last_conv_iter),  "—", as.character(last_conv_iter))
  )

```


```{r summaryCharts, echo=FALSE, include=FALSE}

# ── Activity heatmap + legend ─────────────────────────────────────────────
target_order <- target_summary_df %>%
  arrange(ext_regi, desc(ttot2), emiMktExt) %>%
  pull(target_label) %>% rev()

activity_long$target_f <- factor(activity_long$target_label, levels = target_order)

p_act <- ggplot(activity_long,
                aes(x = iteration, y = target_f, fill = cell_color, text = tooltip)) +
  geom_tile(height = 0.82, color = NA, linewidth = 0) +
  scale_fill_identity() +
  theme_minimal() +
  theme(axis.title.y = element_blank(), panel.grid = element_blank(),
        axis.text.y = element_text(size = 9), axis.title.x = element_text(size = 9),
        plot.margin = margin(4, 4, 4, 4, "pt")) +
  labs(x = "Iteration")

act_h <- max(100, length(target_order) * 38 + 90)
act_plotly <- ggplotly(p_act, tooltip = "text", height = act_h) %>%
  layout(
    showlegend = TRUE,
    legend = list(orientation = "h", x = 0.98, xanchor = "right",
                  y = -0.04, yanchor = "top",
                  font = list(size = 10), bgcolor = "rgba(0,0,0,0)", borderwidth = 0),
    shapes = lapply(conv_first$first_conv, function(i)
      list(type = "line", x0 = i - 0.5, x1 = i - 0.5, y0 = 0, y1 = 1,
           yref = "paper", line = list(color = "rgba(0,0,0,0.18)", width = 1, dash = "dot")))
  )

present_in_summary <- intersect(names(methodPalette), unique(activity_long$primary))
for (meth in names(methodPalette)) {
  if (!meth %in% present_in_summary) next
  act_plotly <- act_plotly %>% add_trace(
    type = "scatter", mode = "markers",
    x = -999L, y = target_order[1],
    name = meth,
    marker = list(color = methodPalette[[meth]], symbol = "circle", size = 8),
    showlegend = TRUE, hoverinfo = "skip",
    inherit = FALSE
  )
}
g[["summary_activity"]] <- act_plotly %>%
  layout(xaxis = list(range = c(0.5, max(activity_long$iteration) + 0.5),
                      fixedrange = TRUE))

# ── Final deviation strip ──────────────────────────────────────────────────
dev_h         <- max(140, nrow(target_summary_df) * 42 + 80)
tgt_order_dev <- target_summary_df %>% arrange(ext_regi, ttot2, emiMktExt) %>% pull(target_label)

dev_data <- target_summary_df %>%
  arrange(ext_regi, ttot2, emiMktExt) %>%
  mutate(dot_color = ifelse(in_tol, "#00BFC4", "#F8766D"))

g[["summary_conv_dev"]] <- plot_ly(dev_data, height = dev_h, width = 550) %>%
  add_segments(x = ~-tol_pct, xend = ~tol_pct, y = ~target_label, yend = ~target_label,
               line = list(color = "rgba(0,191,196,0.12)", width = 14),
               showlegend = FALSE, hoverinfo = "skip") %>%
  add_segments(x = 0, xend = ~final_dev_pct, y = ~target_label, yend = ~target_label,
               line = list(color = "#cccccc", width = 1),
               showlegend = FALSE, hoverinfo = "skip") %>%
  add_markers(x = ~final_dev_pct, y = ~target_label,
              marker = list(color = ~dot_color, size = 10),
              text = ~paste0(target_label, "<br>Final dev: ", final_dev_pct, "%",
                             "<br>Tolerance: ±", tol_pct, "%<br>", final_conv),
              hoverinfo = "text", showlegend = FALSE) %>%
  layout(
    shapes = list(list(type = "line", x0 = 0, x1 = 0, y0 = 0, y1 = 1,
                       yref = "paper", line = list(color = "#999", width = 1))),
    xaxis = list(title = "Final deviation (%)", zeroline = FALSE, gridcolor = "#e8e8e8",
                 mirror = TRUE, ticks = "outside", showline = TRUE,
                 linecolor = "#adadad", linewidth = 2),
    yaxis = list(title = "", zeroline = FALSE, gridcolor = "#e8e8e8",
                 categoryorder = "array", categoryarray = rev(tgt_order_dev))
  )

# ── Summary reactable ──────────────────────────────────────────────────────
convCell <- function(v) if (identical(v, "✓"))
  htmltools::span(style = "color:#00BFC4;font-weight:bold;", v) else
  htmltools::span(style = "color:#F8766D;font-weight:bold;", v)

devCell <- function(v, row) {
  tol <- target_summary_df$tol_pct[row]
  col <- if (!is.na(v) && abs(v) <= tol) "#00BFC4" else "#F8766D"
  htmltools::span(style = paste0("color:", col, ";"), paste0(v, "%"))
}

summary_react <- reactable::reactable(
  target_summary_df %>%
    select(target_label, final_conv, first_conv_disp, final_dev_pct,
           tol_pct, n_damped, n_conv_iters, methods_used),
  columns = list(
    target_label    = reactable::colDef(name = "Target",         minWidth = 140),
    final_conv      = reactable::colDef(name = "Conv.",          align = "center", width = 56,
                                        html = TRUE, cell = convCell),
    first_conv_disp = reactable::colDef(name = "First conv.",    align = "center", width = 80),
    final_dev_pct   = reactable::colDef(name = "Final dev (%)",  align = "right",  width = 95,
                                        cell = devCell),
    tol_pct         = reactable::colDef(name = "Tolerance",      align = "right",  width = 80,
                                        cell = function(v) paste0("±", v, "%")),
    n_damped        = reactable::colDef(name = "Damped",         align = "center", width = 70),
    n_conv_iters    = reactable::colDef(name = "Iters conv.",    align = "center", width = 80),
    methods_used    = reactable::colDef(name = "Methods used",
                                        style = list(fontSize = "0.82em", color = "#555"))
  ),
  striped = TRUE, highlight = TRUE, compact = TRUE, pagination = FALSE,
  theme = reactable::reactableTheme(
    headerStyle = list(borderBottom = "2px solid #ccc", fontSize = "0.88em",
                       background = "#eeeeee")
  )
)

```


## {.tabset}

```{r, include=FALSE}

# Creating sections
out <- "### Summary\n\n"

out <- paste0(out, "##### Activity and Convergence Timeline\n\n")
out <- paste0(out, "<div class='collapsible-section'>\n")
out <- paste0(out, "<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
out <- paste0(out, "<div class='collapsible-content'>\n\n")
out <- paste0(out, "```{r, results='asis'}\ng$summary_activity\n```\n\n")
out <- paste0(out, "</div></div>\n\n")

out <- paste0(out, "##### Final Deviation vs Tolerance\n\n")
out <- paste0(out, "<div class='collapsible-section'>\n")
out <- paste0(out, "<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
out <- paste0(out, "<div class='collapsible-content'>\n\n")
out <- paste0(out, "```{r, results='asis'}\ng$summary_conv_dev\n```\n\n")
out <- paste0(out, "</div></div>\n\n")

out <- paste0(out, "##### Summary Table\n\n")
out <- paste0(out, "<div class='collapsible-section'>\n")
out <- paste0(out, "<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
out <- paste0(out, "<div class='collapsible-content'>\n\n")
out <- paste0(out, "```{r, results='asis'}\nsummary_react\n```\n\n")
out <- paste0(out, "</div></div>\n\n")

for(reg in tgtRegions){
  out <- paste0(out,knitr::knit_expand(text=paste0("### ", reg, " {.tabset}\n\n")))
  for(subGroup in tgts %>% filter(ext_regi == reg) %>% pull(group)){

    tgtPeriod    <- strsplit(subGroup,"_",fixed=T)[[1]][1]
    unit         <- pm_emiMktTarget %>% filter(ext_regi == reg, ttot2 == tgtPeriod) %>% pull(emi_type_47)
    emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == subGroup) %>% pull(emiMktExt)

    out <- paste0(out,knitr::knit_expand(text=paste0("#### ", subGroup, "\n\n")))

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Emissions and Carbon Price (", unit, ")\n\n")))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"<div style='margin-bottom:-18px'>\n")
    out <- paste0(out,"```{r , results = 'asis', fig.height=10}\n\n")
    out <- paste0(out,"g$qtty_and_price[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div>\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$method_timeline[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis', fig.height=10}\n\n")
    out <- paste0(out,"g$deviation[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , echo=FALSE}\n\n")
    out <- paste0(out,"makeRescaleTable(tableData, '",reg,"', '",tgtPeriod,"', '",emiMktExtVal,"')\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Rescale Factor\n\n")))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$rescale_factor[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Phase Diagram\n\n")))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$phase_diagram[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Slope and Clamp Evolution\n\n")))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$slope_clamp[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Convergence Rate\n\n")))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$convergence_rate[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")
  }
}


```

`r paste(knitr::knit(text = paste(out, collapse = '\n')))`


