---
title: "CO2 budget target convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gdx)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r setup, echo=FALSE, include=FALSE}

emptyPlot <- function(var, df = NULL) {
  cat(paste0("str() of raw data for '", var, "':\n"))
  str(df)
  p <- ggplot() +
    annotate("text", x = 0, y = 0, label = paste0("no data for '", var, "' in this gdx")) +
    theme_void()
  ggplotly(p)
}

tol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "cm_budgetCO2_absDevTol", react = "silent")))
if (length(tol) == 0 || is.na(tol)) tol <- NA_real_

# whether damages are internalized -> global budget check is one-sided (positive only)
m2r <- suppressWarnings(gdx::readGDX(params$gdx, "module2realisation", react = "silent"))
damagesOn <- !is.null(m2r) && m2r[m2r$modules == "internalizeDamages", 2] != "off"

```

The global CO2 budget criterion requires the cumulated CO2 emissions to stay within
**+-`r ifelse(is.na(tol), "cm_budgetCO2_absDevTol", tol)` Gt CO2** of the target budget
(`cm_budgetCO2_absDevTol`). Positive deviation = target budget **exceeded**.

`r if (damagesOn) "**Damages are internalized in this run**, so only *positive* deviation (budget overshoot) is checked - a budget undershoot is acceptable." else "Both positive and negative deviations are checked (symmetric band)."`

## Convergence summary — global budget deviation

Absolute deviation of cumulated global CO2 emissions from the target budget, over
iterations. The shaded band is the tolerance; green points are inside it (converged),
red are outside. This is the criterion that drives convergence.

```{r global, results = "asis"}

df <- mip::getPlotData("p80_globalBudget_absDev_iter", params$gdx)

if (nrow(df) == 0) {
  emptyPlot("p80_globalBudget_absDev_iter", df)
} else {
  df <- df %>%
    rename("value" = "p80_globalBudget_absDev_iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "converged" = if (is.na(tol)) "yes" else ifelse(.data$value > tol | (!damagesOn & .data$value < -tol), "no", "yes"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>Deviation: ", round(.data$value, 3), " Gt CO2",
                         ifelse(.data$value > 0, " (budget exceeded)", " (below budget)"))
    )

  p <- ggplot(df, aes_(x = ~iteration, y = ~value, text = ~tooltip, group = 1)) +
    { if (!is.na(tol)) {
        if (damagesOn) {
          annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = tol,
                   fill = "#00BFC4", alpha = 0.15)
        } else {
          annotate("rect", xmin = -Inf, xmax = Inf, ymin = -tol, ymax = tol,
                   fill = "#00BFC4", alpha = 0.15)
        }
      } else NULL } +
    { if (!is.na(tol)) geom_hline(yintercept = if (damagesOn) tol else c(-tol, tol),
                                  linetype = "dashed", color = "grey40") else NULL } +
    geom_hline(yintercept = 0, color = "grey80") +
    geom_line(alpha = 0.7) +
    suppressWarnings(geom_point(aes_(fill = ~converged), size = 2, shape = 21, color = "white")) +
    scale_fill_manual(values = c("yes" = "#00BFC4", "no" = "#F8766D")) +
    theme_bw() +
    labs(x = "iteration", y = "deviation from target budget (Gt CO2)")

  ggplotly(p, tooltip = c("text"))
}

```

## Regional budget deviation

Where a regional budget target is set, the absolute deviation per region over
iterations. Dashed lines mark the +-`r ifelse(is.na(tol), "tolerance", tol)` Gt CO2 band.

```{r regional, results = "asis"}

dfr <- mip::getPlotData("p80_regionalBudget_absDev_iter", params$gdx)

if (nrow(dfr) == 0) {
  cat("No regional budget target active in this run.")
} else {
  dfr <- dfr %>%
    rename("value" = "p80_regionalBudget_absDev_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))

  # drop regions that are all-zero across iterations (no regional target there)
  active <- dfr %>%
    group_by(.data$all_regi) %>%
    summarise(nz = any(abs(.data$value) > 1e-9)) %>%
    filter(.data$nz) %>%
    pull(.data$all_regi)

  dfr <- dfr %>% filter(.data$all_regi %in% active)

  if (nrow(dfr) == 0) {
    cat("No region has a non-zero regional budget deviation in this run.")
  } else {
    p <- ggplot(dfr, aes_(x = ~iteration, y = ~value, group = ~all_regi)) +
      { if (!is.na(tol)) geom_hline(yintercept = c(-tol, tol),
                                    linetype = "dashed", color = "grey50") else NULL } +
      geom_hline(yintercept = 0, color = "grey80") +
      geom_line(alpha = 0.7) +
      geom_point(size = 0.8, alpha = 0.7) +
      facet_wrap(~all_regi, scales = "free_y") +
      theme_bw() +
      labs(x = "iteration", y = "deviation from regional budget (Gt CO2)")

    ggplotly(p)
  }
}

```
