---
title: "Damage internalization convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gdx)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r setup, echo=FALSE, include=FALSE}

emptyPlot <- function(var, df = NULL) {
  cat(paste0("str() of raw data for '", var, "':\n"))
  str(df)
  p <- ggplot() +
    annotate("text", x = 0, y = 0, label = paste0("no data for '", var, "' in this gdx")) +
    theme_void()
  ggplotly(p)
}

booleanColor <- c("yes" = "#00BFC4", "no" = "#F8766D")

sccTol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "cm_sccConvergence", react = "silent")))
tempTol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "cm_tempConvergence", react = "silent")))
if (length(sccTol) == 0) sccTol <- NA_real_
if (length(tempTol) == 0) tempTol <- NA_real_

# helper: single iteration-indexed series with a threshold line
plotConvSeries <- function(var, tol, ylab) {
  df <- mip::getPlotData(var, params$gdx)
  if (nrow(df) == 0) return(emptyPlot(var, df))

  df <- df %>%
    rename("value" = !!var) %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "converged" = if (is.na(tol)) "yes" else ifelse(.data$value > tol, "no", "yes"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>", ylab, ": ", round(.data$value, 4),
                         if (!is.na(tol)) paste0("<br>threshold: ", tol) else "")
    )

  p <- ggplot(df, aes_(x = ~iteration, y = ~value, group = 1)) +
    { if (!is.na(tol)) geom_hline(yintercept = tol, linetype = "dashed", color = "grey40") else NULL } +
    geom_line(alpha = 0.6) +
    suppressWarnings(geom_point(aes_(fill = ~converged, text = ~tooltip), size = 2, shape = 21, color = "white")) +
    scale_fill_manual(values = booleanColor) +
    theme_bw() +
    labs(x = "iteration", y = ylab)

  ggplotly(p, tooltip = c("text"))
}

```

Damage internalization converges only when **both** sub-criteria are met in the same
iteration:

- the **social cost of carbon (SCC)** is stable between iterations
  (`p80_sccConvergenceMaxDeviation_iter <= cm_sccConvergence`
  = `r ifelse(is.na(sccTol), "NA", sccTol)`), and
- the **global mean temperature** is stable
  (`p80_gmt_conv_iter <= cm_tempConvergence` = `r ifelse(is.na(tempTol), "NA", tempTol)`).

The two charts below show each gate separately - the one still above its dashed
threshold is the one holding up convergence.

## Convergence summary

Per iteration, whether damage internalization converged (i.e. **both** the SCC and the
temperature gate are within tolerance). Green = converged, red = not.

```{r summary, results = "asis"}

scc <- mip::getPlotData("p80_sccConvergenceMaxDeviation_iter", params$gdx)
gmt <- mip::getPlotData("p80_gmt_conv_iter", params$gdx)

if (nrow(scc) == 0 || nrow(gmt) == 0) {
  cat("No damage-internalization convergence data in this gdx (module internalizeDamages is off in this run).")
} else {
  scc <- scc %>% rename("scc" = "p80_sccConvergenceMaxDeviation_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))
  gmt <- gmt %>% rename("gmt" = "p80_gmt_conv_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))

  comb <- full_join(scc, gmt, by = "iteration") %>%
    arrange(.data$iteration) %>%
    mutate(
      "converged" = ifelse((is.na(sccTol) | .data$scc <= sccTol) &
                             (is.na(tempTol) | .data$gmt <= tempTol), "yes", "no"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>SCC max dev: ", round(.data$scc, 4),
                         "<br>GMT conv: ", round(.data$gmt, 4),
                         "<br>", ifelse(.data$converged == "yes", "converged", "not converged"))
    )

  p <- suppressWarnings(ggplot(comb, aes_(x = ~iteration, y = "Damage\nInternalization",
                                          fill = ~converged, text = ~tooltip)) +
    geom_point(size = 3, shape = 21, color = "white") +
    scale_fill_manual(values = booleanColor) +
    theme_minimal() +
    labs(x = "iteration", y = NULL))

  ggplotly(p, tooltip = c("text"))
}

```

## SCC convergence

Maximum deviation of the social cost of carbon from the previous iteration [%]. Red
points are above the threshold (not converged).

```{r scc, results = "asis"}
plotConvSeries("p80_sccConvergenceMaxDeviation_iter", sccTol, "SCC max deviation")
```

## Temperature convergence

Global mean temperature convergence metric per iteration. Red points are above the
threshold (not converged).

```{r temp, results = "asis"}
plotConvSeries("p80_gmt_conv_iter", tempTol, "GMT convergence")
```

## Combined verdict

Per iteration, whether each gate passes and whether damage internalization as a whole
is converged.

```{r combined, results = "asis"}

scc <- mip::getPlotData("p80_sccConvergenceMaxDeviation_iter", params$gdx)
gmt <- mip::getPlotData("p80_gmt_conv_iter", params$gdx)

if (nrow(scc) == 0 || nrow(gmt) == 0) {
  cat("no data")
} else {
  scc <- scc %>% rename("scc" = "p80_sccConvergenceMaxDeviation_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))
  gmt <- gmt %>% rename("gmt" = "p80_gmt_conv_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))

  comb <- full_join(scc, gmt, by = "iteration") %>%
    arrange(.data$iteration) %>%
    mutate(
      "SCC gate" = ifelse(is.na(sccTol) | .data$scc <= sccTol, "pass", "fail"),
      "Temp gate" = ifelse(is.na(tempTol) | .data$gmt <= tempTol, "pass", "fail"),
      "converged" = ifelse(.data$`SCC gate` == "pass" & .data$`Temp gate` == "pass", "yes", "no")
    )

  knitr::kable(
    comb %>% transmute(
      iteration = .data$iteration,
      `SCC max dev` = round(.data$scc, 4),
      `GMT conv` = round(.data$gmt, 4),
      `SCC gate` = .data$`SCC gate`,
      `Temp gate` = .data$`Temp gate`,
      converged = .data$converged
    ),
    row.names = FALSE
  )
}

```
