---
title: "Tax convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gdx)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r setup, echo=FALSE, include=FALSE}

# placeholder for absent/empty variables (ggplotly errors on zero-row data)
emptyPlot <- function(var, df = NULL) {
  cat(paste0("str() of raw data for '", var, "':\n"))
  str(df)
  p <- ggplot() +
    annotate("text", x = 0, y = 0, label = paste0("no data for '", var, "' in this gdx")) +
    theme_void()
  ggplotly(p)
}

booleanColor <- c("yes" = "#00BFC4", "no" = "#F8766D")

# getPlotData() errors when a symbol is entirely absent from the gdx (e.g. the
# per-component parameter on runs that predate it). Return a 0-row frame instead so the
# nrow() == 0 guards handle it gracefully.
getPlotDataSafe <- function(var, gdx) {
  tryCatch(mip::getPlotData(var, gdx), error = function(e) data.frame())
}

# Tolerance: read cm_TaxConv_tolerance from the gdx (exported by nash); fall back to the
# historical hardcoded value (0.1% of GDP) for older gdx files.
tol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "cm_TaxConv_tolerance", react = "silent")))
if (length(tol) == 0 || is.na(tol)) tol <- 0.001

df <- mip::getPlotData("p80_convNashTaxrev_iter", params$gdx)
if (nrow(df) > 0) {
  df <- df %>%
    rename("value" = "p80_convNashTaxrev_iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "ttot" := as.numeric(as.character(.data$ttot))
    )
}

```

The tax convergence criterion requires the **net tax revenue in every region and
period to be below `r round(100 * tol, 2)`% of GDP** (`|p80_convNashTaxrev_iter| <=
`r tol``). A non-zero tax revenue means the region is still collecting/paying a net
implicit tax that should net out at the Nash equilibrium, so any region above the band
blocks convergence.

## Convergence summary

Per iteration, the largest absolute net tax revenue across all regions and periods,
versus the tolerance (dashed). The criterion converges once the worst region/period is
inside the band. Green = converged, red = not.

```{r summary, results = "asis"}

if (nrow(df) == 0) {
  emptyPlot("p80_convNashTaxrev_iter", df)
} else {
  s <- df %>%
    group_by(.data$iteration) %>%
    summarise("worst" = max(abs(.data$value)), .groups = "drop") %>%
    mutate(
      "converged" = ifelse(.data$worst > tol, "no", "yes"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>Worst |tax revenue|: ", round(100 * .data$worst, 3), "% of GDP<br>",
                         ifelse(.data$worst > tol,
                                paste0("not converged (> ", round(100 * tol, 2), "%)"),
                                "converged"))
    )

  p <- ggplot(s, aes_(x = ~iteration, y = ~worst, group = 1, text = ~tooltip)) +
    geom_hline(yintercept = tol, linetype = "dashed", color = "grey40") +
    geom_line(alpha = 0.6) +
    suppressWarnings(geom_point(aes_(fill = ~converged), size = 2, shape = 21, color = "white")) +
    scale_fill_manual(values = booleanColor) +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(x = "iteration", y = "worst |tax revenue| (share of GDP)")

  ggplotly(p, tooltip = c("text"))
}

```

## Deviation per region and iteration

Signed tax revenue as a share of GDP. Blue = below target (paying), red = above
(collecting); saturated cells are outside the `r round(100 * tol, 2)`% band. One panel
per period. Hover for the exact value.

```{r heatmap, results = "asis"}

if (nrow(df) == 0) {
  emptyPlot("p80_convNashTaxrev_iter", df)
} else {
  cap <- 5 * tol
  hd <- df %>%
    mutate(
      "fillv" = pmax(pmin(.data$value, cap), -cap),
      "tooltip" = paste0(
        .data$all_regi, " ", .data$ttot,
        "<br>Iteration: ", .data$iteration,
        "<br>Tax revenue: ", round(100 * .data$value, 3), "% of GDP",
        "<br>", ifelse(abs(.data$value) > tol, "outside", "within"), " +-",
        round(100 * tol, 2), "% band"
      )
    )

  p <- ggplot(hd, aes_(x = ~iteration, y = ~all_regi, fill = ~fillv, text = ~tooltip)) +
    geom_tile(color = "white", linewidth = 0.1) +
    facet_wrap(~ttot) +
    scale_fill_gradient2(
      low = "#2c7bb6", mid = "#ffffbf", high = "#d7191c", midpoint = 0,
      name = "share of GDP", labels = scales::percent
    ) +
    theme_minimal() +
    labs(x = "iteration", y = NULL) +
    theme(panel.grid = element_blank())

  ggplotly(p, tooltip = c("text"))
}

```

## Convergence trajectory per region

Tax revenue over iterations, one panel per region (free y-axis). The dashed lines mark
the +-`r round(100 * tol, 2)`% tolerance band; colour distinguishes the target period.

```{r lines, results = "asis"}

if (nrow(df) == 0) {
  emptyPlot("p80_convNashTaxrev_iter", df)
} else {
  p <- ggplot(df, aes_(x = ~iteration, y = ~value, color = ~factor(ttot),
                       group = ~interaction(all_regi, ttot))) +
    geom_hline(yintercept = c(-tol, tol), linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, color = "grey80") +
    geom_line(alpha = 0.7) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~all_regi, scales = "free_y") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(x = "iteration", y = "tax revenue (share of GDP)", color = "period")

  ggplotly(p)
}

```

## Worst offender per iteration

For each iteration, the single region/period furthest outside the tolerance band -
the cell that most needs attention.

```{r worst, results = "asis"}

if (nrow(df) == 0) {
  cat("no data")
} else {
  worst <- df %>%
    group_by(.data$iteration) %>%
    slice(which.max(abs(.data$value))) %>%
    ungroup() %>%
    mutate(
      "converged" = ifelse(abs(.data$value) > tol, "no", "yes"),
      "abs_share" = abs(.data$value)
    ) %>%
    arrange(.data$iteration)

  knitr::kable(
    worst %>%
      transmute(
        iteration = .data$iteration,
        region = .data$all_regi,
        period = .data$ttot,
        `tax revenue (% of GDP)` = round(100 * .data$value, 3),
        converged = .data$converged
      ),
    row.names = FALSE
  )
}

```

## Tax revenue components (final iteration)

The convergence check is on the **total** net tax revenue, but that total sums many
tax/subsidy components. Only the final-iteration value of each component is stored in the
gdx (the `v21_taxrev*` variables), so this shows the **end-of-run composition** — which
taxes reach the largest magnitude and are therefore most able to push the net revenue out
of balance. Values are a share of GDP (`vm_cesIO(...,"inco")`), matching the convergence
metric. Components absent for this run's realisation are skipped (no error).

```{r components_read, echo=FALSE, include=FALSE}

# GDP denominator (final-iteration level), to express components as a share of GDP
gdpMag <- suppressWarnings(gdx::readGDX(params$gdx, "vm_cesIO", field = "l",
                                        restore_zeros = FALSE, react = "silent"))

# Tax revenue components read directly from the gdx (final-iteration levels, per the
# vm_taxrev composition in modules/21_tax). Each is summed over any sub-index (dimSums
# over the data dimension); sign = -1 for a subsidy that enters revenue with a minus.
# Robust to missing symbols (skipped).
compDefs <- list(
  list(sym = "v21_taxrevGHG",                  label = "GHG",                sign = 1),
  list(sym = "v21_taxrevCO2Sector",            label = "CO2 sector",         sign = 1),
  list(sym = "v21_taxrevCO2luc",               label = "CO2 LUC",            sign = 1),
  list(sym = "v21_taxemiMkt",                  label = "Emission market",    sign = 1),
  list(sym = "v21_taxrevPE",                   label = "PE",                 sign = 1),
  list(sym = "v21_taxrevResEx",                label = "Resource extr.",     sign = 1),
  list(sym = "v21_taxrevBioEF",                label = "Bioenergy fee",      sign = 1),
  list(sym = "v21_taxrevSE",                   label = "SE",                 sign = 1),
  list(sym = "v21_taxrevFE",                   label = "FE",                 sign = 1),
  list(sym = "v21_taxrevImport",               label = "PE import",          sign = 1),
  list(sym = "v21_taxrevPE2SE",                label = "PE2SE",              sign = 1),
  list(sym = "v21_taxrevSO2",                  label = "SO2",                sign = 1),
  list(sym = "v21_taxrevFlex",                 label = "Flexibility",        sign = 1),
  list(sym = "v21_taxrevCCS",                  label = "CCS",                sign = 1),
  list(sym = "v21_taxrevNetNegEmi",            label = "Net-neg emi",        sign = 1),
  list(sym = "v21_taxrevBioSust",              label = "Bio sustainability", sign = 1),
  list(sym = "v21_taxrevEI",                   label = "Energy intensity",   sign = 1),
  list(sym = "v21_taxrevChProdStartYear",      label = "Prod. change",       sign = 1),
  list(sym = "vm_costSubsidizeLearning",       label = "Learning subsidy",   sign = -1),
  list(sym = "v21_taxrevCES",                  label = "CES",                sign = 1),
  list(sym = "vm_taxrevimplicitQttyTargetTax", label = "Implicit qty target", sign = 1),
  list(sym = "vm_taxrevimplicitPriceTax",      label = "Implicit FE price",  sign = 1),
  list(sym = "vm_taxrevimplicitPePriceTax",    label = "Implicit PE price",  sign = 1)
)

readComp <- function(def) {
  x <- suppressWarnings(gdx::readGDX(params$gdx, def$sym, field = "l",
                                     restore_zeros = FALSE, react = "silent"))
  if (is.null(x)) return(NULL)
  df <- tryCatch(quitte::as.quitte(x), error = function(e) NULL)
  if (is.null(df) || nrow(df) == 0) return(NULL)
  # sum over any sub-index (e.g. entyPe, emiMkt, emi_sectors) down to region x period
  df <- df %>%
    group_by(.data$region, .data$period) %>%
    summarise("value" = def$sign * sum(.data$value, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(.data$value))
  if (nrow(df) == 0) return(NULL)
  df$component <- def$label
  df
}

comp <- dplyr::bind_rows(lapply(compDefs, readComp))

if (!is.null(gdpMag) && !is.null(comp) && nrow(comp) > 0) {
  gdp <- quitte::as.quitte(gdpMag[, , "inco"]) %>%
    group_by(.data$region, .data$period) %>%
    summarise("gdp" = sum(.data$value, na.rm = TRUE), .groups = "drop") %>%
    filter(.data$gdp != 0)

  comp <- comp %>%
    inner_join(gdp, by = c("region", "period")) %>%
    mutate(
      "share" = .data$value / .data$gdp,
      "period" := as.numeric(as.character(.data$period))
    ) %>%
    filter(abs(.data$share) > 1e-9)
}

```

Heatmap of each component's **peak signed contribution** per region (the value at the
period where that component is largest in magnitude). Deep red / blue cells are the
components most in play; components are ordered left-to-right by overall magnitude.

```{r components_heatmap, results = "asis"}

if (is.null(comp) || !("share" %in% names(comp)) || nrow(comp) == 0) {
  cat("No tax revenue component data in this gdx (module 21_tax is not `on`).")
} else {
  peak <- comp %>%
    group_by(.data$region, .data$component) %>%
    slice(which.max(abs(.data$share))) %>%
    ungroup() %>%
    mutate("tooltip" = paste0(.data$component, " · ", .data$region,
                              "<br>period ", .data$period,
                              "<br>", round(100 * .data$share, 3), "% of GDP"))

  compOrder <- peak %>% group_by(.data$component) %>%
    summarise("m" = max(abs(.data$share)), .groups = "drop") %>%
    arrange(.data$m) %>% pull(.data$component)
  peak$component <- factor(peak$component, levels = compOrder)

  hm <- ggplot(peak, aes_(x = ~component, y = ~region, fill = ~share, text = ~tooltip)) +
    geom_tile(color = "white", linewidth = 0.1) +
    scale_fill_gradient2(low = "#2c7bb6", mid = "#ffffbf", high = "#d7191c",
                         midpoint = 0, labels = scales::percent, name = "share of GDP") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid = element_blank()) +
    labs(x = NULL, y = NULL)

  ggplotly(hm, tooltip = c("text"))
}

```

### Largest contributors

Each component's peak magnitude (share of GDP) and where it peaks — the rows at the top
are the tax components most able to unbalance the net revenue and hold up convergence.

```{r components_table, results = "asis"}

if (is.null(comp) || !("share" %in% names(comp)) || nrow(comp) == 0) {
  cat("No component data.")
} else {
  tbl <- comp %>%
    group_by(.data$component) %>%
    slice(which.max(abs(.data$share))) %>%
    ungroup() %>%
    arrange(desc(abs(.data$share))) %>%
    transmute(
      component = .data$component,
      `peak |share| of GDP (%)` = round(100 * abs(.data$share), 3),
      `signed (%)` = round(100 * .data$share, 3),
      region = .data$region,
      period = .data$period
    )
  knitr::kable(tbl, row.names = FALSE)
}

```
