---
title: "Price anticipation convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gridExtra)
library(knitr)
library(mip)
library(plotly)
library(remind2)
library(scales)
library(svglite)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```


```{r setup, echo=FALSE, include=FALSE}

adjustSliderAnimation <- function(p) {
  return(list(p[[1]] %>% plotly::animation_opts(frame = 1)))
}

```


## Convergence summary

The price-anticipation criterion converges when the maximum monetary trade-expenditure
deviation until 2100 falls below **10% of the trade-good surplus tolerance**. The dashed
line is that threshold; green points pass, red points fail.

```{r summary, results = "asis"}

threshold <- NA_real_
mt <- mip::getPlotData("p80_surplusMaxTolerance", params$gdx)
if (nrow(mt) > 0 && "good" %in% mt$all_enty) {
  threshold <- 0.1 * mt$p80_surplusMaxTolerance[mt$all_enty == "good"]
}

df <- mip::getPlotData("p80_DevPriceAnticipGlobAllMax2100Iter", params$gdx)

if (nrow(df) == 0) {
  cat("p80_DevPriceAnticipGlobAllMax2100Iter is empty in this gdx.")
} else {
  df <- df %>%
    rename("value" = "p80_DevPriceAnticipGlobAllMax2100Iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "converged" = ifelse(!is.na(threshold) & .data$value > threshold, "no", "yes"),
      "tooltip" = paste0("Iteration: ", .data$iteration,
                         "<br>Max deviation: ", round(.data$value, 5), " T$",
                         if (all(!is.na(threshold))) paste0("<br>threshold: ", round(threshold, 5)) else "")
    )

  p <- ggplot(df, aes_(x = ~iteration, y = ~value, group = 1, text = ~tooltip)) +
    { if (!is.na(threshold)) geom_hline(yintercept = threshold, linetype = "dashed", color = "grey40") else NULL } +
    geom_line(alpha = 0.6) +
    suppressWarnings(geom_point(aes_(fill = ~converged), size = 2, shape = 21, color = "white")) +
    scale_fill_manual(values = c("yes" = "#00BFC4", "no" = "#F8766D")) +
    scale_y_log10() +
    theme_bw() +
    labs(x = "iteration", y = "max trade expenditure deviation until 2100 [T$]")

  ggplotly(p, tooltip = c("text"))
}

```


## Monetary trade deviation

How much regions' monetary trade expenditure moves purely because each region
*anticipates* its own impact on world prices. At a valid Nash equilibrium this term
should fade to ~0, so across iterations these curves should shrink and flatten.

### Max trade expenditure deviation

- Max deviation of the yearly monetary export/import expenditure due to price change anticipation effect until 2100. [Unit: trillion Dollar]
- p80_DevPriceAnticipGlobMax2100Iter

*One line per traded good plus `all` (the aggregate); x = iteration, log y-axis. Watch
the worst-case anticipation error up to 2100 fall towards zero — a line that plateaus
high is the good whose anticipation is not settling.*

```{r results = "asis"}

df <-
  mip::getPlotData("p80_DevPriceAnticipGlobMax2100Iter", params$gdx)

df.all <-
  mip::getPlotData("p80_DevPriceAnticipGlobAllMax2100Iter", params$gdx) %>%
  mutate(all_enty = "all") %>%
  rename(p80_DevPriceAnticipGlobMax2100Iter = p80_DevPriceAnticipGlobAllMax2100Iter)

df <- rbind(df, df.all)

p <- mip::mipIterations(
  plotData = df,
  returnGgplots = TRUE,
  facetScales = "free_y",
  xAxis = "iteration",
  facets = "all_enty",
  color = NULL,
  slider = NULL
)

# add logarithmic scale
p[[1]] + ggplot2::scale_y_log10(limits = c(1e-4, 1e+1))

```


### Per iteration trade expenditure deviation  

- Per iteration deviation of the yearly monetary export/import expenditure due to price change anticipation effect. [Unit: trillion Dollar]
- p80_DevPriceAnticipGlobIter

*x = model year, one facet per good, log y-axis, with an iteration slider. Advance the
slider and watch the curve collapse toward the axis; periods that stay elevated are
where the anticipation effect lingers.*

```{r results = "asis"}

df <- mip::getPlotData("p80_DevPriceAnticipGlobIter", params$gdx) %>%
  mutate(ttot := as.numeric(ttot))

p <- mip::mipIterations(
  plotData = df,
  xAxis = "ttot",
  facets = "all_enty",
  slider = "iteration",
  facetScales = "free_y",
  returnGgplots = TRUE
)

# add logarithmic scale and then convert to plotly
plots <- p[[1]] + ggplot2::scale_y_log10(limits = c(1e-4, 1e+1))
plots <- list(plots)
plots <- lapply(plots, plotly::ggplotly) %>% adjustSliderAnimation()

htmltools::tagList(plots)

```


## Price

The underlying price signal driving the deviations above.

### Price change due to antecipation

- Price change of a trade good due to the price anticipation effect. [Unit: Percent]
- p80_PriceChangePriceAnticipReg

*Percent price change attributable to anticipation; x = year, one facet per region, one
line per good. Values approaching 0 across the horizon mean anticipation is no longer
distorting prices.*

```{r results = "asis"}

df <-
  mip::getPlotData("p80_PriceChangePriceAnticipReg", params$gdx) %>%
  mutate(ttot := as.numeric(ttot)) %>%
  select(-"iteration")

plots <- mip::mipIterations(
  plotData = df,
  returnGgplots = TRUE,
  xAxis = "ttot",
  facets = "all_regi",
  color = "all_enty",
  slider = NULL,
  facetScales = "free_y"
)

plots[[1]]

```

## Regional monetary trade deviation

The same expenditure deviation, resolved per region so you can see which regions
contribute most.

### Regional trade expenditure deviation

- Regional and per iteration deviation of the yearly monetary export/import expenditure due to price change anticipation effec until 2100. [Unit: trillion Dollar]
- p80_DevPriceAnticipReg

*x = year, one facet per region, one line per good. Larger, more persistent curves mark
the regions and goods most responsible for the remaining anticipation deviation.*

```{r results = "asis"}

df <- mip::getPlotData("p80_DevPriceAnticipReg", params$gdx) %>%
  mutate(ttot := as.numeric(ttot)) %>%
  select(-"iteration")

plots <- mip::mipIterations(
  plotData = df,
  returnGgplots = TRUE,
  xAxis = "ttot",
  facets = "all_regi",
  color = "all_enty",
  slider = NULL,
  facetScales = "free_y"
)

plots[[1]]

```

## Global monetary trade deviation

Globally aggregated anticipation deviation over the whole horizon (not just the max
until 2100), for context on its shape over time.

### Global sum of trade expenditure deviation

- Global sum of deviation of the yearly monetary export/import expenditure due to price change anticipation effect. [Unit: trillion Dollar]
- p80_DevPriceAnticipGlob

*x = year, faceted by good (`all` = sum over goods, `allmax` = its running maximum).
Read the height and persistence over time as the size of the remaining global
anticipation distortion.*

```{r results = "asis"}

df <- mip::getPlotData("p80_DevPriceAnticipGlob", params$gdx) %>%
  mutate(ttot := as.numeric(ttot)) %>%
  select(-"iteration")

df.all <-
  mip::getPlotData("p80_DevPriceAnticipGlobAll", params$gdx) %>%
  mutate(ttot := as.numeric(ttot), all_enty = "all") %>%
  rename(p80_DevPriceAnticipGlob = p80_DevPriceAnticipGlobAll) %>%
  select(-"iteration")

df.allmax <-
  mip::getPlotData("p80_DevPriceAnticipGlobAllMax", params$gdx) %>%
  mutate(ttot := as.numeric(ttot), all_enty = "allmax") %>%
  rename(p80_DevPriceAnticipGlob = p80_DevPriceAnticipGlobAllMax) %>%
  select(-"iteration")

df <- rbind(df, df.all, df.allmax)

plots <- mip::mipIterations(
  plotData = df,
  returnGgplots = TRUE,
  xAxis = "ttot",
  facets = "all_enty",
  color = NULL,
  slider = NULL,
  facetScales = "free_y"
)

plots[[1]]

```

### Max global sum of trade expenditure deviation

- Max global sum of deviation of the yearly monetary export/import expenditure due to price change anticipation effect. [Unit: trillion Dollar]
- p80_DevPriceAnticipGlobMax

*The running maximum of the global sum over the horizon (a monotone envelope), one
facet per good. The lower this envelope sits, the closer the run is to eliminating the
anticipation distortion.*

```{r results = "asis"}

df.max <-
  mip::getPlotData("p80_DevPriceAnticipGlobMax", params$gdx) %>%
  select(-"iteration")

plots <- mip::mipIterations(
  plotData = df.max,
  returnGgplots = TRUE,
  xAxis = "ttot",
  facets = "all_enty",
  color = NULL,
  slider = NULL,
  facetScales = "free_y"
)

plots[[1]]

```

