---
title: "CO2 price convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }

</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(gdx)
library(ggplot2)
library(magclass)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```


```{r setup, echo=FALSE, include=FALSE}

# Value column returned by mip::getPlotData is named after the variable, while
# the set dimensions keep their gdx names ("iteration" and "ttot" here). The
# helpers below abstract the read-and-plot step per dimensionality so each
# variable section stays a one-liner.

# placeholder shown when a variable is absent/empty in this gdx (ggplotly()
# errors on a zero-row data frame, so we substitute an annotated blank plot).
# Also dumps str() of the raw data for debugging why it came back empty.
emptyPlot <- function(var, df = NULL) {
  cat(paste0("str() of raw data for '", var, "':\n"))
  str(df)
  p <- ggplot() +
    annotate("text", x = 0, y = 0, label = paste0("no data for '", var, "' in this gdx")) +
    theme_void()
  ggplotly(p)
}

# only an iteration dimension -> static line/point over iterations
plotOverIterations <- function(var, gdx = params$gdx) {
  df <- mip::getPlotData(var, gdx) %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)))

  if (nrow(df) == 0) return(emptyPlot(var, df))

  p <- ggplot(df, aes(x = .data$iteration, y = .data[[var]])) +
    geom_line(alpha = 0.6) +
    geom_point(size = 1) +
    theme_bw() +
    labs(x = "iteration", y = var)

  ggplotly(p)
}

# combine several iteration-indexed flags into one strip plot: iteration on x,
# one row per variable on y, discrete colour (blue = 0, red = nonzero). Missing
# iterations are filled with 0 so every variable spans the full iteration axis
# (cf. "Max. Trade Surplus" in plotNashConvergence.R).
plotFlipflopStrip <- function(vars, gdx = params$gdx) {
  lastIteration <- readGDX(gdx, "o_iterationNumber", react = "error")[[1]]
  iterations <- data.frame(iteration = seq_len(lastIteration))

  df <- lapply(vars, function(v) {
    d <- mip::getPlotData(v, gdx)
    have <- if (nrow(d) == 0) {
      data.frame(iteration = numeric(0), value = numeric(0))
    } else {
      data.frame(iteration = as.numeric(as.character(d$iteration)), value = d[[v]])
    }
    full <- left_join(iterations, have, by = "iteration")
    full$value[is.na(full$value)] <- 0
    full$variable <- v
    full
  }) %>% bind_rows()

  df$variable <- factor(df$variable, levels = rev(vars))
  df$state <- factor(ifelse(df$value == 0, "0", "1"), levels = c("0", "1"))

  p <- suppressWarnings(ggplot(df, aes(
    x = .data$iteration, y = .data$variable,
    color = .data$state, text = paste0("value: ", .data$value)
  )) +
    geom_point(size = 2, alpha = 0.6) +
    scale_color_manual(values = c("0" = "blue", "1" = "red"), drop = FALSE) +
    scale_y_discrete(drop = FALSE) +
    theme_minimal() +
    labs(x = NULL, y = NULL, color = "value"))

  pl <- ggplotly(p, tooltip = c("x", "y", "text"), height = 120 + 70 * length(vars)) %>%
    hide_legend() %>%
    config(displayModeBar = FALSE, displaylogo = FALSE)

  htmltools::tagList(pl)
}

# only a period dimension -> static line/point over periods
plotOverPeriod <- function(var, gdx = params$gdx) {
  df <- mip::getPlotData(var, gdx) %>%
    mutate("ttot" := as.numeric(as.character(.data$ttot)))

  if (nrow(df) == 0) return(emptyPlot(var, df))

  p <- ggplot(df, aes(x = .data$ttot, y = .data[[var]])) +
    geom_line(alpha = 0.6) +
    geom_point(size = 1) +
    theme_bw() +
    labs(x = "period", y = var)

  ggplotly(p)
}

# used by the animated period-vs-iteration plots below
adjustSliderAnimation <- function(p) {
  return(list(p[[1]] %>% plotly::animation_opts(frame = 1)))
}

# period x iteration -> animated, period on x with a slider over iterations.
# Returns a tagList so the calling chunk can render it directly.
plotOverPeriodBySlider <- function(var, gdx = params$gdx) {
  df <- mip::getPlotData(var, gdx) %>%
    mutate("ttot" := as.numeric(as.character(.data$ttot)))

  if (nrow(df) == 0) return(htmltools::tagList(emptyPlot(var, df)))

  p <- mip::mipIterations(
    plotData = df,
    xAxis = "ttot",
    facets = NULL,
    color = NULL,
    slider = "iteration"
  ) %>% adjustSliderAnimation()

  htmltools::tagList(p)
}

```


## Model configuration

Initial scalar configuration read from the gdx (missing entries show as `NA`).

*These switches define the CO2 target and the price-adjustment algorithm — the context
for every chart below. Check them first to confirm what the run was actually solving.*

```{r config_read}
configVars <- c(
  "cm_budgetCO2from2020", "carbonprice", "cm_taxCO2_functionalForm", "cm_taxCO2_peakBudgYr",
  "cm_peakBudgYr", "cm_taxCO2_IncAfterPeakBudgYr", "cm_nash_autoconverge", "cm_budgetCO2_absDevTol",
  "sm_peakbudget_diff_tolerance", "cm_iterative_target_adj"
)

readScalar <- function(v) {
  x <- suppressWarnings(gdx::readGDX(params$gdx, v, react = "silent"))
  if (is.null(x)) return(NA_character_)
  paste(as.character(as.vector(x)), collapse = ", ")
}

configTable <- data.frame(variable = configVars, value = vapply(configVars, readScalar, character(1)))
```

```{r config_table}
knitr::kable(configTable, row.names = FALSE)
```


## Actual peak emissions year

peak budget year calculated based on maximum of cumulative CO2 emissions, used to check adjustment algorithm [year]

*The year where cumulative CO2 peaks in the final solution. Compare it with the target
`cm_peakBudgYr` above — a mismatch (or a value that keeps jumping between iterations) is
exactly what the peak-budget algorithm is trying to resolve.*

```{r peak_year, results = "asis"}
p45_peakBudgYr_check <- readGDX(params$gdx, "p45_peakBudgYr_check")
allperiods <- getYears(p45_peakBudgYr_check, as.integer = TRUE)
actualPeakBudgYr <- allperiods[p45_peakBudgYr_check == 1]
cat(actualPeakBudgYr)
```


## CO2 tax anchor

The CO2 price path REMIND rescales each iteration to hit the carbon budget. These charts
track the anchor price and its iteration-to-iteration change.

### p45_taxCO2eq_anchor_iter

save p45_taxCO2eq_anchor in each iteration (before entering functionalForm/postsolve.gms) for debugging

<!-- animated: x: period, slider: iteration -->

*x = period, slider = iteration: the CO2 price trajectory as it is re-anchored each
iteration. Advancing the slider, the curve should stop shifting once the budget is met.*

```{r results = "asis"}
plotOverPeriodBySlider("p45_taxCO2eq_anchor_iter")
```

### o45_taxCO2eq_anchor_iterDiff_Itr

track pm_taxCO2eq_anchor_iterationdiff in 2100 over iterations

*x = iteration: how much the 2100 anchor price still changed from the previous
iteration. It should decay toward 0 — a flat non-zero or oscillating line means the
price is not settling.*

```{r}
plotOverIterations("o45_taxCO2eq_anchor_iterDiff_Itr")
```

### p45_taxCO2eq_anchor_iterationdiff_tmp

help parameter for iterative adjustment of taxes

*x = period, slider = iteration: the per-period tax adjustment applied that iteration.
Shrinking magnitudes across the slider indicate convergence.*

```{r results = "asis"}
plotOverPeriodBySlider("p45_taxCO2eq_anchor_iterationdiff_tmp")
```


## CO2 budget and peak budget year

Whether the cumulative-emissions budget and the timing of its peak are being met, and
how they move across iterations.

### o45_diff_to_Budg

Difference between actual CO2 budget and target CO2 budget

*x = iteration; converges when it reaches ~0 (within `cm_budgetCO2_absDevTol`). Positive
= the budget was overshot.*

```{r}
plotOverIterations("o45_diff_to_Budg")
```

### o45_totCO2emi_peakBudgYr

Total CO2 emissions in the peakBudgYr

*x = iteration: emissions in the peak year; watch them settle as the price path
stabilises.*

```{r}
plotOverIterations("o45_totCO2emi_peakBudgYr")
```

### o45_peakBudgYr_Itr

Year in which the CO2 budget is supposed to peak. Is changed in iterative_target_adjust = 9

*x = iteration: the assumed peak year. Back-and-forth jumps here are the flip-flopping
the algorithm tries to damp (see Flip-flop tracking below).*

```{r}
plotOverIterations("o45_peakBudgYr_Itr")
```

### o45_totCO2emi_allYrs

Global CO2 emissions over time and iterations. Needed to check the procedure to find the peakBudgYr

*x = period, slider = iteration: the full emissions trajectory; the peak of this curve
is what determines the peak-budget year.*

<!-- animated: x: period, slider: iteration -->

```{r results = "asis"}
plotOverPeriodBySlider("o45_totCO2emi_allYrs")
```

### o45_change_totCO2emi_peakBudgYr

Measure for how much the CO2 emissions change around the peakBudgYr

*x = iteration: curvature of emissions around the peak year — near-flat means a
well-defined, stable peak.*

```{r}
plotOverIterations("o45_change_totCO2emi_peakBudgYr")
```

## CO2 price rescaling

The multiplicative factors applied to the CO2 price each iteration to steer it toward
the budget. A factor near 1 means little further adjustment is needed.

### p45_factorRescale_taxCO2

Multiplicative factor for rescaling the CO2 price to reach the target

*x = iteration; should converge to 1. Values far from 1, or oscillating around it, mean
the price is still being pushed hard.*

```{r}
plotOverIterations("p45_factorRescale_taxCO2")
```

### p45_factorRescale_taxCO2_Funneled

Multiplicative factor for rescaling the CO2 price to reach the target - limited by an iteration-dependent funnel

*The same factor after the funnel caps how far the price may move in one iteration.
Compare it with the un-funneled factor above to see when the cap is binding.*

```{r}
plotOverIterations("p45_factorRescale_taxCO2_Funneled")
```


## Flip-flop tracking

Iteration on the x-axis, one row per variable; blue = 0, red = nonzero. Missing iterations are filled with 0 (hover for the actual value).

*A red dot flags an iteration where that counter fired — the peak-budget logic detected
or absorbed flip-flopping of the peak year. Long red stretches mean the algorithm is
struggling to pin down a stable peak; all-blue is the healthy case.*

- o45_pkBudgYr_flipflop: Counter that tracks if flipfloping of cm_peakBudgYr occured in the last iterations
- o45_delay_increase_peakBudgYear: Counter that tracks if flip-flopping of peakBudgYr happened. Starts an inner loop to try and overcome this
- o45_reached_until2150pricepath: Counter that tracks if the inner loop of increasing the CO2 price AFTER peakBudgYr goes beyond the initial trajectory
- o45_factorRescale_taxCO2_afterPeakBudgYr: Multiplicative factor for rescaling the CO2 price in the year after peakBudgYr - only needed if flip-flopping of peakBudgYr occurs

```{r results = "asis"}
plotFlipflopStrip(c(
  "o45_pkBudgYr_flipflop",
  "o45_delay_increase_peakBudgYear",
  "o45_reached_until2150pricepath",
  "o45_factorRescale_taxCO2_afterPeakBudgYr"
))
```
