---
title: "Trade convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }

</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gridExtra)
library(knitr)
library(mip)
library(plotly)
library(remind2)
library(scales)
library(svglite)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```


```{r setup, echo=FALSE, include=FALSE}

adjustSliderAnimation <- function(p) {
  return(list(p[[1]] %>% plotly::animation_opts(frame = 1)))
}

mipIterationsSurplus <- function(plotData,
                                 xAxis = "year",
                                 color = NULL,
                                 slider = "iteration",
                                 facets = "region",
                                 facetScales = "fixed",
                                 rectData) {
  p <-
    mip::mipIterations(plotData,
                       returnGgplots = TRUE,
                       xAxis,
                       color,
                       slider,
                       facets,
                       facetScales)[[1]]
  
  p <- p + geom_rect(
    data = rectData,
    aes(
      xmin = min(plotData[xAxis]),
      xmax = max(plotData[xAxis]),
      ymin = p80_surplusMaxTolerance * -1,
      ymax = p80_surplusMaxTolerance,
      text = all_enty
    ),
    fill = "#00BFC4",
    alpha = 0.5,
    inherit.aes = FALSE
  )
  
  p <- plotly::ggplotly(p) %>%
    plotly::animation_opts(frame = 1) %>%
    hide_legend()
  
  htmltools::tagList(p)
}

```

## Surplus convergence history {.tabset}

For each traded good, the worst residual market surplus over iterations (`p80_surplusMax_iter`) is checked against two criteria in REMIND's `postsolve.gms`:
1. **Primary check (until 2100)**: worst surplus up to 2100 against tolerance (`p80_surplusMaxTolerance`).
2. **Secondary check (until 2150)**: worst surplus up to 2150 against 10x tolerance (`10 * p80_surplusMaxTolerance`).

Log y-axis: the good whose line stays above its tolerance longest is the one holding up trade convergence. (Values that have reached exactly zero are dropped, as they cannot be shown on a log scale.)

### Primary check (until 2100, 1x tolerance)

```{r conv_history_2100, results = "asis", fig.height=6}

sm <- mip::getPlotData("p80_surplusMax_iter", params$gdx)

if (nrow(sm) == 0) {
  cat("p80_surplusMax_iter is empty in this gdx.")
} else {
  sm <- sm %>%
    rename("value" = "p80_surplusMax_iter") %>%
    mutate("iteration" := as.numeric(as.character(.data$iteration)),
           "tall" := as.numeric(as.character(.data$tall)))

  horizon <- if (2100 %in% sm$tall) 2100 else max(sm$tall)

  smtol <- mip::getPlotData("p80_surplusMaxTolerance", params$gdx) %>%
    rename("tol" = "p80_surplusMaxTolerance") %>%
    select("all_enty", "tol")

  d <- sm %>%
    filter(.data$tall == horizon) %>%
    left_join(smtol, by = "all_enty") %>%
    filter(.data$value > 0) %>%
    mutate("tooltip" = paste0(.data$all_enty,
                              "<br>Iteration: ", .data$iteration,
                              "<br>Worst surplus until ", horizon, ": ", round(.data$value, 4),
                              "<br>Tolerance (1x): ", round(.data$tol, 4),
                              "<br>", ifelse(.data$value > .data$tol, "outside tolerance", "within tolerance")))

  hlines <- smtol %>% filter(.data$all_enty %in% unique(d$all_enty))

  p <- ggplot(d, aes_(x = ~iteration, y = ~value, group = 1, text = ~tooltip)) +
    geom_hline(data = hlines, aes_(yintercept = ~tol), linetype = "dashed", color = "#F8766D") +
    geom_line(alpha = 0.6) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~all_enty, scales = "free_y") +
    scale_y_log10() +
    theme_bw() +
    labs(x = "iteration", y = paste0("worst surplus until ", horizon, " (log scale)"))

  ggplotly(p, tooltip = c("text"))
}

```

### Secondary check (until 2150, 10x tolerance)

```{r conv_history_2150, results = "asis", fig.height=6}

if (nrow(sm) == 0) {
  cat("p80_surplusMax_iter is empty in this gdx.")
} else {
  horizon2150 <- if (2150 %in% sm$tall) 2150 else max(sm$tall)

  smtol2150 <- smtol %>%
    mutate("tol10" = .data$tol * 10)

  d2150 <- sm %>%
    filter(.data$tall == horizon2150) %>%
    left_join(smtol2150, by = "all_enty") %>%
    filter(.data$value > 0) %>%
    mutate("tooltip" = paste0(.data$all_enty,
                              "<br>Iteration: ", .data$iteration,
                              "<br>Worst surplus until ", horizon2150, ": ", round(.data$value, 4),
                              "<br>10x Tolerance: ", round(.data$tol10, 4),
                              "<br>", ifelse(.data$value > .data$tol10, "outside 10x tolerance", "within 10x tolerance")))

  hlines2150 <- smtol2150 %>% filter(.data$all_enty %in% unique(d2150$all_enty))

  p2150 <- ggplot(d2150, aes_(x = ~iteration, y = ~value, group = 1, text = ~tooltip)) +
    geom_hline(data = hlines2150, aes_(yintercept = ~tol10), linetype = "dashed", color = "#F8766D") +
    geom_line(alpha = 0.6) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~all_enty, scales = "free_y") +
    scale_y_log10() +
    theme_bw() +
    labs(x = "iteration", y = paste0("worst surplus until ", horizon2150, " (log scale)"))

  ggplotly(p2150, tooltip = c("text"))
}

```

## Surplus on commodity market {.tabset}

- p80_surplus

Residual market surplus (supply minus demand) for each traded good; it must fall inside
the tolerance band for trade to converge. The tabs show the same data four ways — most
include the tolerance band as a shaded rectangle, so a good whose line leaves the band
is not yet cleared.

```{r p80_surplus___READ, echo=FALSE, include=FALSE}

p80_surplus <- mip::getPlotData("p80_surplus", params$gdx) %>%
  mutate(tall := as.numeric(tall))
str(p80_surplus)

p80_surplusMaxTolerance <-
  mip::getPlotData("p80_surplusMaxTolerance", params$gdx) %>%
  filter(all_enty %in% unique(p80_surplus[["all_enty"]]))

```

### all goods per iteration
<!-- ### x: time, slider: iter, color: enty -->

*All goods on one chart — x = model year, one line per good, slider over iterations. Watch the lines settle as the run converges.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = p80_surplus,
  xAxis = "tall",
  facets = NULL,
  color = "all_enty",
  slider = "iteration"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per iteration
<!-- ### x: time, slider: iter, facet: enty -->

*One facet per good, x = model year, slider over iterations — the same data separated so each good's own scale is visible.*

```{r}

mipIterationsSurplus(
  plotData = p80_surplus,
  xAxis = "tall",
  facets = "all_enty",
  color = NULL,
  slider = "iteration",
  facetScales = "free_y",
  p80_surplusMaxTolerance
)

```

### good per period
<!-- ### x: iter, slider: time, facet: enty -->

*One facet per good, x = iteration, slider over model years — the convergence view: each curve should flatten as iterations proceed.*

```{r  results = "asis"}

mipIterationsSurplus(
  plotData = filter(p80_surplus, tall >= 2025),
  # for this to work, we starting year must be available for all facets
  xAxis = "iteration",
  slider = "tall",
  color = NULL,
  facets = "all_enty",
  facetScales = "free_y",
  p80_surplusMaxTolerance
)

```

### good per period and iteration
<!-- ### x: iter, facets: time -->

*One facet per model year, x = iteration (no slider) — the full convergence trajectory for every period at once.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = filter(p80_surplus, tall >= 2025),
  returnGgplots = TRUE,
  xAxis = "iteration",
  facets = "tall",
  color = NULL,
  slider = NULL
)

for (i in p) {
  rectData <- p80_surplusMaxTolerance %>%
    filter(all_enty == unique(i$data$all_enty))
  
  i <- i + suppressWarnings(geom_rect(
    data = rectData,
    aes(
      xmin = 0,
      xmax = max(p80_surplus$iteration),
      ymin = p80_surplusMaxTolerance * -1,
      ymax = p80_surplusMaxTolerance,
      text = all_enty
    ),
    fill = "#00BFC4",
    alpha = 0.5,
    inherit.aes = FALSE
  ))
  
  print(i)
}

```


## Trade price {.tabset}

- p80_pvp_itr

The Nash price (Negishi weight) of each traded good over iterations. As surpluses clear,
these prices should stop moving between iterations. Same four views as above.

```{r p80_pvp_itr___READ, echo=FALSE, include=FALSE}

p80_pvp_itr <- mip::getPlotData("p80_pvp_itr", params$gdx) %>%
  mutate(ttot := as.numeric(ttot)) %>%
  filter(ttot >= 2005)
str(p80_pvp_itr)

```

### all goods per iteration
<!-- ### x: time, slider: iter, color: enty -->

*All goods on one chart — x = model year, one line per good, slider over iterations. Watch the lines settle as the run converges.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = p80_pvp_itr,
  xAxis = "ttot",
  facets = NULL,
  color = "all_enty",
  slider = "iteration"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per iteration
<!-- ### x: time, slider: iter, facet: enty -->

*One facet per good, x = model year, slider over iterations — the same data separated so each good's own scale is visible.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = p80_pvp_itr,
  xAxis = "ttot",
  facets = "all_enty",
  color = NULL,
  slider = "iteration",
  facetScales = "free_y"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per period
<!-- ### x: iter, slider: time, facet: enty -->

*One facet per good, x = iteration, slider over model years — the convergence view: each curve should flatten as iterations proceed.*

```{r  results = "asis"}

p <- mip::mipIterations(
  filter(p80_pvp_itr, ttot >= 2025),
  # for this to work, we starting year must be available for all facets
  xAxis = "iteration",
  slider = "ttot",
  color = NULL,
  facets = "all_enty",
  facetScales = "free_y"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per period and iteration
<!-- ### x: iter, facets: time -->

*One facet per model year, x = iteration (no slider) — the full convergence trajectory for every period at once.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = filter(p80_pvp_itr, ttot >= 2025),
  returnGgplots = TRUE,
  xAxis = "iteration",
  facets = "ttot",
  color = NULL,
  slider = NULL,
  facetScales = "free_y"
)

for (i in p) {
  print(i)
}

```

## Price not discounted {.tabset}

The trade prices above re-expressed without the intertemporal discount factor (each good
relative to the `good` numeraire), so price *levels* are comparable across periods. Same
four views.

```{r results = "asis", echo=FALSE, include=FALSE}

price_not_discounted_itr <- left_join(
  filter(p80_pvp_itr, all_enty != "good"),
  filter(p80_pvp_itr, all_enty == "good"),
  by = c("ttot", "iteration")
) %>%
  mutate(all_enty = all_enty.x,
         p80_pvp_itr_no_discount = p80_pvp_itr.x / p80_pvp_itr.y) %>%
  select("ttot", "iteration", "all_enty", "p80_pvp_itr_no_discount")

# adjust "good" value: p80_pvp_itr(t, "good",iteration) * (1.05 ** (t.val - 2005) )

price_not_discounted_itr_good <- p80_pvp_itr %>%
  filter(all_enty == "good") %>%
  mutate("p80_pvp_itr_no_discount" := .data$p80_pvp_itr * 1.05 ** (.data$ttot - 2005)) %>%
  select(-"p80_pvp_itr")

price_not_discounted_itr <-
  rbind(price_not_discounted_itr, price_not_discounted_itr_good)

```

### all goods per iteration
<!-- ### x: time, slider: iter, color: enty -->

*All goods on one chart — x = model year, one line per good, slider over iterations. Watch the lines settle as the run converges.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = price_not_discounted_itr,
  xAxis = "ttot",
  facets = NULL,
  color = "all_enty",
  slider = "iteration"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per iteration
<!-- ### x: time, slider: iter, facet: enty -->

*One facet per good, x = model year, slider over iterations — the same data separated so each good's own scale is visible.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = price_not_discounted_itr,
  xAxis = "ttot",
  facets = "all_enty",
  color = NULL,
  slider = "iteration",
  facetScales = "free_y"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```

### good per period
<!-- ### x: iter, slider: time, facet: enty -->

*One facet per good, x = iteration, slider over model years — the convergence view: each curve should flatten as iterations proceed.*

```{r  results = "asis"}

p <- mip::mipIterations(
  filter(price_not_discounted_itr, ttot >= 2025),
  # for this to work, we starting year must be available for all facets
  xAxis = "iteration",
  slider = "ttot",
  color = NULL,
  facets = "all_enty",
  facetScales = "free_y"
) %>% adjustSliderAnimation()

htmltools::tagList(p)

```


### good per period and iteration
<!-- ### x: iter, facets: time -->

*One facet per model year, x = iteration (no slider) — the full convergence trajectory for every period at once.*

```{r  results = "asis"}

p <- mip::mipIterations(
  plotData = filter(price_not_discounted_itr, ttot >= 2025),
  returnGgplots = TRUE,
  xAxis = "iteration",
  facets = "ttot",
  color = NULL,
  slider = NULL,
  facetScales = "free_y"
)

for (i in p) {
  print(i)
}

```

## Prices and surplus in one plot {.tabset}

Surplus and (undiscounted) price overlaid for each good, so you can see cause and effect
together: when the price stops moving the surplus should close. The `iter+1/+2/+3`
colours show the same series shifted forward, to check whether a price change in one
iteration clears the surplus a few iterations later.

```{r  results = "asis", echo=FALSE, include=FALSE}

prices_and_surplus <-
  left_join(p80_surplus,
            price_not_discounted_itr,
            by = c("tall" = "ttot", "all_enty", "iteration")) %>%
  filter(!is.na(p80_pvp_itr_no_discount)) %>%
  tidyr::pivot_longer(c("p80_surplus", "p80_pvp_itr_no_discount"), names_to = "variable")

# calculate iteration +1 and +2

.step <- function(data, i) {
  data %>%
    filter(.data$iteration > i) %>%
    mutate("iteration" := .data$iteration - i,
           "variable" := paste0(.data$variable, "iter+", i)) %>%
    return()
}

prices_and_surplus_steps <- rbind(
  prices_and_surplus,
  .step(prices_and_surplus, 1),
  .step(prices_and_surplus, 2),
  .step(prices_and_surplus, 3)
) %>%
  mutate(
    "group" = ifelse(
      grepl("^p80_surplus", .data$variable),
      "p80_surplus",
      "p80_pvp_itr_no_discount"
    ),
    "step" = gsub(
      "^(p80_surplus|p80_pvp_itr_no_discount) ?",
      "",
      .data$variable
    ),
    "step" = ifelse(.data$step == "", "iter", .data$step)
  ) %>%
  relocate(where(is.numeric), .after = last_col()) %>%
  select(-"variable")

```

### goods per iteration (1)
<!-- ### x: time, slider: iter, facet: var -->

```{r  results = "asis"}

for (v in unique(prices_and_surplus_steps$all_enty)) {
  p <- mip::mipIterations(
    plotData = filter(prices_and_surplus_steps, all_enty == v, tall >= 2005),
    returnGgplots = TRUE,
    xAxis = "tall",
    facets = "group",
    color = "step",
    slider = "iteration",
    facetScales = "free_y"
  )
  
  # manually override colors
  plots <- p[[1]] +
    ggplot2::scale_colour_manual(
      values = c(
        "iter" = "#000066",
        "iter+1" = "#39418d",
        "iter+2" = "#7b8cba",
        "iter+3" = "#bdd7e7"
      )
    )
  
  # add surplus tolerance as background
  dataRect <- filter(p80_surplusMaxTolerance, all_enty == v) %>%
    mutate(group = "p80_surplus")
  
  plots <- plots + suppressWarnings(geom_rect(
    data = dataRect,
    aes(
      xmin = 2005,
      xmax = max(p80_surplus$tall),
      ymin = p80_surplusMaxTolerance * -1,
      ymax = p80_surplusMaxTolerance,
      text = all_enty
    ),
    fill = "#00BFC4",
    inherit.aes = FALSE,
    alpha = 0.2
  ))
  
  plots <-
    plotly::ggplotly(plots) %>% plotly::animation_opts(frame = 1)
  
  print(htmltools::tagList(plots))
}

```

### good per iteration (2)
<!-- ### x: time, slider: iter, facet: var + all_enty -->

```{r  results = "asis"}

df <- prices_and_surplus_steps %>%
  filter(.data$all_enty != "peur", .data$tall >= 2005) %>%
  mutate("group" = paste0(.data$all_enty, "-", .data$group)) %>%
  select(-"all_enty")

dataRect <- p80_surplusMaxTolerance %>%
  mutate("group" = paste0(.data$all_enty, "-p80_surplus")) %>%
  filter(group %in% unique(df$group))

p <- ggplot(df) +
  geom_line(aes_string(
    x = "tall",
    y = "value",
    color = "step",
    frame = "iteration"
  )) +
  facet_wrap("group", scales = "free_y", ncol = 2) +
  ggplot2::scale_color_manual(
    breaks = c("iter", "iter+1", "iter+2", "iter+3"),
    values = c("#000066", "#39418d", "#7b8cba", "#bdd7e7")
  ) +
  suppressWarnings(geom_rect(
    data = dataRect,
    aes(
      xmin = min(df$tall),
      xmax = max(df$tall),
      ymin = p80_surplusMaxTolerance * -1,
      ymax = p80_surplusMaxTolerance,
      text = all_enty
    ),
    fill = "#00BFC4",
    inherit.aes = FALSE,
    alpha = 0.2
  )) +
  theme_bw() +
  theme(strip.background = element_blank())

plotly::ggplotly(p, autosize = FALSE, height = 800) %>%
  plotly::animation_opts(frame = 1)

```

### good per period
<!-- ### x: iter, slider: time, facet: var -->

```{r  results = "asis"}

for (v in unique(prices_and_surplus$all_enty)) {
  dataRect <- p80_surplusMaxTolerance %>%
    filter(all_enty == v) %>%
    mutate(variable = "p80_surplus")
  
  df <- filter(prices_and_surplus, all_enty == v, tall >= 2025)
  
  p <- mip::mipIterations(
    df,
    xAxis = "iteration",
    facets = "variable",
    color = NULL,
    slider = "tall",
    facetScales = "free_y",
    returnGgplots = TRUE
  )[[1]] +
    suppressWarnings(geom_rect(
      data = dataRect,
      aes(
        xmin = min(df$iteration),
        xmax = max(df$iteration),
        ymin = p80_surplusMaxTolerance * -1,
        ymax = p80_surplusMaxTolerance,
        text = all_enty
      ),
      fill = "#00BFC4",
      inherit.aes = FALSE,
      alpha = 0.2
    ))
  
  p <- plotly::ggplotly(p) %>%
    plotly::animation_opts(frame = 1)
  
  
  print(htmltools::tagList(p))
}

```

### good per period and iteration
<!-- ### x: iter, facet: time, color: var -->

```{r  results = "asis", fig.width=12, fig.height=6}

prices_and_surplus_scaled <- left_join(p80_surplus,
                                       price_not_discounted_itr,
                                       by = c("tall" = "ttot", "all_enty", "iteration")) %>%
  filter(!is.na(p80_pvp_itr_no_discount), tall >= 2025)

for (v in setdiff(unique(prices_and_surplus_scaled$all_enty), "good")) {
  dfv <- prices_and_surplus_scaled %>% filter(all_enty == v)
  plots <- list()
  
  for (t in unique(dfv$tall)) {
    df <- filter(dfv, tall == t)
    
    secScaleDiff <-
      max(df$p80_pvp_itr_no_discount) - min(df$p80_pvp_itr_no_discount)
    secScaleMin <-
      min(df$p80_pvp_itr_no_discount) - 0.05 * secScaleDiff
    secScaleMax <-
      max(df$p80_pvp_itr_no_discount) + 0.05 * secScaleDiff
    
    p <- ggplot(df) +
      geom_line(aes(x = iteration, y = p80_surplus), color = "#00BFC4") +
      geom_line(aes(
        x = iteration,
        y = rescale(p80_pvp_itr_no_discount, to = c(min(p80_surplus), max(p80_surplus)))
      ), color = "#F8766D") +
      scale_y_continuous("", sec.axis = sec_axis( ~ rescale(., to = c(
        secScaleMin, secScaleMax
      )), name = "")) +
      theme(axis.title.x = element_blank(),
            plot.title = element_text(size = 7)) +
      ggtitle(t)
    
    plots <- append(plots, list(p))
  }
  
  do.call("grid.arrange", c(plots, ncol = 4, top = v))
  cat('<span style="color:#00BFC4">primary y-axis: p80_surplus</span><br/>')
  cat('<span style="color:#F8766D">secondary y-axis: p80_pvp_itr_no_discount</span>')
  
}

```
