---
title: "Regional emission target convergence"
output:
  html_document:
    toc: false
    toc_float: true
    toc_depth: 5
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 16
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
  .code-folding-btn {
    display: none;
  }
  .collapsible-section {
    position: relative;
    margin-bottom: 8px;
  }
  .collapse-btn {
    position: absolute;
    top: 5px;
    right: 5px;
    z-index: 100;
    background: rgba(255,255,255,0.9);
    border: 1px solid #ccc;
    border-radius: 3px;
    cursor: pointer;
    font-size: 11px;
    padding: 2px 6px;
    line-height: 1.4;
    color: #555;
  }
  .collapse-btn:hover {
    background: #f0f0f0;
  }
</style>

<script type="text/javascript">
$(document).ready(function() {
  var $tocItems = $(".tocify-subheader li.tocify-item"); // selector for all TOC items
  var $tabs     = $("a[role=\"tab\"]");                  // selector for all tabs
  var $panels   = $("div[role=\"tabpanel\"]");           // selector for all tabpanels

  $tocItems.hide();  // hide all TOC items

  // get the tab name for each section header (e.g. mammals)
  // and assign it to its corresponding toc item
  $panels.find("div[data-unique]").each(function() {
    var key = $(this).attr("data-unique");
    var tab = $(this).closest("div[role=\"tabpanel\"]").attr("id");
    $tocItems.filter("[data-unique=\"" + key + "\"]").attr("tab", tab)
  })
  // now each toc item knows to which tab panel it is pointing

  // show the toc items that point to sections on the first panel
  $tocItems.filter(function() {
    return($(this).attr("tab") === $tabs.first().text());
  }).toggle();

  // assign an onclick event to the tabs..
  $tabs.on("click", function() {
    $tocItems.hide();  // ... hide all TOC items
    var key = $(this).text(); // ... get the name of the tab clicked
    $tocItems.filter(function() {  // ... filter for the toc items pointing to that tab
      return($(this).attr("tab") === key);
    }).toggle();  // ... and show them
  });
});

function toggleSection(btn) {
  var content = btn.nextElementSibling;
  if (content.style.display === 'none') {
    content.style.display = '';
    btn.innerHTML = '&#9660;';
  } else {
    content.style.display = 'none';
    btn.innerHTML = '&#9654;';
  }
}
</script>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(plotly)
library(mip)
library(gdx)
library(reactable)
library(htmltools)
library(tidyr)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r data, echo=FALSE, include=FALSE}

pm_emiMktTarget_tolerance <-
  mip::getPlotData("pm_emiMktTarget_tolerance", params$gdx)
pm_emiMktTarget_dev <-
  mip::getPlotData("pm_emiMktTarget_dev", params$gdx)
pm_factorRescaleemiMktCO2Tax <-
  mip::getPlotData("pm_factorRescaleemiMktCO2Tax", params$gdx)
pm_emiMktCurrent <- mip::getPlotData("pm_emiMktCurrent", params$gdx)
p47_emiMktCurrent_iter <-
  mip::getPlotData("p47_emiMktCurrent_iter", params$gdx)
pm_emiMktTarget <- mip::getPlotData("pm_emiMktTarget", params$gdx)
pm_emiMktRefYear <- mip::getPlotData("pm_emiMktRefYear", params$gdx)
pm_emiMktTarget_dev_iter <-
  mip::getPlotData("pm_emiMktTarget_dev_iter", params$gdx)
pm_taxemiMkt_iteration <-
  mip::getPlotData("pm_taxemiMkt_iteration", params$gdx)
pm_taxemiMkt <- mip::getPlotData("pm_taxemiMkt", params$gdx)
regi <- gdx::readGDX(params$gdx, "regi")
ext_regi <- gdx::readGDX(params$gdx, "ext_regi")
regi_groupExt <- gdx::readGDX(params$gdx, "regi_groupExt")
regi_group <- gdx::readGDX(params$gdx, "regi_group")
emiMktGroup <- gdx::readGDX(params$gdx, "emiMktGroup")

tgts <-
  pm_emiMktTarget %>% select(ext_regi, ttot2, emiMktExt) %>% mutate_at(c('ttot2'), as.numeric) %>% mutate(group = paste0(ttot2, "_", emiMktExt))
tgtRegions <- unique(tgts$ext_regi)

p47_targetConverged_iter <-
  mip::getPlotData("p47_targetConverged_iter", params$gdx)
p47_emiTargetMkt_iter <-
  mip::getPlotData("p47_emiTargetMkt_iter", params$gdx)
p47_factorRescaleSlope_iter <-
  mip::getPlotData("p47_factorRescaleSlope_iter", params$gdx)
p47_factorRescaleemiMktCO2Tax_iter <-
  mip::getPlotData("p47_factorRescaleemiMktCO2Tax_iter", params$gdx)
#p47_ClampedRescaleSlope <- mip::getPlotData("p47_ClampedRescaleSlope", params$gdx)
regiEmiMktRescaleType <-
  gdx::readGDX(params$gdx, "regiEmiMktRescaleType")

# How each target converged. The algorithm guarantees EXACTLY ONE label per target-iteration, in the
# precedence order  unmetAtCap > unmetFrozen > bestAchievable > lowerThanTolerance  (smallPrice is exclusive),
# so the last recorded label is the verdict and nothing here needs to tie-break.
#   lowerThanTolerance  inside the raw user tolerance                      - no residual
#   smallPrice          price at the floor, emissions BELOW target         - no residual (over-achieved)
#   bestAchievable      frozen/held outside the tolerance but inside the exit band - CARRIES A RESIDUAL
#   unmetFrozen         frozen outside the tolerance                       - CARRIES A RESIDUAL
#   unmetAtCap          still outside the tolerance at the iteration cap   - CARRIES A RESIDUAL
# The label describes the DEVIATION, not the machinery that produced it; whether a target was deliberately
# abandoned is p47_giveUp, surfaced separately in the summary table.
regiEmiMktconvergenceType <- suppressWarnings(gdx::readGDX(params$gdx, "regiEmiMktconvergenceType"))
convTypeByIter <- if (!is.null(regiEmiMktconvergenceType) && nrow(regiEmiMktconvergenceType) > 0) {
  .ct <- regiEmiMktconvergenceType
  if ("ttot1" %in% names(.ct)) .ct <- .ct %>% select(-ttot) %>% rename(ttot2 = ttot1)
  .typeCol <- if ("convergenceType" %in% names(.ct)) "convergenceType" else
    tail(setdiff(names(.ct), c("iteration", "ttot2", "ext_regi", "emiMktExt", "value", "ttot")), 1)
  .ct %>%
    select(-any_of("value")) %>%
    mutate(iteration = as.integer(as.character(iteration))) %>%
    group_by(iteration, ttot2, ext_regi, emiMktExt) %>%
    summarise(conv_type = paste(sort(unique(.data[[.typeCol]])), collapse = " + "), .groups = "drop")
} else {
  data.frame(iteration = integer(), ttot2 = character(), ext_regi = character(),
             emiMktExt = character(), conv_type = character(), stringsAsFactors = FALSE)
}
bestAchByTarget <- if (nrow(convTypeByIter) > 0) {
  convTypeByIter %>%
    group_by(ext_regi, ttot2, emiMktExt) %>%
    summarise(is_bestAch = any(grepl("bestAchievable|unmetFrozen", conv_type)), .groups = "drop") %>%
    mutate(ttot2 = as.numeric(as.character(ttot2)))
} else {
  data.frame(ext_regi = character(), ttot2 = numeric(), emiMktExt = character(),
             is_bestAch = logical(), stringsAsFactors = FALSE)
}
# "unmetAtCap": the run ended by hitting the iteration cap with this target still outside its tolerance. The
# algorithm records it at the last iteration and drops any stale lowerThanTolerance label, so a target that was
# certified in-band earlier and then drifted (or is oscillating) can no longer be read as a clean convergence.
unmetAtCapByTarget <- if (nrow(convTypeByIter) > 0) {
  convTypeByIter %>%
    group_by(ext_regi, ttot2, emiMktExt) %>%
    summarise(is_unmet = grepl("unmetAtCap|unmetFrozen", conv_type[which.max(iteration)]), .groups = "drop") %>%
    mutate(ttot2 = as.numeric(as.character(ttot2)))
} else {
  data.frame(ext_regi = character(), ttot2 = numeric(), emiMktExt = character(),
             is_unmet = logical(), stringsAsFactors = FALSE)
}
# a market is "non-binding" when it ended at the carbon-price floor (smallPrice on the final iteration):
# emissions are already below its target so the price collapsed to ~0 and carries no information - the
# region is really steered by its other (binding) market.
nonBindingByTarget <- if (nrow(convTypeByIter) > 0) {
  convTypeByIter %>%
    group_by(ext_regi, ttot2, emiMktExt) %>%
    summarise(is_nonBinding = grepl("smallPrice", conv_type[which.max(iteration)]), .groups = "drop") %>%
    mutate(ttot2 = as.numeric(as.character(ttot2)))
} else {
  data.frame(ext_regi = character(), ttot2 = numeric(), emiMktExt = character(),
             is_nonBinding = logical(), stringsAsFactors = FALSE)
}

emiMktTarget_tolerance <-
  setNames(
    pm_emiMktTarget_tolerance$pm_emiMktTarget_tolerance,
    pm_emiMktTarget_tolerance$ext_regi
  )

tryGetPlotData <- function(name, gdx) {
  tryCatch(mip::getPlotData(name, gdx), error = function(e) data.frame())
}
p47_slopeReferenceIteration_iter     <- tryGetPlotData("p47_slopeReferenceIteration_iter",     params$gdx)

# ── algorithm state added after data8; all read defensively so older gdx files still render ────────────
# THE ZERO-RECORD TRAP: GAMS drops zero records, so p47_marketMet_iter only ever stores 1s. A missing
# (iteration, target) row means NOT met, not "unknown" - joining it and leaving NA would make every target
# look met on every iteration it happened not to be. zeroFill() below is what makes that explicit.
p47_marketMet_iter  <- tryGetPlotData("p47_marketMet_iter",  params$gdx)   # inside the RAW tolerance

# consolidated algorithm: the PERSISTENT per-target state lives in one set-indexed container
# p47_targetState(targetState, ...) with targetState in
# {reopenCount, settledCount, refreshCount, releaseCount, divBrakeCount, aimTries, giveUp, divArmed,
#  rolledBack, reopenIter, bestDevIter, rollFromIter, bestDevAbs, rollFromDev}.
# These were 14 standalone p47_* parameters before 2026-08-04, and every gdx older than that still carries
# them that way, so extractState() falls back to the legacy name - same pattern as extractTrace() below.
p47_targetState <- tryGetPlotData("p47_targetState", params$gdx)
extractState <- function(el, legacyName) {
  if (!is.null(p47_targetState) && nrow(p47_targetState) > 0) {
    typeCol <- names(p47_targetState)[
      sapply(p47_targetState, function(col) any(as.character(col) %in%
        c("reopenCount", "giveUp", "aimTries", "divArmed", "rolledBack")))
    ][1]
    if (!is.na(typeCol)) {
      out <- p47_targetState[as.character(p47_targetState[[typeCol]]) == el, , drop = FALSE]
      if (nrow(out) > 0) {
        out[[typeCol]] <- NULL
        names(out)[names(out) == "p47_targetState"] <- legacyName
        return(out)
      }
    }
  }
  tryGetPlotData(legacyName, params$gdx)   # gdx predates the consolidation
}
p47_giveUp          <- extractState("giveUp",        "p47_giveUp")          # deliberately abandoned
p47_reopenCount     <- extractState("reopenCount",   "p47_reopenCount")     # corrections spent
p47_releaseCount    <- extractState("releaseCount",  "p47_releaseCount")    # uncharged un-freezes
p47_refreshCount    <- extractState("refreshCount",  "p47_refreshCount")    # re-open budget earned back
p47_aimTries        <- extractState("aimTries",      "p47_aimTries")        # attempts at the AIM band
p47_divArmed        <- extractState("divArmed",      "p47_divArmed")        # divergence test live
p47_divBrakeCount   <- extractState("divBrakeCount", "p47_divBrakeCount")   # divergence brakes applied
p47_rolledBack      <- extractState("rolledBack",    "p47_rolledBack")      # 0 none / 1 rolled / 2 undone

# collapse a target-indexed parameter to a named lookup keyed "ext_regi|ttot2|emiMktExt"
targetKeyed <- function(df, valcol) {
  if (is.null(df) || nrow(df) == 0) return(setNames(numeric(0), character(0)))
  d <- df
  if ("ttot1" %in% names(d)) d <- d %>% select(-any_of("ttot")) %>% rename(ttot2 = ttot1)
  if (!"ttot2" %in% names(d) && "ttot" %in% names(d)) d <- d %>% rename(ttot2 = ttot)
  vc <- if (valcol %in% names(d)) valcol else tail(setdiff(names(d), c("ttot2","ext_regi","emiMktExt","ttot")), 1)
  setNames(as.numeric(d[[vc]]),
           paste(d$ext_regi, as.character(d$ttot2), d$emiMktExt, sep = "|"))
}
giveUpByTarget  <- targetKeyed(p47_giveUp,        "p47_giveUp")
reopenByTarget  <- targetKeyed(p47_reopenCount,   "p47_reopenCount")
releaseByTarget <- targetKeyed(p47_releaseCount,  "p47_releaseCount")
refreshByTarget <- targetKeyed(p47_refreshCount,  "p47_refreshCount")
aimByTarget     <- targetKeyed(p47_aimTries,      "p47_aimTries")
armedByTarget   <- targetKeyed(p47_divArmed,      "p47_divArmed")
brakeByTarget   <- targetKeyed(p47_divBrakeCount, "p47_divBrakeCount")
rolledByTarget  <- targetKeyed(p47_rolledBack,    "p47_rolledBack")
lookup0 <- function(tbl, key) { v <- tbl[key]; ifelse(is.na(v), 0, v) }

# consolidated algorithm: per-iteration diagnostics live in one set-indexed container
# p47_slopeTrace_iter(slopeTrace, ...) with slopeTrace in
# {rawSlope, fitR2, preDamp, rollIter, rollUndo, divBrake, parked}.
# The last four are EVENTS rather than measurements - the iteration a price was rolled back, a rollback was
# undone, a divergence brake fired, or the parked-target stop gave up - and they are the only record of those
# decisions anywhere in the gdx. Split the container back into the individual frames the rest expects.
p47_slopeTrace_iter <- tryGetPlotData("p47_slopeTrace_iter", params$gdx)
extractTrace <- function(tp, valname) {
  if (is.null(p47_slopeTrace_iter) || nrow(p47_slopeTrace_iter) == 0) return(data.frame())
  typeCol <- names(p47_slopeTrace_iter)[
    sapply(p47_slopeTrace_iter, function(col) any(as.character(col) %in% c("rawSlope", "fitR2", "preDamp")))
  ][1]
  if (is.na(typeCol)) return(data.frame())
  out <- p47_slopeTrace_iter[as.character(p47_slopeTrace_iter[[typeCol]]) == tp, , drop = FALSE]
  out[[typeCol]] <- NULL
  names(out)[names(out) == "p47_slopeTrace_iter"] <- valname
  out
}
p47_clampedRescaleSlope_iter             <- extractTrace("rawSlope", "p47_clampedRescaleSlope_iter")
p47_slopeFitR2_iter                      <- extractTrace("fitR2",    "p47_slopeFitR2_iter")
p47_dampedFactorRescaleemiMktCO2Tax_iter <- extractTrace("preDamp",  "p47_dampedFactorRescaleemiMktCO2Tax_iter")
# backward compatibility: fall back to the legacy separate parameters for older gdx files
if (nrow(p47_clampedRescaleSlope_iter) == 0)
  p47_clampedRescaleSlope_iter <- tryGetPlotData("p47_clampedRescaleSlope_iter", params$gdx)
if (nrow(p47_slopeFitR2_iter) == 0)
  p47_slopeFitR2_iter <- tryGetPlotData("p47_slopeFitR2_iter", params$gdx)
if (nrow(p47_dampedFactorRescaleemiMktCO2Tax_iter) == 0)
  p47_dampedFactorRescaleemiMktCO2Tax_iter <- tryGetPlotData("p47_dampedFactorRescaleemiMktCO2Tax_iter", params$gdx)

# minR2 threshold for the R2 chart (from the config container; legacy scalar as fallback)
slopeParamData <- tryGetPlotData("p47_slopeParam", params$gdx)
minR2_threshold <- tryCatch({
  if (!is.null(slopeParamData) && nrow(slopeParamData) > 0) {
    typeCol <- names(slopeParamData)[sapply(slopeParamData, function(col) any(as.character(col) == "minR2"))][1]
    slopeParamData[[grep("^p47_slopeParam$|value", names(slopeParamData), value = TRUE)[1]]][as.character(slopeParamData[[typeCol]]) == "minR2"][1]
  } else NA_real_
}, error = function(e) NA_real_)
if (is.na(minR2_threshold)) minR2_threshold <- 0.5

# legacy clamp parameters (present only in older gdx files from the adaptive-clamp algorithm;
# empty data.frames for the consolidated algorithm, which the downstream joins skip gracefully)
p47_slopeUpperClampBound_iter <- tryGetPlotData("p47_slopeUpperClampBound_iter", params$gdx)
p47_upperClampActive_iter     <- tryGetPlotData("p47_upperClampActive_iter", params$gdx)

```

```{r initializeCharts, echo=FALSE, include=FALSE}

#aesthetics
booleanColor <- c("yes" = "#00BFC4", "no" = "#F8766D")
qttyPriceColor <- c("qtty" = "#00BFC4", "price" = "#F8766D")
methodPalette <- c(
  # consolidated (least-squares slope) algorithm
  "slope"     = "#1f78b4", "sq·spread" = "#984ea3", "sq·fit"   = "#a65628",
  # shared / legacy (adaptive-clamp algorithm) codes kept for backward compatibility
  "sq·1st"   = "#e41a1c", "sq·match" = "#ff7f00", "sq·small" = "#f7c842",
  "sq·win"   = "#984ea3", "sq·+sl"   = "#a65628", "sq·clamp" = "#f781bf",
  "sq·degen" = "#aaaaaa",
  "sl·ref"   = "#1f78b4", "sl·1st"   = "#377eb8", "sl·rep"   = "#00BFC4",
  # iterations with no rescale method recorded (target frozen after convergence, or another period being solved)
  "converged"= "#93c47d", "inactive" = "#d9d9d9"
)

g <- NULL

```



<!-- emissions and prices -->
```{r emiAndPrice chart, echo=FALSE, include=FALSE}

#Convert regi to ext_regi and vice-versa
convGroupExt <- NULL
for (tgtReg in unique(tgts$ext_regi)) {
  if (tgtReg %in% regi)
    convGroupExt <- c(convGroupExt, setNames(tgtReg, tgtReg))
  else
    convGroupExt <-
      c(convGroupExt,
        setNames(
          regi_groupExt %>% filter(ext_regi == tgtReg,!(all_regi %in% tgtReg)) %>% slice(1) %>% pull(all_regi),
          tgtReg
        ))
}

emissions <- p47_emiMktCurrent_iter %>%
  rename(emissions = p47_emiMktCurrent_iter) %>%
  mutate(emissions = emissions * 3.66667 * 1000) %>%
  select(-ttot)

convEmiMktGroup <-
  setNames(emiMktGroup$all_emiMktExt, emiMktGroup$all_emiMkt)
tmpPrice <- pm_taxemiMkt_iteration %>%
  mutate(pm_taxemiMkt_iteration = pm_taxemiMkt_iteration * 1000 * 12 / 44) %>%
  filter(all_regi %in% convGroupExt)
price <- rbind(
  tmpPrice %>%
    mutate(
      all_emiMktExt = convEmiMktGroup[all_emiMkt],
      ext_regi = setNames(names(convGroupExt), convGroupExt)[all_regi]
    ),
  tmpPrice %>%
    filter(all_emiMkt == "ETS") %>%
    mutate(
      all_emiMktExt = "all",
      ext_regi = setNames(names(convGroupExt), convGroupExt)[all_regi]
    )
) %>%
  rename(ttot2 = ttot,
         #ext_regi = all_regi,
         emiMktExt = all_emiMktExt) %>%
  select(-all_emiMkt,-all_regi)

plotData <-
  left_join(
    left_join(
      left_join(
        left_join(
          left_join(
            left_join(
              emissions,
              price,
              by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
            ),
            regiEmiMktRescaleType %>% select(-ttot) %>% rename(ttot2 = ttot1) %>% mutate(iteration = as.integer(iteration)),
            by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
          ),
          p47_factorRescaleSlope_iter %>% select(-ttot),
          by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
        ),
        p47_factorRescaleemiMktCO2Tax_iter %>% select(-ttot),
        by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
      ),
      p47_targetConverged_iter %>% rename(ttot2 = ttot),
      by = c("iteration", "ttot2", "ext_regi")
    ),
    pm_emiMktTarget_dev_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  mutate(
    tooltip = paste0(
      "Iteration : ",
      iteration,
      "<br>",
      "Emissions : ",
      round(emissions),
      " Mt CO2<br>",
      "Price : ",
      round(pm_taxemiMkt_iteration),
      " US$2005/t CO2<br>",
      "Convergence : ",
      ifelse(
        is.na(p47_targetConverged_iter),
        ifelse(
          is.na(rescaleType),
          "not active",
          paste0(
            "not converged<br>",
            "Deviation: ",
            round(100 * pm_emiMktTarget_dev_iter, 2),
            "%",
            ifelse(pm_emiMktTarget_dev_iter >= 0, " > ", " < -"),
            round(100 * emiMktTarget_tolerance[ext_regi], 2),
            "%<br>",
            "Rescale for next iteration: ",
            round(p47_factorRescaleemiMktCO2Tax_iter, 2),
            "<br>",
            "Rescale Type : ",
            ifelse(
              grepl("^slope", rescaleType),
              paste0(
                rescaleType,
                " (slope=",
                round(p47_factorRescaleSlope_iter, 2),
                ")"
              ),
              rescaleType
            )
          )
        ),
        "converged"
      )
    )
  )

g[["qtty_and_price"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    # iterate over the group string (period_market), so a region with several markets
    # in the same period (e.g. DEU ETS + ESR @2030) gets one chart per market instead
    # of both markets interleaved into a single (sawtoothing) price trace.
    subGroups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(subGroups, subGroups), function(grp) {
      tgtPeriod <- as.numeric(sub("_.*$", "", grp))
      tgtMkt    <- sub("^[0-9]+_", "", grp)
      d <- plotData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt)
      # aimed target as an emission level (same transform as the emissions series) -> horizontal target line
      tgtEmi <- pm_emiMktTarget %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt) %>%
        slice(1) %>% pull(pm_emiMktTarget)
      tgtEmi <- if (length(tgtEmi) == 1 && !is.na(tgtEmi)) tgtEmi * 3.66667 * 1000 else NA_real_
      emiVals <- d %>% pull(emissions)
      emiHi <- max(c(emiVals, if (!is.na(tgtEmi)) tgtEmi), na.rm = TRUE) * 1.05
      emiLo <- min(c(0, emiVals, if (!is.na(tgtEmi)) tgtEmi), na.rm = TRUE)
      # dark-gray target line: neutral reference that contrasts with both the teal emissions and red price lines
      tgtShapes <- if (!is.na(tgtEmi)) list(list(type = "line", xref = "paper", x0 = 0, x1 = 1,
        yref = "y", y0 = tgtEmi, y1 = tgtEmi, line = list(color = "#404040", dash = "dot", width = 1.5))) else list()
      tgtAnn <- if (!is.na(tgtEmi)) list(list(text = "target", xref = "paper", x = 0.99, xanchor = "right",
        yref = "y", y = tgtEmi, yanchor = "bottom", showarrow = FALSE,
        font = list(size = 10, color = "#404040"), bgcolor = "rgba(255,255,255,0.6)")) else list()
      secAxisScale <-
        max(d %>% pull(pm_taxemiMkt_iteration)) / max(d %>% pull(emissions))
      
      p <- plot_ly(data = d, height = 350) %>% #
        add_trace(
          x = ~ iteration,
          y = ~ emissions,
          name = "Emissions",
          mode = "lines",
          type = "scatter",
          line = list(color = qttyPriceColor["qtty"]),
          hoverinfo = "none",
          hoverinfo = 'skip'
        ) %>%
        add_trace(
          x = ~ iteration,
          y = ~ pm_taxemiMkt_iteration,
          name = "Price",
          yaxis = "y2",
          mode = "lines",
          type = "scatter",
          line = list(color = qttyPriceColor["price"]),
          hoverinfo = "none",
          hoverinfo = 'skip'
        ) %>%
        # add trace only for hover info (hack to disable line in hovermode = "x unified"
        add_trace(
          x = ~ iteration,
          y = ~ emissions,
          mode = "lines",
          type = "scatter",
          hoverinfo = "text",
          line = list(width = 0),
          hovertext = d$tooltip
        ) %>%
        layout(
          hovermode = "x unified",
          showlegend = FALSE,
          margin = list(r = 60),
          shapes = tgtShapes,
          annotations = tgtAnn,
          xaxis = list(
            zeroline = FALSE,
            range = c(0.5, max(d$iteration) + 0.5),
            title = "Iteration",
            gridcolor = "#e8e8e8",
            mirror = TRUE,
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          ),
          yaxis = list(
            zeroline = TRUE,
            zerolinecolor = "#e8e8e8",
            zerolinewidth = 3,
            title = list(text = "Emissions (Mt CO2)", font = list(color =
                                                                    qttyPriceColor["qtty"])),
            range = c(emiLo, emiHi),
            gridcolor = "#e8e8e8",
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          ),
          yaxis2 = list(
            zeroline = TRUE,
            zerolinecolor = "#e8e8e8",
            zerolinewidth = 3,
            showgrid = FALSE,
            title = list(text = "Price (US$2005/t CO2)", font = list(color =
                                                                       qttyPriceColor["price"])),
            overlaying = "y",
            side = "right",
            range = c(emiLo * secAxisScale, emiHi * secAxisScale),
            ticks = 'outside',
            showline = TRUE,
            linecolor = "#adadad",
            linewidth = 2#, tickcolor = "#e8e8e8", tickwidth = 3
          )
        )
    })
  })


```


 <!-- deviation -->
```{r deviation chart, echo=FALSE, include=FALSE}

# deviation chart
plotData <- pm_emiMktTarget_dev_iter %>%
  mutate(
    converged = ifelse((.data$pm_emiMktTarget_dev_iter <=  emiMktTarget_tolerance[ext_regi]) &
                         (.data$pm_emiMktTarget_dev_iter >= -emiMktTarget_tolerance[ext_regi]),
                       "yes",
                       "no"
    ),
    tooltip = paste0(
      "Iteration: ",
      .data$iteration,
      "<br>",
      ifelse((.data$pm_emiMktTarget_dev_iter <=  emiMktTarget_tolerance[ext_regi]) &
               (.data$pm_emiMktTarget_dev_iter >= -emiMktTarget_tolerance[ext_regi]),
             paste0(
               "Converged<br>",
               "Deviation: ",
               round(100 * .data$pm_emiMktTarget_dev_iter, 2),
               "%"
             ),
             paste0(
               "Not converged<br>",
               "Deviation: ",
               round(100 * .data$pm_emiMktTarget_dev_iter, 2),
               "%",
               ifelse(pm_emiMktTarget_dev_iter >= 0, " > ", " < -"),
               round(100 * emiMktTarget_tolerance[ext_regi], 2),
               "%"
             )
      )
    )
  )

g[["deviation"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    # one chart per market (period_market), see qtty_and_price note above
    subGroups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(subGroups, subGroups), function(grp) {
      tgtPeriod <- as.numeric(sub("_.*$", "", grp))
      tgtMkt    <- sub("^[0-9]+_", "", grp)
      d <- plotData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt)
      p <-
        ggplot(d,
               aes_string(x = "iteration", y = "pm_emiMktTarget_dev_iter")) +
        suppressWarnings(geom_rect(
          aes(
            xmin = .data$iteration - 0.5,
            xmax = .data$iteration + 0.5,
            ymin = emiMktTarget_tolerance[ext_regi] * -1,
            ymax = emiMktTarget_tolerance[ext_regi],
            fill = converged,
            text = tooltip
          ),
          alpha = 0.2
        )) +
        suppressWarnings(geom_line(aes_string(group = 1, text = "tooltip"))) +
        scale_y_continuous(
          breaks = c(
            -5 * emiMktTarget_tolerance[ext_regi],
            -emiMktTarget_tolerance[ext_regi],
            0,
            emiMktTarget_tolerance[ext_regi],
            5 * emiMktTarget_tolerance[ext_regi]
          ),
          labels = scales::percent_format(accuracy = .1)
        ) +
        scale_fill_manual(values = booleanColor) +
        theme_bw() +
        theme(
          strip.background = element_blank(),
          axis.title.x = element_blank(),
          legend.position = "none"
        ) +
        coord_cartesian(ylim = c(-5 * emiMktTarget_tolerance[reg],+5 * emiMktTarget_tolerance[reg])) +
        ylab("Deviation (%)") +
        scale_x_continuous(expand = c(0, 0),
                           breaks = 1:max(d$iteration)) +
        theme(plot.margin = margin(0, 42, 0, 22, "pt")) # t,r,b,l
      
      tol_frac         <- emiMktTarget_tolerance[[reg]]
      tol_pct          <- round(100 * tol_frac, 2)
      refYear_val      <- pm_emiMktRefYear %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt) %>%
        slice(1) %>% pull(pm_emiMktRefYear)
      tol_MtCO2 <- if (length(refYear_val) > 0 && !is.na(refYear_val))
        round(tol_frac * refYear_val * 3.66667 * 1000, 1) else NA
      tol_label <- paste0("Tolerance: ±", tol_pct, "%",
                          if (!is.na(tol_MtCO2)) paste0(" (±", tol_MtCO2, " MtCO₂)") else "")

      # flag a non-binding market: it converged at the carbon-price floor (smallPrice), so its carbon
      # price carries no information and the region is really steered by its other (binding) market.
      is_nonBinding_here <- nrow(nonBindingByTarget %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt, is_nonBinding)) > 0
      if (is_nonBinding_here) {
        tol_label <- paste0("● Non-binding market · carbon price floored    |    ", tol_label)
      }

      # flag a best-achievable (noise-floor) stop: target "converged" with residual OUTSIDE the band
      is_bestAch_here <- nrow(bestAchByTarget %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt, is_bestAch)) > 0
      annot_color <- if (is_nonBinding_here) "#3b7dd8" else "#666"
      if (is_bestAch_here) {
        residual_pct <- round(100 * (d %>% arrange(iteration) %>% slice(n()) %>% pull(pm_emiMktTarget_dev_iter)), 2)
        tol_label <- paste0("⚠ Best-achievable stop · residual ", residual_pct,
                            "% (outside tolerance)    |    ", tol_label)
        annot_color <- "#e07b39"
      }

      # flag a target that was still outside tolerance when the run hit the iteration cap - this is NOT a
      # convergence, it is where the target happened to stand when the run was stopped.
      is_unmet_here <- nrow(unmetAtCapByTarget %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == tgtMkt, is_unmet)) > 0
      if (is_unmet_here) {
        residual_pct <- round(100 * (d %>% arrange(iteration) %>% slice(n()) %>% pull(pm_emiMktTarget_dev_iter)), 2)
        tol_label <- paste0("✗ NOT converged · iteration cap reached with deviation ", residual_pct,
                            "%    |    ", tol_label)
        annot_color <- "#c0392b"
      }

      plotly::ggplotly(p, tooltip = c("text"), height = 200) %>%
        plotly::layout(annotations = list(list(
          text      = tol_label,
          xref      = "paper", yref = "paper",
          x = 0.99, y = 0.02,
          xanchor   = "right", yanchor = "bottom",
          showarrow = FALSE,
          font      = list(size = 10, color = annot_color),
          bgcolor   = "rgba(255,255,255,0.75)"
        )))
    })
  })


```



```{r rescaleTableData, echo=FALSE, include=FALSE}

rescaleTypeAbbrev <- c(
  # consolidated (least-squares slope) algorithm
  slope                          = "slope",
  squareDev_lowPriceSpread       = "sq·spread",
  squareDev_unstableFit          = "sq·fit",
  # shared / legacy (adaptive-clamp algorithm) codes kept for backward compatibility
  squareDev_firstIteration       = "sq·1st",
  squareDev_perfectMatch         = "sq·match",
  squareDev_smallChange          = "sq·small",
  squareDev_outsideWindow        = "sq·win",
  squareDev_noNonPositiveSlope   = "sq·+sl",
  squareDev_adaptiveClamp        = "sq·clamp",
  squareDev_degenerateSlope      = "sq·degen",
  slope_refIteration             = "sl·ref",
  slope_firstIteration           = "sl·1st",
  slope_repeatPrev_positiveSlope = "sl·rep"
)

safeJoin <- function(base, extra, by, extra_col_rename = NULL) {
  if (nrow(extra) == 0) return(base)
  if (!is.null(extra_col_rename)) extra <- dplyr::rename(extra, !!extra_col_rename)
  dplyr::left_join(base, extra, by = by)
}

tableData <- pm_emiMktTarget_dev_iter %>%
  select(-ttot) %>%
  mutate(iteration = as.integer(iteration)) %>%
  left_join(
    regiEmiMktRescaleType %>%
      select(-ttot) %>% rename(ttot2 = ttot1) %>%
      mutate(iteration = as.integer(iteration)) %>% select(-any_of("value")) %>%
      group_by(iteration, ttot2, ext_regi, emiMktExt) %>%
      summarise(rescaleType = paste(sort(unique(rescaleType)), collapse = " + "), .groups = "drop"),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  left_join(
    p47_targetConverged_iter %>% rename(ttot2 = ttot),
    by = c("iteration", "ttot2", "ext_regi")
  ) %>%
  left_join(
    p47_factorRescaleSlope_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  left_join(
    p47_factorRescaleemiMktCO2Tax_iter %>% select(-ttot),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_slopeReferenceIteration_iter) > 0) p47_slopeReferenceIteration_iter %>% rename(ttot2 = ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi")
  ) %>%
  safeJoin(
    if (nrow(p47_slopeUpperClampBound_iter) > 0) p47_slopeUpperClampBound_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_upperClampActive_iter) > 0) p47_upperClampActive_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_clampedRescaleSlope_iter) > 0) p47_clampedRescaleSlope_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_dampedFactorRescaleemiMktCO2Tax_iter) > 0) p47_dampedFactorRescaleemiMktCO2Tax_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_slopeFitR2_iter) > 0) p47_slopeFitR2_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(convTypeByIter) > 0) convTypeByIter else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  safeJoin(
    if (nrow(p47_marketMet_iter) > 0) p47_marketMet_iter %>% select(-ttot) else data.frame(),
    by = c("iteration", "ttot2", "ext_regi", "emiMktExt")
  ) %>%
  arrange(ext_regi, ttot2, emiMktExt, iteration) %>%
  group_by(ext_regi, ttot2, emiMktExt) %>%
  mutate(delta_dev_raw = pm_emiMktTarget_dev_iter - lag(pm_emiMktTarget_dev_iter)) %>%
  ungroup() %>%
  mutate(
    method     = sapply(strsplit(rescaleType, " + ", fixed = TRUE), function(types) {
      codes <- rescaleTypeAbbrev[types]
      paste(codes[!is.na(codes)], collapse = " + ")
    }),
    dev_pct    = round(100 * pm_emiMktTarget_dev_iter, 2),
    converged  = ifelse(!is.na(p47_targetConverged_iter) & p47_targetConverged_iter == 1, "✓", "✗"),
    used_slope = round(p47_factorRescaleSlope_iter, 3),
    final_rsc  = round(p47_factorRescaleemiMktCO2Tax_iter, 3),
    delta_dev  = round(abs(100 * delta_dev_raw), 2)
  )

# MET answers "is the target actually achieved", measured against the RAW tolerance with no hysteresis.
# CONVERGED answers "should this target keep steering its carbon price". They differ exactly where a target is
# frozen while unmet - which several give-up branches do deliberately - so showing only the second is how a run
# reports success while sitting far outside its tolerance.
# A missing value is the zero record GAMS dropped, i.e. NOT met - never "unknown".
tableData$met <- if ("p47_marketMet_iter" %in% names(tableData)) {
  ifelse(!is.na(tableData$p47_marketMet_iter) & tableData$p47_marketMet_iter == 1, "✓", "✗")
} else {
  NA_character_
}
tableData$slope_ref  <- if ("p47_slopeReferenceIteration_iter" %in% names(tableData)) tableData$p47_slopeReferenceIteration_iter else NA_real_
tableData$raw_slope  <- if ("p47_clampedRescaleSlope_iter" %in% names(tableData)) round(tableData$p47_clampedRescaleSlope_iter, 3) else NA_real_
tableData$fit_r2     <- if ("p47_slopeFitR2_iter" %in% names(tableData)) round(tableData$p47_slopeFitR2_iter, 3) else NA_real_
tableData$clamp_bnd  <- if ("p47_slopeUpperClampBound_iter" %in% names(tableData)) round(tableData$p47_slopeUpperClampBound_iter, 3) else NA_real_
tableData$clamp_act  <- if ("p47_upperClampActive_iter" %in% names(tableData)) {
  ifelse(is.na(tableData$p47_upperClampActive_iter), NA_character_,
         ifelse(tableData$p47_upperClampActive_iter == 1, "✓", "—"))
} else {
  NA_character_
}
tableData$pre_damp   <- if ("p47_dampedFactorRescaleemiMktCO2Tax_iter" %in% names(tableData)) round(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter, 3) else NA_real_
tableData$damped     <- if ("p47_dampedFactorRescaleemiMktCO2Tax_iter" %in% names(tableData)) {
  ifelse(is.na(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter), NA_character_,
         ifelse(abs(tableData$p47_dampedFactorRescaleemiMktCO2Tax_iter - tableData$p47_factorRescaleemiMktCO2Tax_iter) > 1e-4, "✓", "—"))
} else {
  NA_character_
}

# flag best-achievable (noise-floor) convergence distinctly from a clean tolerance hit
tableData$conv_type <- if ("conv_type" %in% names(tableData)) ifelse(is.na(tableData$conv_type), "", tableData$conv_type) else ""
# ✗ = outside the tolerance and not steering any more; ≈ = accepted short of the tolerance but inside the
# exit band. Both carry a residual and neither may be shown as a clean ✓.
tableData$converged <- ifelse(grepl("unmetAtCap", tableData$conv_type), "✗",
                       ifelse(grepl("unmetFrozen", tableData$conv_type), "✗",
                       ifelse(grepl("bestAchievable", tableData$conv_type), "≈", tableData$converged)))
tableData$conv_type <- ifelse(grepl("unmetAtCap", tableData$conv_type), "unmet@cap",
                       ifelse(grepl("unmetFrozen", tableData$conv_type), "unmet-frozen",
                       ifelse(grepl("bestAchievable", tableData$conv_type), "best-ach",
                       ifelse(grepl("smallPrice", tableData$conv_type), "small-price",
                       ifelse(grepl("lowerThanTolerance", tableData$conv_type), "in-tol", "")))))

tableData <- tableData %>%
  select(iteration, ttot2, ext_regi, emiMktExt,
         method, rescaleType, dev_pct, delta_dev, converged, met, conv_type,
         slope_ref, raw_slope, fit_r2, clamp_bnd, clamp_act, used_slope,
         pre_damp, final_rsc, damped)

legendRow <- function(code, fullname, what, when) {
  htmltools::tags$tr(
    style = "border-bottom: 1px solid #f0f0f0;",
    htmltools::tags$td(style = "padding: 2px 8px; font-family: monospace; white-space: nowrap;", code),
    htmltools::tags$td(style = "padding: 2px 8px; font-family: monospace; white-space: nowrap; color: #888; font-size: 0.85em;", fullname),
    htmltools::tags$td(style = "padding: 2px 8px;", what),
    htmltools::tags$td(style = "padding: 2px 8px; color: #666;", when)
  )
}

makeRescaleTable <- function(data, reg, tgtPeriod, emiMktExtVal) {
  d <- data %>%
    filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
    arrange(iteration)

  tol_frac    <- emiMktTarget_tolerance[[reg]]
  tol_pct     <- round(100 * tol_frac, 2)
  refYear_val <- pm_emiMktRefYear %>%
    filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
    slice(1) %>% pull(pm_emiMktRefYear)
  tol_MtCO2 <- if (length(refYear_val) > 0 && !is.na(refYear_val))
    round(tol_frac * refYear_val * 3.66667 * 1000, 1) else NA
  caption_text <- paste0(
    "Tolerance: ±", tol_pct, "%",
    if (!is.na(tol_MtCO2)) paste0(" (±", tol_MtCO2, " MtCO2)") else ""
  )

  # ── parameter metadata ────────────────────────────────────────────────────
  param_order  <- c("method","dev_pct","delta_dev","converged","met","conv_type",
                    "slope_ref","raw_slope","fit_r2","used_slope",
                    "pre_damp","final_rsc","damped")
  param_labels <- c(
    method     = "Method",
    dev_pct    = "Deviation (%)",
    delta_dev  = "|Δdev| (%)",
    converged  = "Converged",
    met        = "Met (raw tol.)",
    conv_type  = "Conv. type",
    slope_ref  = "Ref. iter.",
    raw_slope  = "Raw slope",
    fit_r2     = "Fit R²",
    used_slope = "Used slope",
    pre_damp   = "Pre-damp factor",
    final_rsc  = "Applied factor",
    damped     = "Damped"
  )
  param_groups <- c(
    method="Convergence", dev_pct="Convergence", delta_dev="Convergence", converged="Convergence", met="Convergence", conv_type="Convergence",
    slope_ref="Slope", raw_slope="Slope", fit_r2="Slope", used_slope="Slope",
    pre_damp="Rescale Factor", final_rsc="Rescale Factor", damped="Rescale Factor"
  )
  group_colors <- c(Convergence="#4a9eca", Slope="#e07b39", `Rescale Factor`="#5cb85c")

  # ── transpose: rows = parameters, cols = iterations ───────────────────────
  iter_cols <- as.character(d$iteration)
  d_t <- d %>%
    select(iteration, all_of(param_order)) %>%
    mutate(across(everything(), as.character)) %>%
    tidyr::pivot_longer(cols = -iteration, names_to = "param", values_to = "value") %>%
    tidyr::pivot_wider(names_from = iteration, values_from = value) %>%
    mutate(group = param_groups[param], label = param_labels[param]) %>%
    arrange(match(param, param_order)) %>%
    select(group, param, label, all_of(iter_cols))

  # ── per-iteration lookups for cell styling ─────────────────────────────────
  method_per_iter    <- setNames(d$method,    iter_cols)
  converged_per_iter <- setNames(d$converged, iter_cols)

  # ── column definitions ─────────────────────────────────────────────────────
  label_coldef <- list(
    group = reactable::colDef(show = FALSE),
    param = reactable::colDef(show = FALSE),
    label = reactable::colDef(
      name  = "",
      width = 145,
      style = function(value, index) {
        list(fontWeight  = "600",
             fontSize    = "0.82em",
             borderLeft  = paste0("3px solid ", group_colors[d_t$group[index]]),
             paddingLeft = "6px",
             whiteSpace  = "nowrap")
      }
    )
  )

  iter_coldefs <- lapply(setNames(iter_cols, iter_cols), function(iter) {
    method  <- method_per_iter[[iter]]
    is_conv <- identical(converged_per_iter[[iter]], "✓")
    is_sq   <- !is.na(method) && grepl("^sq", method)

    reactable::colDef(
      name  = iter,
      width = 46,
      align = "center",
      html  = TRUE,
      header = htmltools::div(
        style = paste0("font-weight:bold; font-size:0.88em;",
                       if (is_conv) " color:#c0c0c0;" else ""),
        iter
      ),
      cell = function(value, index) {
        if (is.na(value) || value == "NA")
          return(htmltools::span(style = "color:#e8e8e8;", "—"))
        pname <- d_t$param[index]
        if (pname == "method")
          gsub(" \\+ ", "<br><span style='color:#aaa; font-size:0.85em;'>+</span>",
               value, perl = TRUE)
        else if (pname == "converged")
          if (value == "✓")
            htmltools::span(style = "color:#00BFC4; font-weight:bold;", value)
          else if (value == "≈")
            htmltools::span(title = "best-achievable (noise-floor) stop — residual is outside tolerance",
                            style = "color:#e07b39; font-weight:bold;", value)
          else if (value == "✗")
            htmltools::span(title = "iteration cap reached with this target still outside its tolerance — not converged",
                            style = "color:#c0392b; font-weight:bold;", value)
          else
            htmltools::span(style = "color:#F8766D; font-weight:bold;", value)
        else if (pname == "conv_type")
          if (value == "best-ach")
            htmltools::span(style = "color:#e07b39; font-weight:bold;", value)
          else if (value == "unmet@cap")
            htmltools::span(style = "color:#c0392b; font-weight:bold;", value)
          else
            htmltools::span(style = "color:#999; font-size:0.85em;", value)
        else
          value
      },
      style = function(value, index) {
        pname <- d_t$param[index]
        is_slope_row <- param_groups[pname] == "Slope"
        if (is.na(value) || value == "NA") return(list(color = "#ebebeb"))
        if (is_conv)  return(list(color = "#c8c8c8", fontSize = "0.78em"))
        if (is_slope_row && is_sq) return(list(color = "#d0d0d0"))
        NULL
      }
    )
  })

  tbl <- reactable::reactable(
    d_t,
    columns  = c(label_coldef, iter_coldefs),
    rowStyle = function(index) {
      base <- list(fontSize = "0.84em", lineHeight = "1.4")
      if (index > 1 && d_t$group[index] != d_t$group[index - 1])
        base <- c(base, list(borderTop = "2px solid #d8d8d8", marginTop = "4px"))
      base
    },
    sortable   = FALSE,
    bordered   = FALSE,
    compact    = TRUE,
    highlight  = TRUE,
    pagination = FALSE,
    theme = reactable::reactableTheme(
      headerStyle = list(borderBottom = "2px solid #ccc", fontSize = "0.88em",
                         background = "#eeeeee")
    )
  )

  legend <- htmltools::div(
    style = "font-size:0.8em; color:#444; margin-top:8px;",
    htmltools::tags$details(
      htmltools::tags$summary(
        style = "cursor:pointer; font-weight:bold; color:#555; padding:4px 0;",
        "Method legend"
      ),
      htmltools::tags$table(
        style = "border-collapse:collapse; margin-top:6px; width:100%;",
        htmltools::tags$thead(
          htmltools::tags$tr(
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "Code"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "Full name"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "What it does"),
            htmltools::tags$th(style = "text-align:left; padding:3px 8px; border-bottom:2px solid #ccc;", "When it fires")
          )
        ),
        htmltools::tags$tbody(
          legendRow("slope",     "slope",                          "Rescale = (target − current) / (slope × price) + 1, capped to the trust region [0.5, 2]", "Consolidated algorithm — windowed least-squares slope is negative and well-fit (R² ≥ s47_slopeMinR2): the normal Newton path"),
          legendRow("sq·spread", "squareDev_lowPriceSpread",       "Rescale = (1 + dev)²",                                            "Consolidated algorithm — carbon price barely moved across the window, so the slope is ill-conditioned (replaces perfectMatch / smallChange / outsideWindow)"),
          legendRow("sq·fit",    "squareDev_unstableFit",          "Rescale = (1 + dev)²",                                            "Consolidated algorithm — windowed slope is non-negative (wrong sign) or the fit is poor (R² < s47_slopeMinR2), so it is unreliable for a Newton step (replaces repeatPrev / noNonPositiveSlope)"),
          legendRow("sq·1st",    "squareDev_firstIteration",       "Rescale = (1 + dev)²",                                            "First iteration, or slope window just restarted — no slope history yet"),
          legendRow("sq·match",  "squareDev_perfectMatch",         "Rescale = (1 + dev)²",                                            "Tax identical to both reference and iteration 1 — division-by-zero guard"),
          legendRow("sq·small",  "squareDev_smallChange",          "Rescale = (1 + dev)²",                                            "Tax change vs both reference and iteration 1 is < 1e−2 — avoids stagnation from negligible price variation"),
          legendRow("sq·win",    "squareDev_outsideWindow",        "Rescale = (1 + dev)²",                                            "Reference near-zero denominator AND using iteration 1 would exceed the slope window"),
          legendRow("sq·+sl",    "squareDev_noNonPositiveSlope",   "Rescale = (1 + dev)²",                                            "Slope is positive (price rise → more emissions) AND no prior slope to repeat — sign assumption violated"),
          legendRow("sq·clamp",  "squareDev_adaptiveClamp",        "Rescale = (1 + dev)²; slope reference reset",                     "Adaptive Upper Clamp Bound fired twice consecutively at same level — slope-based convergence exhausted"),
          legendRow("sq·degen",  "squareDev_degenerateSlope",      "Rescale = (1 + dev)²; slope reference reset",                     "Emission change negligible relative to 2005 emissions — degenerate slope from near-net-zero starting point"),
          legendRow("sl·ref",    "slope_refIteration",             "Rescale = (target − current) / (slope × price) + 1",         "Normal path — slope computed against the current Slope Reference Iteration"),
          legendRow("sl·1st",    "slope_firstIteration",           "Rescale = (target − current) / (slope × price) + 1",         "Reference denominator near-zero, but iteration 1 is within slope window — uses iteration 1 as baseline"),
          legendRow("sl·rep",    "slope_repeatPrev_positiveSlope", "Rescale = (target − current) / (slope × price) + 1 (prev.)", "Computed slope is positive, but prior slope exists with same reference — repeats previous slope to avoid sign anomaly")
        )
      )
    )
  )

  htmltools::tagList(
    tbl,
    legend
  )
}

```

```{r rescaleFactorChart, echo=FALSE, include=FALSE}

g[["rescale_factor"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>% arrange(iteration)

      hasDamp <- !all(is.na(d$pre_damp))
      p <- plot_ly(data = d, height = 280) %>%
        add_segments(x = min(d$iteration) - 0.5, xend = max(d$iteration) + 0.5,
                     y = 1, yend = 1, showlegend = FALSE,
                     line = list(color = "#aaaaaa", dash = "dot", width = 1)) %>%
        add_trace(x = ~iteration, y = ~final_rsc,
                  name = "Final rescale", mode = "lines+markers", type = "scatter",
                  line = list(color = "#5cb85c", width = 2),
                  marker = list(color = "#5cb85c", size = 6),
                  hovertemplate = "Iter %{x}<br>Final: %{y:.3f}<extra></extra>")

      if (hasDamp) {
        d_damp <- d %>% filter(!is.na(pre_damp) & abs(pre_damp - final_rsc) > 1e-4)
        if (nrow(d_damp) > 0) {
          p <- p %>% add_trace(data = d_damp,
                               x = ~iteration, y = ~pre_damp,
                               name = "Pre-damp", mode = "markers", type = "scatter",
                               marker = list(color = "#e07b39", size = 10, symbol = "circle-open",
                                             line = list(width = 2, color = "#e07b39")),
                               opacity = 0.7,
                               hovertemplate = "Iter %{x}<br>Pre-damp: %{y:.3f}<extra></extra>")
        }
      }

      p %>% layout(
        hovermode = "x unified", showlegend = TRUE,
        legend = list(orientation = "h", y = -0.18),
        xaxis = list(title = "Iteration", zeroline = FALSE,
                     range = c(0.5, max(d$iteration) + 0.5),
                     gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                     showline = TRUE, linecolor = "#adadad", linewidth = 2),
        yaxis = list(title = "Rescale factor", gridcolor = "#e8e8e8",
                     ticks = "outside", showline = TRUE, linecolor = "#adadad", linewidth = 2)
      )
    })
  })

```


```{r phaseDiagramChart, echo=FALSE, include=FALSE}

g[["phase_diagram"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)

      emi_d <- p47_emiMktCurrent_iter %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        mutate(emissions = p47_emiMktCurrent_iter * 3.66667 * 1000) %>%
        select(iteration, emissions)

      price_d <- price %>%
        filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        select(iteration, pm_taxemiMkt_iteration) %>%
        rename(carbon_price = pm_taxemiMkt_iteration)

      d_phase <- emi_d %>%
        inner_join(price_d, by = "iteration") %>%
        left_join(tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod,
                                       emiMktExt == emiMktExtVal) %>%
                    select(iteration, method, converged), by = "iteration") %>%
        arrange(iteration)

      tgt_emi_MtCO2 <- tryCatch({
        v <- pm_emiMktTarget %>%
          filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
          slice(1) %>% pull(pm_emiMktTarget)
        if (length(v) == 1 && !is.na(v)) v * 3.66667 * 1000 else NULL
      }, error = function(e) NULL)

      p <- plot_ly(data = d_phase, height = 350)

      if (!is.null(tgt_emi_MtCO2)) {
        p <- p %>% add_segments(
          x = min(d_phase$carbon_price, na.rm = TRUE),
          xend = max(d_phase$carbon_price, na.rm = TRUE),
          y = tgt_emi_MtCO2, yend = tgt_emi_MtCO2,
          showlegend = FALSE,
          line = list(color = "#e07b39", dash = "dash", width = 1.5)
        )
      }

      d_phase <- d_phase %>%
        mutate(
          primary_method = sub(" \\+.*", "", ifelse(is.na(method) | method == "", NA, method)),
          # iterations with no rescale method: the target was already converged (its price frozen), or a
          # different period was being solved that iteration. Label them by convergence, not "unknown".
          primary_method = ifelse(!is.na(primary_method) & primary_method %in% names(methodPalette), primary_method,
                            ifelse(converged %in% c("✓", "≈"), "converged", "inactive")),
          point_color    = ifelse(primary_method %in% names(methodPalette),
                                  methodPalette[primary_method], "#aaaaaa")
        )

      p <- p %>%
        add_trace(x = ~carbon_price, y = ~emissions,
                  type = "scatter", mode = "lines",
                  line = list(color = "#cccccc", width = 1),
                  showlegend = FALSE, hoverinfo = "skip")

      for (meth in unique(d_phase$primary_method)) {
        d_sub <- d_phase %>% filter(primary_method == meth)
        col   <- if (meth %in% names(methodPalette)) methodPalette[[meth]] else "#aaaaaa"
        p <- p %>% add_trace(
          data = d_sub,
          x = ~carbon_price, y = ~emissions,
          type = "scatter", mode = "markers+text",
          name = meth,
          text = ~as.character(iteration), textposition = "top right",
          textfont = list(size = 9, color = "#555"),
          marker = list(size = 8, color = col),
          customdata = ~paste0("Iter: ", iteration,
                               "<br>Price: ", round(carbon_price), " US$2005/t CO2",
                               "<br>Emi: ", round(emissions), " Mt CO2",
                               "<br>Method: ", ifelse(is.na(method), "—", method)),
          hovertemplate = "%{customdata}<extra></extra>"
        )
      }

      seg_arrows <- if (nrow(d_phase) > 1)
        lapply(seq_len(nrow(d_phase) - 1), function(i)
          list(
            ax       = d_phase$carbon_price[i],
            ay       = d_phase$emissions[i],
            x        = d_phase$carbon_price[i + 1],
            y        = d_phase$emissions[i + 1],
            xref = "x", yref = "y", axref = "x", ayref = "y",
            arrowhead = 2, arrowsize = 0.8, arrowwidth = 1.2,
            arrowcolor = "#aaaaaa",
            showarrow = TRUE, text = ""
          )
        ) else list()

      p %>% layout(
          showlegend = TRUE,
          legend = list(x = 0.98, y = 0.98, xanchor = "right", yanchor = "top",
                        bgcolor = "rgba(255,255,255,0.65)", bordercolor = "#cccccc",
                        borderwidth = 1, font = list(size = 10)),
          annotations = seg_arrows,
          xaxis = list(title = "Carbon price (US$2005/t CO2)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "Emissions (Mt CO2)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2)
        )
    })
  })

```


```{r summaryTargetRow, echo=FALSE, include=FALSE}

# Reusable summary row for one target: the ALREADY-BUILT emissions & carbon-price chart (left) and phase
# diagram (right), side by side. Reuses the g$ objects so there is no chart-code duplication, and both
# already carry the aimed-target line.
summaryTargetRow <- function(reg, grp) {
  htmltools::div(
    style = "display:flex; flex-wrap:nowrap; gap:14px; align-items:flex-start; margin:2px 0 18px 0;",
    htmltools::div(style = "flex:1 1 50%; min-width:0;", g$qtty_and_price[[reg]][[grp]]),
    htmltools::div(style = "flex:1 1 50%; min-width:0;", g$phase_diagram[[reg]][[grp]])
  )
}

```


```{r methodTimelineChart, echo=FALSE, include=FALSE}

g[["method_timeline"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        arrange(iteration) %>%
        mutate(
          primary = sub(" \\+.*", "", method),
          primary = ifelse(!is.na(converged) & converged == "✓", "converged",
                           ifelse(is.na(primary) | nchar(trimws(primary)) == 0, "inactive", primary)),
          color   = ifelse(primary == "inactive", "#eeeeee",
                           ifelse(primary %in% names(methodPalette), as.character(methodPalette[primary]), "#aaaaaa")),
          label = ifelse(is.na(method) | nchar(trimws(method)) == 0, "target not active", method)
        )

      present_methods <- intersect(names(methodPalette), unique(d$primary))

      # Single bar trace — continuous strip, per-bar color from d$color
      p <- plot_ly(height = 70) %>%
        add_bars(
          data = d,
          x = ~iteration, y = rep(1, nrow(d)),
          marker = list(color = d$color, line = list(width = 0)),
          hovertemplate = ~paste0("Iter ", iteration, ": ", label, "<extra></extra>"),
          showlegend = FALSE
        )

      # Circle legend entries — placed off-chart so they appear without fading
      for (meth in names(methodPalette)) {
        if (!meth %in% present_methods) next
        col <- methodPalette[[meth]]
        p <- p %>% add_trace(
          type = "scatter", mode = "markers",
          x = -999L, y = -999,
          name = meth,
          marker = list(color = col, symbol = "circle", size = 8),
          showlegend = TRUE,
          hoverinfo = "skip"
        )
      }

      p %>% layout(
        barmode  = "stack",
        showlegend = TRUE,
        legend = list(
          orientation = "h",
          x = 0.98, xanchor = "right",
          y = 0.45, yanchor = "top",
          font = list(size = 10),
          traceorder = "normal",
          bgcolor = "rgba(0,0,0,0)",
          borderwidth = 0
        ),
        xaxis = list(title = "", zeroline = FALSE,
                     range = c(0.5, max(d$iteration) + 0.5),
                     fixedrange = TRUE,
                     showticklabels = FALSE, showgrid = FALSE),
        yaxis = list(title = "", showticklabels = FALSE, zeroline = FALSE,
                     range = c(0, 1.01), fixedrange = TRUE, showgrid = FALSE),
        margin = list(t = 2, b = 50, l = 80, r = 60)
      ) %>%
        plotly::config(displayModeBar = FALSE)
    })
  })

```


```{r slopeClampChart, echo=FALSE, include=FALSE}

g[["slope_clamp"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>% arrange(iteration)

      p <- plot_ly(data = d, height = 280)

      # legacy adaptive-clamp bound (only present for older gdx files)
      if ("clamp_bnd" %in% names(d) && !all(is.na(d$clamp_bnd))) {
        p <- p %>% add_trace(x = ~iteration, y = ~clamp_bnd,
                             name = "Adaptive clamp bound (legacy)", type = "scatter", mode = "lines",
                             line = list(color = "#e07b39", width = 1.5, dash = "dot", shape = "hv"),
                             hovertemplate = "Iter %{x}<br>Clamp bound: %{y:.3f}<extra></extra>")
      }
      if ("raw_slope" %in% names(d) && !all(is.na(d$raw_slope))) {
        p <- p %>% add_trace(x = ~iteration, y = ~raw_slope,
                             name = "Raw slope (pre-cap)", type = "scatter", mode = "markers",
                             marker = list(color = "#e07b39", size = 9, symbol = "circle-open",
                                           line = list(width = 2, color = "#e07b39")),
                             opacity = 0.7,
                             hovertemplate = "Iter %{x}<br>Raw slope: %{y:.3f}<extra></extra>")
      }

      p <- p %>%
        add_trace(x = ~iteration, y = ~used_slope,
                  name = "Used slope", type = "scatter", mode = "lines+markers",
                  line = list(color = "#4a9eca", width = 2),
                  marker = list(color = "#4a9eca", size = 6),
                  hovertemplate = "Iter %{x}<br>Used slope: %{y:.3f}<extra></extra>") %>%
        add_segments(x = min(d$iteration) - 0.5, xend = max(d$iteration) + 0.5,
                     y = 0, yend = 0, showlegend = FALSE,
                     line = list(color = "#e8e8e8", width = 2))

      # fit R2 on a secondary axis, with the acceptance threshold (only when available)
      hasR2 <- "fit_r2" %in% names(d) && !all(is.na(d$fit_r2))
      if (hasR2) {
        p <- p %>%
          add_trace(x = ~iteration, y = ~fit_r2,
                    name = "Fit R²", type = "scatter", mode = "lines+markers",
                    yaxis = "y2",
                    line = list(color = "#5cb85c", width = 1.5, dash = "dash"),
                    marker = list(color = "#5cb85c", size = 5, symbol = "diamond"),
                    hovertemplate = "Iter %{x}<br>Fit R²: %{y:.3f}<extra></extra>") %>%
          add_segments(x = min(d$iteration) - 0.5, xend = max(d$iteration) + 0.5,
                       y = minR2_threshold, yend = minR2_threshold, yaxis = "y2",
                       showlegend = FALSE, hoverinfo = "skip",
                       line = list(color = "#5cb85c", width = 1, dash = "dot"))
      }

      p %>%
        layout(
          hovermode = "x unified", showlegend = TRUE,
          legend = list(orientation = "h", y = -0.2),
          xaxis = list(title = "Iteration", zeroline = FALSE,
                       range = c(0.5, max(d$iteration) + 0.5),
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "Slope", gridcolor = "#e8e8e8",
                       ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis2 = list(title = list(text = "Fit R²", font = list(color = "#5cb85c")),
                        overlaying = "y", side = "right",
                        range = c(0, 1.02), showgrid = FALSE, zeroline = FALSE,
                        tickfont = list(color = "#5cb85c"))
        )
    })
  })

```


```{r convergenceRateChart, echo=FALSE, include=FALSE}

g[["convergence_rate"]] <-
  lapply(setNames(tgtRegions, tgtRegions), function(reg) {
    groups <- tgts %>% filter(ext_regi == reg) %>% pull(group)
    lapply(setNames(groups, groups), function(grp) {
      tgtPeriod    <- as.numeric(strsplit(grp, "_", fixed = TRUE)[[1]][1])
      emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == grp) %>% pull(emiMktExt)
      d <- tableData %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>%
        arrange(iteration) %>% filter(!is.na(delta_dev)) %>%
        mutate(abs_delta = delta_dev)

      plot_ly(data = d, height = 220) %>%
        add_bars(x = ~iteration, y = ~abs_delta,
                 marker = list(color = "#4a9eca", opacity = 0.85),
                 hovertemplate = "Iter %{x}<br>|Δdev|: %{y:.2f}%<extra></extra>",
                 showlegend = FALSE) %>%
        layout(
          xaxis = list(title = "Iteration", zeroline = FALSE,
                       range = c(0.5, max(d$iteration) + 0.5),
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2),
          yaxis = list(title = "|Δdeviation| (%)", zeroline = FALSE,
                       gridcolor = "#e8e8e8", mirror = TRUE, ticks = "outside",
                       showline = TRUE, linecolor = "#adadad", linewidth = 2)
        )
    })
  })

```


```{r summaryData, echo=FALSE, include=FALSE}

activity_long <- tableData %>%
  mutate(
    target_label = paste(ext_regi, ttot2, emiMktExt, sep = " · "),
    primary      = sub(" \\+.*", "", method),
    primary      = ifelse(!is.na(converged) & converged == "✓", "converged",
                          ifelse(is.na(primary) | nchar(trimws(primary)) == 0, "inactive", primary)),
    cell_color   = ifelse(primary == "inactive", "#eeeeee",
                          ifelse(primary %in% names(methodPalette), as.character(methodPalette[primary]), "#aaaaaa")),
    is_conv = !is.na(converged) & converged == "✓",
    tooltip = ifelse(
      primary == "inactive",
      paste0(
        gsub(" · ", " | ", target_label), " — Iter ", iteration,
        "<br>Method: not active"
      ),
      paste0(
        gsub(" · ", " | ", target_label), " — Iter ", iteration,
        "<br>Method: ", ifelse(is.na(method) | nchar(trimws(method)) == 0, "not active", method),
        "<br>Deviation: ", ifelse(is.na(dev_pct), "—", paste0(dev_pct, "%")),
        "<br>|Δdev|: ", ifelse(is.na(delta_dev), "—", paste0(delta_dev, "%")),
        "<br>Applied factor: ", ifelse(is.na(final_rsc), "—", as.character(final_rsc)),
        "<br>Converged: ", ifelse(is_conv, "✓", "✗")
      )
    )
  )

conv_first <- activity_long %>%
  filter(is_conv) %>%
  group_by(target_label) %>%
  summarise(first_conv = min(iteration), .groups = "drop")

target_summary_df <- tableData %>%
  group_by(ext_regi, ttot2, emiMktExt) %>%
  arrange(iteration) %>%
  summarise(
    n_total_iters   = max(iteration),
    n_active        = sum(!is.na(method)),
    n_conv_iters    = sum(!is.na(converged) & converged == "✓"),
    first_conv_iter = { ci <- iteration[!is.na(converged) & converged == "✓"]
                        if (length(ci) > 0) min(ci) else NA_integer_ },
    last_conv_iter  = { ci <- iteration[!is.na(converged) & converged == "✓"]
                        if (length(ci) > 0) max(ci) else NA_integer_ },
    pct_conv        = round(100 * sum(!is.na(converged) & converged == "✓") /
                            max(1, max(iteration)), 1),
    final_dev_pct   = last(dev_pct),
    final_conv      = last(converged),
    peak_abs_dev    = round(max(abs(dev_pct), na.rm = TRUE), 2),
    n_damped        = sum(!is.na(damped) & damped == "✓"),
    methods_used    = {
      pm <- sub(" \\+.*", "", method[!is.na(method)])
      pm <- pm[nchar(trimws(pm)) > 0]
      if (length(pm) == 0) "—"
      else { mc <- sort(table(pm), decreasing = TRUE)
             paste(paste0(names(mc), " (", as.integer(mc), ")"), collapse = " · ") }
    },
    .groups = "drop"
  ) %>%
  mutate(
    tol_pct         = round(100 * emiMktTarget_tolerance[ext_regi], 2),
    target_label    = paste(ext_regi, ttot2, emiMktExt, sep = " · "),
    in_tol          = !is.na(final_dev_pct) & abs(final_dev_pct) <= tol_pct,
    first_conv_disp = ifelse(is.na(first_conv_iter), "—", as.character(first_conv_iter)),
    last_conv_disp  = ifelse(is.na(last_conv_iter),  "—", as.character(last_conv_iter)),
    # ── what the algorithm actually DID to this target ─────────────────────────────────────────────────
    # A residual on its own does not say whether the algorithm ran out of road, deliberately stopped, or
    # never engaged. These are the counters that answer that, and none of them was visible in this report
    # before. `gave_up` is the important one: it is the difference between "still trying and failing" and
    # "stopped trying on purpose", and the run-end signal depends on it.
    tkey            = paste(ext_regi, as.character(ttot2), emiMktExt, sep = "|"),
    gave_up         = lookup0(giveUpByTarget,  tkey) > 0,
    n_reopen        = lookup0(reopenByTarget,  tkey),
    n_release       = lookup0(releaseByTarget, tkey),
    n_refresh       = lookup0(refreshByTarget, tkey),
    n_aim           = lookup0(aimByTarget,     tkey),
    div_armed       = lookup0(armedByTarget,   tkey) > 0,
    n_brake         = lookup0(brakeByTarget,   tkey),
    rolled          = lookup0(rolledByTarget,  tkey),
    outcome         = ifelse(gave_up, "gave up", ifelse(in_tol, "met", "unmet")),
    state_disp      = paste0(
      ifelse(n_reopen  > 0, paste0("reopen ", n_reopen, " "), ""),
      ifelse(n_refresh > 0, paste0("refresh ", n_refresh, " "), ""),
      ifelse(n_release > 0, paste0("release ", n_release, " "), ""),
      ifelse(n_brake   > 0, paste0("brake ", n_brake, " "), ""),
      ifelse(rolled    > 0, ifelse(rolled >= 2, "rollback undone ", "rollback "), "")
    ),
    state_disp      = ifelse(nchar(trimws(state_disp)) == 0, "—", trimws(state_disp))
  )

```


```{r summaryCharts, echo=FALSE, include=FALSE}

# ── Activity heatmap + legend ─────────────────────────────────────────────
target_order <- target_summary_df %>%
  arrange(ext_regi, desc(ttot2), emiMktExt) %>%
  pull(target_label) %>% rev()

activity_long$target_f <- factor(activity_long$target_label, levels = target_order)

p_act <- ggplot(activity_long,
                aes(x = iteration, y = target_f, fill = cell_color, text = tooltip)) +
  geom_tile(height = 0.82, color = NA, linewidth = 0) +
  scale_fill_identity() +
  theme_minimal() +
  theme(axis.title.y = element_blank(), panel.grid = element_blank(),
        axis.text.y = element_text(size = 9), axis.title.x = element_text(size = 9),
        plot.margin = margin(4, 4, 4, 4, "pt")) +
  labs(x = "Iteration")

act_h <- max(100, length(target_order) * 38 + 90)
act_plotly <- ggplotly(p_act, tooltip = "text", height = act_h) %>%
  layout(
    showlegend = TRUE,
    legend = list(orientation = "h", x = 0.98, xanchor = "right",
                  y = -0.04, yanchor = "top",
                  font = list(size = 10), bgcolor = "rgba(0,0,0,0)", borderwidth = 0),
    shapes = lapply(conv_first$first_conv, function(i)
      list(type = "line", x0 = i - 0.5, x1 = i - 0.5, y0 = 0, y1 = 1,
           yref = "paper", line = list(color = "rgba(0,0,0,0.18)", width = 1, dash = "dot")))
  )

present_in_summary <- intersect(names(methodPalette), unique(activity_long$primary))
for (meth in names(methodPalette)) {
  if (!meth %in% present_in_summary) next
  act_plotly <- act_plotly %>% add_trace(
    type = "scatter", mode = "markers",
    x = -999L, y = target_order[1],
    name = meth,
    marker = list(color = methodPalette[[meth]], symbol = "circle", size = 8),
    showlegend = TRUE, hoverinfo = "skip",
    inherit = FALSE
  )
}
g[["summary_activity"]] <- act_plotly %>%
  layout(xaxis = list(range = c(0.5, max(activity_long$iteration) + 0.5),
                      fixedrange = TRUE))

# ── Final deviation strip ──────────────────────────────────────────────────
dev_h         <- max(140, nrow(target_summary_df) * 42 + 80)
tgt_order_dev <- target_summary_df %>% arrange(ext_regi, ttot2, emiMktExt) %>% pull(target_label)

dev_data <- target_summary_df %>%
  arrange(ext_regi, ttot2, emiMktExt) %>%
  mutate(dot_color = ifelse(in_tol, "#00BFC4", "#F8766D"))

g[["summary_conv_dev"]] <- plot_ly(dev_data, height = dev_h, width = 550) %>%
  add_segments(x = ~-tol_pct, xend = ~tol_pct, y = ~target_label, yend = ~target_label,
               line = list(color = "rgba(0,191,196,0.12)", width = 14),
               showlegend = FALSE, hoverinfo = "skip") %>%
  add_segments(x = 0, xend = ~final_dev_pct, y = ~target_label, yend = ~target_label,
               line = list(color = "#cccccc", width = 1),
               showlegend = FALSE, hoverinfo = "skip") %>%
  add_markers(x = ~final_dev_pct, y = ~target_label,
              marker = list(color = ~dot_color, size = 10),
              text = ~paste0(target_label, "<br>Final dev: ", final_dev_pct, "%",
                             "<br>Tolerance: ±", tol_pct, "%<br>", final_conv),
              hoverinfo = "text", showlegend = FALSE) %>%
  layout(
    shapes = list(list(type = "line", x0 = 0, x1 = 0, y0 = 0, y1 = 1,
                       yref = "paper", line = list(color = "#999", width = 1))),
    xaxis = list(title = "Final deviation (%)", zeroline = FALSE, gridcolor = "#e8e8e8",
                 mirror = TRUE, ticks = "outside", showline = TRUE,
                 linecolor = "#adadad", linewidth = 2),
    yaxis = list(title = "", zeroline = FALSE, gridcolor = "#e8e8e8",
                 categoryorder = "array", categoryarray = rev(tgt_order_dev))
  )

# ── Summary reactable ──────────────────────────────────────────────────────
convCell <- function(v) if (identical(v, "✓"))
  htmltools::span(style = "color:#00BFC4;font-weight:bold;", v) else
  htmltools::span(style = "color:#F8766D;font-weight:bold;", v)

# "met" / "unmet" / "gave up" - the last is NOT a failure of the algorithm but a deliberate decision, and it
# is what allows the run to terminate at all, so it is coloured as its own state rather than as a miss.
outcomeCell <- function(v) {
  col <- switch(v, "met" = "#00BFC4", "gave up" = "#e0a030", "#F8766D")
  htmltools::span(style = paste0("color:", col, ";font-weight:bold;"), v)
}

devCell <- function(v, row) {
  tol <- target_summary_df$tol_pct[row]
  col <- if (!is.na(v) && abs(v) <= tol) "#00BFC4" else "#F8766D"
  htmltools::span(style = paste0("color:", col, ";"), paste0(v, "%"))
}

summary_react <- reactable::reactable(
  target_summary_df %>%
    select(target_label, final_conv, outcome, first_conv_disp, final_dev_pct,
           tol_pct, n_aim, state_disp, n_damped, n_conv_iters, methods_used),
  columns = list(
    target_label    = reactable::colDef(name = "Target",         minWidth = 140),
    final_conv      = reactable::colDef(name = "Conv.",          align = "center", width = 56,
                                        html = TRUE, cell = convCell),
    outcome         = reactable::colDef(name = "Outcome",        align = "center", width = 82,
                                        html = TRUE, cell = outcomeCell),
    first_conv_disp = reactable::colDef(name = "First conv.",    align = "center", width = 80),
    final_dev_pct   = reactable::colDef(name = "Final dev (%)",  align = "right",  width = 95,
                                        cell = devCell),
    tol_pct         = reactable::colDef(name = "Tolerance",      align = "right",  width = 80,
                                        cell = function(v) paste0("±", v, "%")),
    n_aim           = reactable::colDef(name = "Aim tries",      align = "center", width = 76),
    state_disp      = reactable::colDef(name = "Interventions",  minWidth = 150,
                                        style = list(fontSize = "0.82em", color = "#555")),
    n_damped        = reactable::colDef(name = "Damped",         align = "center", width = 70),
    n_conv_iters    = reactable::colDef(name = "Iters conv.",    align = "center", width = 80),
    methods_used    = reactable::colDef(name = "Methods used",
                                        style = list(fontSize = "0.82em", color = "#555"))
  ),
  striped = TRUE, highlight = TRUE, compact = TRUE, pagination = FALSE,
  theme = reactable::reactableTheme(
    headerStyle = list(borderBottom = "2px solid #ccc", fontSize = "0.88em",
                       background = "#eeeeee")
  )
)

```


## {.tabset}

```{r, include=FALSE}

# Creating sections
# short "what it shows / how to read it" note rendered under a section title
note <- function(txt) paste0("<p style='color:#555;font-size:0.9em;margin:2px 0 10px 0;'><em>", txt, "</em></p>\n\n")
# collapsible wrapper around a code chunk that evaluates `expr` (matches the report's collapsible style)
collapsibleChunk <- function(expr, opts = "results='asis'") paste0(
  "<div class='collapsible-section'>\n",
  "<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n",
  "<div class='collapsible-content'>\n\n",
  "```{r , ", opts, "}\n\n", expr, "\n\n```\n\n",
  "</div></div>\n\n")

out <- "### Summary\n\n"

out <- paste0(out, "##### Activity and Convergence Timeline\n\n")
out <- paste0(out, note("When each target was the one actively being solved (colored) and the iteration at which it converged (marker), across the whole run."))
out <- paste0(out, collapsibleChunk("g$summary_activity"))

out <- paste0(out, "##### Final Deviation vs Tolerance\n\n")
out <- paste0(out, note("Each target's final deviation from its target versus its tolerance band (shaded). Bars reaching outside the band are unmet, or accepted as best-achievable at the noise floor."))
out <- paste0(out, collapsibleChunk("g$summary_conv_dev"))

# NEW: per-target overview - one row per target, emissions&price (left) + phase diagram (right), placed
# right before the summary table. Reuses the g$ chart objects via summaryTargetRow (no duplication).
out <- paste0(out, "##### Per-target Overview\n\n")
out <- paste0(out, note("One row per target: emissions &amp; carbon price (left) and the emissions-vs-price phase diagram (right), each with the dashed aimed-target line. A converged target shows emissions settling on the target line and the phase path ending on it."))
out <- paste0(out, "<div class='collapsible-section'>\n<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n<div class='collapsible-content'>\n\n")
for (reg in tgtRegions) {
  for (grp in tgts %>% filter(ext_regi == reg) %>% pull(group)) {
    out <- paste0(out, "<p style='margin:6px 0 0 0;'><b>", reg, " &mdash; ", grp, "</b></p>\n\n")
    out <- paste0(out, "```{r , fig.height=4}\n\nsummaryTargetRow('", reg, "','", grp, "')\n\n```\n\n")
  }
}
out <- paste0(out, "</div></div>\n\n")

out <- paste0(out, "##### Summary Table\n\n")
out <- paste0(out, note("One row per target. <b>Conv.</b> is whether the target still steers its price; <b>Outcome</b> is what actually happened to it &mdash; <span style='color:#00BFC4'>met</span> (inside the raw tolerance), <span style='color:#F8766D'>unmet</span>, or <span style='color:#e0a030'>gave up</span> (a give-up branch deliberately abandoned it, which is what lets the run end). <b>Aim tries</b> is attempts spent closing the last stretch to the aim band; a target at the budget converged through the accept fallback rather than by aiming. <b>Interventions</b> lists what the algorithm had to do: re-opens charged, re-open budget refreshed after a long settled stretch, uncharged releases, divergence brakes, and price rollbacks."))
out <- paste0(out, collapsibleChunk("summary_react"))

for(reg in tgtRegions){
  out <- paste0(out,knitr::knit_expand(text=paste0("### ", reg, " {.tabset}\n\n")))
  for(subGroup in tgts %>% filter(ext_regi == reg) %>% pull(group)){

    tgtPeriod    <- strsplit(subGroup,"_",fixed=T)[[1]][1]
    emiMktExtVal <- tgts %>% filter(ext_regi == reg, group == subGroup) %>% pull(emiMktExt)
    # filter by market too, otherwise a region with >1 market in a period returns a
    # length-2 vector here and the section header is emitted twice.
    unit         <- pm_emiMktTarget %>% filter(ext_regi == reg, ttot2 == tgtPeriod, emiMktExt == emiMktExtVal) %>% pull(emi_type_47)

    out <- paste0(out,knitr::knit_expand(text=paste0("#### ", subGroup, "\n\n")))

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Emissions and Carbon Price (", unit, ")\n\n")))
    out <- paste0(out, note("Emissions (teal, left axis) and the carbon price (red, right axis) per Nash iteration; the dashed orange line is the aimed target. Convergence = emissions settling on the target line. The colored squares below mark the rescale method used each iteration; a deviation-vs-tolerance chart and the per-iteration detail table follow."))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"<div style='margin-bottom:-18px'>\n")
    out <- paste0(out,"```{r , results = 'asis', fig.height=10}\n\n")
    out <- paste0(out,"g$qtty_and_price[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div>\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$method_timeline[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis', fig.height=10}\n\n")
    out <- paste0(out,"g$deviation[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , echo=FALSE}\n\n")
    out <- paste0(out,"makeRescaleTable(tableData, '",reg,"', '",tgtPeriod,"', '",emiMktExtVal,"')\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Rescale Factor\n\n")))
    out <- paste0(out, note("The multiplicative change applied to the carbon price each iteration (&gt;1 raises it, &lt;1 lowers it). Values pinned at the trust-region cap [0.5, 2] mark maximal steps; progressive dampening shrinks the step when the price oscillates."))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$rescale_factor[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Phase Diagram\n\n")))
    out <- paste0(out, note("Emissions vs carbon price, one point per iteration with arrows tracing the path; the dashed line is the aimed target. A path flattening onto the target line means the price is still moving emissions; a near-vertical stall away from it means the price no longer helps (noise floor / infeasible)."))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$phase_diagram[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Slope and Fit R² Evolution\n\n")))
    out <- paste0(out, note("The fitted emissions-vs-price slope used for the Newton price step and its R² (fit quality). Low R² (below the dashed threshold) means the slope is unreliable, so a squareDev fallback step is used that iteration instead."))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$slope_clamp[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")

    out <- paste0(out,knitr::knit_expand(text=paste0("##### Convergence Rate\n\n")))
    out <- paste0(out, note("How fast the deviation shrinks from one iteration to the next. A flat line near zero means the target has stalled - either at its noise floor or because it is (locally) infeasible."))
    out <- paste0(out,"<div class='collapsible-section'>\n")
    out <- paste0(out,"<button class='collapse-btn' onclick='toggleSection(this)' title='Hide/Show'>&#9660;</button>\n")
    out <- paste0(out,"<div class='collapsible-content'>\n\n")
    out <- paste0(out,"```{r , results = 'asis'}\n\n")
    out <- paste0(out,"g$convergence_rate[['", reg, "']][['", subGroup, "']]\n\n")
    out <- paste0(out,"```\n\n")
    out <- paste0(out,"</div></div>\n\n")
  }
}


```

`r paste(knitr::knit(text = paste(out, collapse = '\n')))`


